### 2017-02-23 paper.tex
Initial import -- ~50% of the paper is in there

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

\usepackage[utf8x]{inputenc}
\usepackage{graphicx}
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
{\tt underworlds}: Cascading Situation Assessment for Robots
}

\author{Séverin Lemaignan$^{1}$%
\thanks{$^{1}$Séverin Lemaignan is with Centre for Robotics and Neural Systems,
        Plymouth University, Plymouth PL48AA, United Kingdom
        {\tt\small severin.lemaignan@plymouth.ac.uk}}%
}

\usepackage{xspace}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\etal}{et al.\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\vs}{vs.\xspace}
\newcommand{\resp}{\textit{resp.}\xspace}
\newcommand{\cf}{cf.\xspace}

\newcommand{\uwds}{{\tt underworlds}\xspace}

\graphicspath{{figs/}}

\usepackage[cache]{minted}
\renewcommand{\theFancyVerbLine}{
  \sffamily\textcolor[rgb]{0.5,0.5,0.5}{\scriptsize\arabic{FancyVerbLine}}}

\newminted{python}{frame=lines,
                    linenos=true,
                    fontsize=\scriptsize,
                    xleftmargin=1.8em}

\newmintinline[python]{python}{fontsize=\footnotesize}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

    We introduce hereafter \uwds, a novel framework for \emph{cascading}
    spatio-temporal situation assessment in robotics. \uwds allows programmers
    to represent the robot's environment as real-time distributed data
    structures, containing both scene graphs (for 3D representation of geometries)
    and timelines (for representation of temporal events). \uwds supports
    \emph{cascading} representations: the environment is viewed as a set of
    \emph{worlds} that can each have different spatial and temporal
    granularities, and may depend on each others.

    \uwds also provides a set of high-level client libraries and tools to introspect
    and manipulate the environment models.

    This article presents the design and architecture of this open-source tool,
    and explores some applications, along with examples of use.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\uwds is a distributed and lightweight framework that enables robot programmers
to build and refine in real-time spatial and temporal models of the environment
surrounding a robot. \uwds makes it easy, then, to share these models amongst
the other software components running on the robot, to eventually provide a
distributed environment model for the robot.

The components which require such spatial and temporal models of the environment
are typically found in the intermediate layers of robotic architectures, between
the low-level perceptual layers, and the high-level decisional layers. They
typically include geometric reasoners (that compute spatial and topological
relations between objects), motion planners, gestures/sequence recognizers.

Difficulties however arise quickly. These different components might need
different models, at different spatial and temporal resolutions. A 3D motion
planner would typically use coarse 3D models of surrounding objects to lower the
computational load while planning, while a spatial reasoner might need on the
contrary a high-resolution model of the environment for accurate rendering when
attempting to estimate the visibility of an object. Spatial models are also
usually large as they might contain a lot of geometric data (including point
clouds and meshes). Efficient sharing of such datasets is non-trivial,
particularly when the said data is dynamically acquired (\eg in the case of
map building or object recognition).

Traditional robotic middlewares like ROS are not particularly well-suited to
deal with these complexities: geometric data can be represented, but are not
first-class citizens (basic tasks like displaying a 3D mesh at a dynamic
position are non-trivial with ROS, for instance) and representing and reasoning
on \emph{alternative states} of the environment is not directly doable.

Representing \emph{alternative states} is actually often critical, as shown by
the following exemplary situation: software components manipulating environment
models typically perform better if the models are physically consistent.
However, low-level perception inaccuracies often introduce hard-to-avoid
physical inconsistencies (like objects floating in the air, or on the contrary
wrongly inset into other objects). Therefore, a post-process stage (for
instance, using a physics simulation engine) might be desired to correct the
positions of the objects seen by the robot. Implemented with a classical approach using ROS TF
frames, we would see an object {\tt book} located at the origin of the reference
frame {\tt book\_frame}, and a secondary frame {\tt book\_frame\_corrected} computed by the
physics engine. It is easy to see why such an approach would be quickly
confusing (where is my book?) and would not scale well, with the robot's 3D
model cluttered with spurious frames.

; efficiently sharing a dynamic 3D scene is
even more difficult;  bserving a scene from
different viewpoints to compute visibility is not directly possible

One of the main features of \uwds is the ability to store many parallel
worlds: past models of the environment, future models, models with some objects
filtered out, models that are physically consistent, etc.

This package provides the \uwds server, a Python client library, and a set
of tools to interact with the system (viewers, scene loader, etc.).

A handful of useful example applications are also provided, like skeleton
tracking (using OpenNI) or visibility tracking.

\subsection{Situation Assessment}

Anchoring perceptions in a symbolic model requires perception abilities and
their symbolic interpretation. We call \emph{physical situation assessment} the
cognitive skill that a robot exhibits when it assesses the nature and content of its
surroundings and monitors its evolution.

Numerous approaches exist, like amodal (in the sense of modality-independent)
\emph{proxies}~\cite{Jacobsson2008}, grounded amodal
representations~\cite{Mavridis2006}, semantic
maps~\cite{Nuechter2008, Galindo2008,Blodow2011} or affordance-based planning
and object classification~\cite{Lorken2008, Varadarajan2011}.

\section{Related work}

\cite{sisbot2011situation}
\cite{naef2003blue}
\cite{bustos2016unified}

\uwds is inspired by geometric and temporal reasoning components like {\sc
Spark} (\emph{SPAtial Reasoning \&
Knowledge}~\cite{sisbot2011situation} (see~\cite{lemaignan2016artificial} for a
high-level description of the integration of a component like {\sc Spark} in a
complete robotic architecture).

It acts as a situation assessment reasoner that generates symbolic knowledge from the
geometry of the environment with respect to relations between objects, robots
and humans, also
taking into account the different perspective that each agent has on the
environment.  {\sc Spark} embeds an \emph{amodal} (as defined by Mavridis and
Roy in~\cite{Mavridis2006}: the different perceptual modalities are abstracted
away into a blended spatial model) geometric model of the environment that
serves both as basis for the fusion of the perception modalities and as bridge
with the symbolic layer. This geometric model is built from 3D CAD models of the
objects, furnitures and robots, and full body, rigged models of humans
(Figure~\ref{fig:sparkScreenshot}).  It is updated at run-time by the robot's
sensors (usually, a combination of vision-based tracking of 2D fiducial markers
to identify and localise objects, and Kinect-based skeleton tracking of humans,
optionally assisted by motion capture to accurately track the head motion, which
is required to compute what the human is looking at).

\subsection{Relation to existing robotic middlewares}

\section{Design and Architecture}

\subsection{A layered model of the environment}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{overview}
    \caption{Schema of a \uwds network: eight possible, user-written
    \emph{clients} (in blue) are exchanging environment models through four shared
    \emph{worlds} (made of joint spatial and temporal models). This architecture
    enables successive and modular refinement of the models
    (\emph{cascading} situation assessment), effectively adapted to each
    client's needs (yellow nodes represent other possible components in the system that do not
interact directly with the \uwds network).} \label{fig|scene} \end{figure}

(worlds are effectively composed of a scene graph representing the 3D geometry
of the environment, and a timeline storing temporal events)

\subsection{Distributed spatio-temporal models}
\label{arch}

\uwds is not a monolithic piece of software. Instead, it stands for both a
\emph{network of interconnected clients} which manipulate spatial and temporal
models of the robot environment (for instance, a motion planner, a object
detection module, a human skeleton tracker, etc.), and for a {client library}
that makes it easy to interface existing software components with the network.

Critically, the network is essentially hidden to the client: from the user
perspective, the environment model is manipulated as local data structures (the
\emph{worlds}). Modifications to the model are asynchronously synchronised with
a central server (the {\tt underworlded} daemon) and broadcasted to every other
client in the network.

As previously mentioned, worlds are composite data structures comprising of a
scene graph and a timeline. These data structures are synchronized using
Google's gRPC message passing framework\footnote{http://www.grpc.io/}, ensuring
high throughput, reliability, cross-platform/cross-language support. The \uwds
API is specifically discussed hereafter, section~\ref{api}.

\subsubsection*{Performance considerations}

\uwds is meant to broadcast complex environment representations (typically
including large geometric datasets, like meshes) in real-time. \uwds itself does
not perform many CPU intensive tasks (CPU intensive processing tasks -- sensor fusion, physics
simulation, etc. -- are realized by the clients) and as such, the performance
bottleneck is essentially the network's data throughput.

In that regard, one of the simple yet critical optimization performed by \uwds
is automatic caching of mesh data. Mesh data are not transmitted when nodes are
updated; only a hash value of the mesh data. The client can then request the
full data whenever they actually need it.

\subsection{Cascading architecture}

Figure~\ref{fig|scene} pictures a typically \uwds topology.

\subsection{Worlds}
\subsubsection{Scenes}

\subsubsection{Nodes}

\subsubsection{Timelines}

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{timeline}
    \caption{Both events and episodes are stored in the world's \emph{timeline}}

    \label{fig|timeline}
\end{figure}

\section{API \& Clients}

\subsection{API}
\label{api}

As mentioned in section~\ref{arch}, \uwds uses Google's gRPC as message passing
protocol. The protocol is explicitely defined, and bindings to various languages
and platforms can be automatically generated from the protocol definition file
(as of Feb 2017, gRPC can generate bindings for C, C++, C\#, Node.js, PHP, Ruby,
Python, Go and Java on Windows, Mac and Linux).

The cross-platform/cross-language support of gRPC is especially welcomed in the
academic context, as it offers ease and flexibility to plug a variety of
pre-existing components into a \uwds network.

However, the gRPC message passing layer is low-level with respect to the typical
use of \uwds (manipulation of asynchronous, distributed spatio-temporal models
of the robot environment). In particular, the asynchronous fetching (and
conversely, remote updating) of nodes and time-related objects should be
entirely hidden from the user, and managed instead by the \uwds client library.

\uwds currently offers such a high-level client library for Python only. A C++
library is being developed (currently in 'alpha' state).
Listing~\ref{lst|pythonapi} gives a complete example of an \uwds client
performing simple filtering: the client continuously ``listen'' for changes in
an input world, hide some items (in this case, items whose name contain the
string ``{\tt wall}''), and forward all other changes to an output world,
effectively making the output world a copy of the input world with all walls
removed.

\begin{listing}[h!]

\begin{pythoncode}
import underworlds

with underworlds.Context("walls_filter") as ctx:

    in_world = ctx.worlds["example_in"]
    out_world = ctx.worlds["example_out"]

    while True:

        in_world.scene.waitforchanges()

        for node in in_word.scene.nodes:
            if("wall" not in node.name):
                out_world.scene.nodes.update(node)

\end{pythoncode}
    \caption{Example of a simple \uwds Python client named {\tt walls\_filter}:
    the client connects to the \uwds network, efficiently accesses the world
    {\tt example\_in}, filter out some objects, and publish the remaining
    objects in the world {\tt example\_out}.} \label{lst|pythonapi}
\end{listing}

\subsection{Standard Clients}

\subsubsection{3D Visualization}

Interestingly, while \uwds deals with 3D geometries and scenes, it does
represent 3D entities purely as data structures: no visual representation is
involved (and as such, the \uwds server and core libraries do not depend on any
graphics library like OpenGL). However, for all practical purposes, the ability
to visualize the content of a scene is desirable. \uwds provides a standard
client, {\tt uwds view}, that performs real-time 3D rendering of the scene,
using OpenGL (Figure~\ref{fig|uwds-view}).

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{uwds-screenshot}
    \caption{Screenshot of the {\tt uwds view} 3D visualization and manipulation
    tool.}
    \label{fig|uwds-view}
\end{figure}

This tool also support basic object manipulations, that are broadcasted to the
other \uwds clients as expected.

\subsubsection{Assets loading}

{\tt uwds load}

\subsubsection{Introspection and debugging}

{\tt uwds ls}
{\tt uwds show}

{\tt uwds explorer}

\subsubsection{Interface with ROS}

{\tt uwds tf}

\section{Applications}

\subsection{Spatial Reasoning}

Spatial reasoning~\cite{O'Keefe1999} is a field in its own right, and has been
used for natural language processing for applications such as direction
recognition ~\cite{Kollar2010,Matuszek2010} or language
grounding~\cite{Tellex2010}.~\cite{Skubic2004} present a spatial reasoner
integrated in a robot which computes symbolic positions of objects.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{spatialrelations}
    \caption{The {\tt spatial\_relations} client computes perspective-aware
    spatial relations between objects and agents}
    \label{fig|spatialrelations}
\end{figure}

\cite{Ros2010}

\subsubsection{Visibility computation}

{\tt uwds visibility}

\subsubsection{Perspective Taking}

\section{Conclusion}

\subsection{Future work}

Representation of fields

Client libraries in more languages

\section*{Acknowledgment}

This work has been supported by the EU H2020 Marie Sk\l odowska-Curie Actions
project DoRoThy (grant 657227) and the EU H2020 L2TOR project (grant 688014).

\bibliographystyle{IEEEtran}
\bibliography{bibliography}

\end{document}

### 2017-02-24 paper.tex
Update on Overleaf.

@@
 clouds and meshes). Efficient sharing of such datasets is non-trivial,
-particularly when the said data is dynamically acquired (\eg in the case of
+particularly when the said data are dynamically acquired (\eg in the case of
 map building or object recognition).

### 2017-02-24 paper.tex
WIP intro

@@
 \title{\LARGE \bf
-{\tt underworlds}: Cascading Situation Assessment for Robots
+Cascading Situation Assessment for Robots
 }
@@
 
-Representing \emph{alternative states} is actually often critical, as shown by
-the following exemplary situation: software components manipulating environment
-models typically perform better if the models are physically consistent.
-However, low-level perception inaccuracies often introduce hard-to-avoid
-physical inconsistencies (like objects floating in the air, or on the contrary
-wrongly inset into other objects). Therefore, a post-process stage (for
-instance, using a physics simulation engine) might be desired to correct the
-positions of the objects seen by the robot. Implemented with a classical approach using ROS TF
-frames, we would see an object {\tt book} located at the origin of the reference
-frame {\tt book\_frame}, and a secondary frame {\tt book\_frame\_corrected} computed by the
-physics engine. It is easy to see why such an approach would be quickly
-confusing (where is my book?) and would not scale well, with the robot's 3D
-model cluttered with spurious frames.
+Representing \emph{alternative states} is actually often critical. For instance,
+software components manipulating environment models typically perform better if
+the models are physically consistent. However, low-level perception inaccuracies
+at the sensor level often introduce hard-to-avoid physical inconsistencies (like
+detected objects floating in the air, or on the contrary wrongly inset into
+other objects). Therefore, a post-process stage (for instance, using a physics
+simulation engine) might be desired to move the objects seen
+by the robot into physically-correct positions. Implemented with a classical
+approach (for instance, using ROS TF frames), we would get an object {\tt book}
+located at the frame {\tt book\_frame}, and a secondary frame {\tt
+book\_frame\_corrected} computed by the physics engine. It is easy to see why
+such an approach would be quickly confusing (where is in fact my book?) and would not
+scale well, with the robot's 3D model cluttered with spurious frames.
 
-; efficiently sharing a dynamic 3D scene is
-even more difficult;  bserving a scene from
-different viewpoints to compute visibility is not directly possible
-
-One of the main features of \uwds is the ability to store many parallel
-worlds: past models of the environment, future models, models with some objects
-filtered out, models that are physically consistent, etc.
-
-This package provides the \uwds server, a Python client library, and a set
-of tools to interact with the system (viewers, scene loader, etc.).
-
-A handful of useful example applications are also provided, like skeleton
-tracking (using OpenNI) or visibility tracking.
+Another relevant example pertains to geometric task planning: a geometric task
+planner typically need to reason over hypothetical future states of the
+environment (``what happen if I move this glass onto that pile of books?'').
+The planner might want to quickly generate possible future states, that
+themselves might require further processing (for instance, running a physics
+simulation). This use-case evidences the need for a flexible representation
+system, where models are derived from each others, with partial modifications
+and different timescales.
 
@@
 and object classification~\cite{Lorken2008, Varadarajan2011}.
+
+serving a scene from
+different viewpoints to compute visibility is not directly possible
 

### 2017-02-24 paper.tex
Added photos of the L2TOR setup
WIP design architecture

@@
 
-\subsection{Situation Assessment}
+\subsection{Cascading Situation Assessment}
 
@@
 
-\subsection{A layered model of the environment}
+\subsection{Cascading architecture}
+
+Figure~\ref{fig|scene} pictures a typically \uwds topology: a graph (that
+happens to be a directed acyclic graph on Figure~\ref{fig|scene}, but does not
+have to be in the general case -- cycles are permitted) of \emph{clients}
+connected through shared data structures called \emph{worlds}.
 
@@
     client's needs (yellow nodes represent other possible components in the system that do not
-interact directly with the \uwds network).} \label{fig|scene} \end{figure}
+    interact directly with the \uwds network).}
 
-(worlds are effectively composed of a scene graph representing the 3D geometry
-of the environment, and a timeline storing temporal events)
+    \label{fig|scene}
+
+\end{figure}
+
+\subsubsection{Clients}
+
+Software components accessing \uwds worlds are called clients. Some standard
+clients (like a 3D visualization tool) are provided with the \uwds package (see
+section~\ref{std_clients}), but clients are otherwise written by the end users
+using the \uwds client API (see section~\ref{api}).
+
+Clients can both read and write onto the worlds they are connected to, and
+automatically see updates broadcasted by other clients connected to the same
+worlds.
+
+Four specific types of clients can be distinguished: \textbf{root clients} which
+create and update worlds (`write-only' client, like the \emph{Sensor fusion} and
+\emph{ROS interface} clients on Figure~\ref{fig|scene}); \textbf{leaf clients}
+which on the contrary only read worlds, without modifying them (like the
+\emph{Computation of spatial relations} client on Figure~\ref{fig|scene});
+\textbf{filters} that copy an input world into an output world, performing
+some filtering operation in-between (like the client \textit{Physics-based
+position correction}); and \textbf{transformers} which transform one
+representation into another (like the \emph{Semantic mapping} client on
+Figure~\ref{fig|scene}).
+
+\subsection{Worlds}
+
+Worlds are effectively distributed data structures composed of a scene graph
+representing the 3D geometry of the environment, and a timeline storing temporal
+events.
+
+Each world is technically independent from every others. Dependencies between
+worlds arise from the clients' connections. For instance, filters effectively
+create a dependency between worlds. A client like the \textit{Physics-based
+position correction} client create a dependency between the world {\tt base} (which
+represents here the result of raw sensor fusion) and the world {\tt corrected}
+which would be a physically-consistent copy of {\tt base}.
+
+As a result, a \uwds network can be seen as a dependency graph (where
+cyclic dependencies are permissible), and we call the whole architecture a
+\emph{cascading situation assessment} system.
+
+\subsubsection{Scenes}
+
+\subsubsection{Nodes}
+
+\subsubsection{Timelines}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{timeline}
+    \caption{Both events and episodes are stored in the world's \emph{timeline}}
+
+    \label{fig|timeline}
+\end{figure}
 
@@
 full data whenever they actually need it.
-
-\subsection{Cascading architecture}
-
-Figure~\ref{fig|scene} pictures a typically \uwds topology.
-
-\subsection{Worlds}
-\subsubsection{Scenes}
-
-\subsubsection{Nodes}
-
-\subsubsection{Timelines}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\linewidth]{timeline}
-    \caption{Both events and episodes are stored in the world's \emph{timeline}}
-
-    \label{fig|timeline}
-\end{figure}
 
@@
 \subsection{Standard Clients}
+\label{std_clients}
 
@@
 
-\section{Applications}
-
 \subsection{Spatial Reasoning}
@@
 
+\section{Application Example}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{l2tor-photo2}
+    \caption{}
+    \label{}
+\end{figure}
+\uwds has been used within the European L2TOR project to...
+
+...context + picture
+
+...needs for spatial representation
+
+\subsection*{Architecture}
+
+...diagram with the various \uwds clients
+
 \section{Conclusion}

### 2017-02-24 paper.tex
WIP design architecture

@@
 
-\newcommand{\uwds}{{\tt underworlds}\xspace}
+\newcommand{\uwds}{{\sc underworlds}\xspace}
 
@@
     \includegraphics[width=\linewidth]{overview}
-    \caption{Schema of a \uwds network: eight possible, user-written
+    \caption{Schema of a possible \uwds network: eight possible, user-written
     \emph{clients} (in blue) are exchanging environment models through four shared
@@
 which would be a physically-consistent copy of {\tt base}.
+As a result, a \uwds network can be seen as a dependency graph between worlds (where
+cyclic dependencies are permissible).
 
-As a result, a \uwds network can be seen as a dependency graph (where
-cyclic dependencies are permissible), and we call the whole architecture a
-\emph{cascading situation assessment} system.
+This architecture enables what we call \emph{cascading situation assessment}:
+a set of independent software components (the clients) builds and refines
+successive models of the environment by a combination of
+filtering/transformations steps and model branching.
 
@@
 
-\subsubsection{Nodes}
+\emph{Worlds} contain both a geometric model and a temporal model. The geometric
+model is represented as a scene graph. The scene graph has a unique root node,
+to which a tree of other nodes is parented.
+
+Nodes in the scene graph have four possible types: \textbf{objects} that represent
+concrete physical objects (typically with one or several associated 3D meshes);
+\textbf{entities} that represent abstract entities like reference frames or
+groups of objects; \textbf{perspectives} that represent viewpoints onto the
+scene (like cameras or human gaze); and \textbf{fields} that represent scalar or
+vector fields (like the visibility of an object, the working space of robot,
+etc. -- note that fields are not yet implemented in the current version of
+\uwds, see section~\ref{futurework}).
+
+Every node has a unique ID, a parent, a 3D transformation relative to the parent
+and an optional name. \emph{Object} nodes optionally store as well pointers to their
+associated meshes.
 
@@
     \includegraphics[width=\linewidth]{l2tor-photo2}
-    \caption{}
-    \label{}
+    \caption{A user manipulating objects on RFID mat while an operator monitors the interaction via an {\tt uwds view} client}
+    \label{fig|l2torexample}
 \end{figure}
-\uwds has been used within the European L2TOR project to...
+\uwds has been used within the European L2TOR project~\cite{L2TOR} to implement
 
@@
 \subsection*{Architecture}
-
+The scenario consists of a number of \uwds clients that communicate with each other in order to provide an updated virtual representation to the operator.
+As seen in Figure \ref{fig|l2torexample}, the user interacts with some
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{l2tor-timeline}
+    \caption{Schema of the \uwds architecture used in L2TOR project pilot study. }
+    \label{fig|l2torarchitecture}
+\end{figure}
 ...diagram with the various \uwds clients
@@
 \subsection{Future work}
+\label{futurework}
 

### 2017-02-24 paper.tex
Update on Overleaf.

--- 2017-02-24 paper.tex
@@
 \subsection*{Architecture}
-
+The scenario consists of a number of \uwds clients that communicate with each other in order to provide an updated virtual representation to the operator. As seen in {l2tor}  
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{l2tor-timeline}
+    \caption{Schema of the \uwds architecture used in L2TOR project pilot study. }
+    \label{fig|l2torarchitecture}
+\end{figure}
 ...diagram with the various \uwds clients

### 2017-02-24 paper.tex
Update on Overleaf.
Section 'relation to middlewares'

--- 2017-02-24 paper.tex
@@
 
-\subsection{Relation to existing robotic middlewares}
+\subsection*{Relation to existing robotic middlewares}
+
+Like traditional robotic middlewares, \uwds offers a form of distributed
+computation based on message passing. However, it distinguishes itself from
+existing middleware in several significant ways.
+
+First and foremost, \uwds purposefully does not offer any \emph{general}
+capability at distributing computation and data flows amongst independent
+components: it focuses specifically on distributing environment models, both
+spatial (geometric models) and temporal (events and situations).
+
+While using standard middlewares as \emph{underlying transport} for \uwds would be
+technically feasible and relatively easy to implement, it does not offer any
+clear advantage over lighter and dedicated message passing libraries like ZeroMQ
+or gRPC (the later being the one used by \uwds).
 
@@
 \end{figure}
-\uwds has been used within the European L2TOR project~\cite{L2TOR} to implement
+\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor} in order to conceptualize and visualize the spatial relations and visibility of the physical objects that users interact with.
+In one of the scenarios of this project, users are instructed by a NAO robot to manipulate a number of Duplo-like animals on top of a printed background paper.
+The robot instructions are based on pedagogical strategies to promote spatial relation learning by object manipulation.
+Such instructions include commands like to place the elephant \textit{on top} or \textit{next to} the cube or put the giraffe \textit{inside} the house.
 
-...context + picture
-
+i will finish this section later today
 ...needs for spatial representation
@@
 The scenario consists of a number of \uwds clients that communicate with each other in order to provide an updated virtual representation to the operator.
-As seen in Figure \ref{fig|l2torexample}, the user interacts with some
+As seen in Figure \ref{fig|l2torexample}, the user manipulates the physical objects on top of an RFID mat as instructed by the NAO robot.
 \begin{figure}

### 2017-02-24 paper.tex
WIP design -- timelines

@@
 
-\subsection{Worlds}
+\subsubsection{Worlds}
 
@@
 which would be a physically-consistent copy of {\tt base}.
-As a result, a \uwds network can be seen as a dependency graph between worlds (where
+As a result, an \uwds network can be seen as a dependency graph between worlds (where
 cyclic dependencies are permissible).
@@
 This architecture enables what we call \emph{cascading situation assessment}:
-a set of independent software components (the clients) builds and refines
-successive models of the environment by a combination of
-filtering/transformations steps and model branching.
+independent software components (the clients) build, refine and share successive
+models of the environment by a combination of filtering/transformations steps
+and model branching.
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\linewidth]{timeline}
-    \caption{Both events and episodes are stored in the world's \emph{timeline}}
+Complementing the spatial representation encapsulated in the scene graph, each
+world also store the world's \emph{timeline}. This data structure is shared
+and synchronized amongst the clients in the same way as the scene graph.
+Client can record in the timeline both \emph{events} (duration-less state) and
+\emph{situations}, \ie states with a start time and a (possibly open-ended) end time.
 
-    \label{fig|timeline}
-\end{figure}
+Importantly, the \uwds server automatically generates a snapshot of
+the scene graph whenever an event or situation is added to the timeline. The
+snapshot is associated to the event, which allows clients to effectively retrieve past
+states of the world. This capability is thought to be primarly used by \uwds clients
+performing action recognition.
 
@@
 academic context, as it offers ease and flexibility to plug a variety of
-pre-existing components into a \uwds network.
+pre-existing components into an \uwds network.
 

### 2017-02-24 paper.tex
Update on Overleaf.

--- 2017-02-24 paper.tex
@@
 
-\subsection*{Relation to existing robotic middlewares}
-
-Like traditional robotic middlewares, \uwds offers a form of distributed
-computation based on message passing. However, it distinguishes itself from
-existing middleware in several significant ways.
-
-First and foremost, \uwds purposefully does not offer any \emph{general}
-capability at distributing computation and data flows amongst independent
-components: it focuses specifically on distributing environment models, both
-spatial (geometric models) and temporal (events and situations).
-
-While using standard middlewares as \emph{underlying transport} for \uwds would be
-technically feasible and relatively easy to implement, it does not offer any
-clear advantage over lighter and dedicated message passing libraries like ZeroMQ
-or gRPC (the later being the one used by \uwds).
+\subsection{Relation to existing robotic middlewares}
 
@@
 
-i will finish this section later today
+i will finish it later today
 ...needs for spatial representation

### 2017-02-24 paper.tex
Update on Overleaf.
uwds tf

--- 2017-02-24 paper.tex
@@
 \usepackage{graphicx}
+
+\usepackage{hyperref}
+
 % See the \addtolength command later in the file to balance the column lengths
@@
 scene graph and a timeline. These data structures are synchronized using
-Google's gRPC message passing framework\footnote{http://www.grpc.io/}, ensuring
+Google's gRPC message passing framework\footnote{\url{http://www.grpc.io/}}, ensuring
 high throughput, reliability, cross-platform/cross-language support. The \uwds
@@
 As mentioned in section~\ref{arch}, \uwds uses Google's gRPC as message passing
-protocol. The protocol is explicitely defined, and bindings to various languages
+protocol. The protocol is explicitely defined (using the \emph{protocol
+buffers}\footnote{\url{https://developers.google.com/protocol-buffers/}}
+interface definition language), and bindings to various languages
 and platforms can be automatically generated from the protocol definition file
@@
 
+%%TODO add screenshot of uwds explorer
+
 \subsubsection{Interface with ROS}
 
-{\tt uwds tf}
+\uwds is meant to integrate as easily as possible into existing robot
+architectures, and interfaces transparently with ROS' TF frame system through
+the {\tt uwds tf} client.
+
+The {\tt uwds tf} root client continuously monitor the ROS TF tree, and mirror
+TF frames as nodes in a given \uwds world. A node is first created if none
+matches a given TF frame, and its transformation is subsequently updated,
+mirroring the TF frame. A regular expression can be provided to only mirror a
+subset of the TF tree into \uwds.
+
+Currently, the process is unidirectional: the {\tt uwds tf} client performs TF
+to \uwds updates, but not the opposite.
 
@@
 As seen in Figure \ref{fig|l2torexample}, the user manipulates the physical objects on top of an RFID mat as instructed by the NAO robot.
+
+L2TOR \uwds clients:
+\subsubsection{Kinect Head Tracker}
+\subsubsection{Object RFID tracker}
+\subsubsection{NAO model}
+\subsubsection{Visibility computation}
+\subsubsection{Spatial Relations computation}
+\subsubsection{Game/Rules computation}
+
 \begin{figure}

### 2017-02-24 paper.tex
Added Fotios to authors

@@
 
-\author{Séverin Lemaignan$^{1}$%
-\thanks{$^{1}$Séverin Lemaignan is with Centre for Robotics and Neural Systems,
+\author{Séverin Lemaignan$^{1}$, Fotios Papadopoulos$^{1}$%
+\thanks{$^{1}$Authors are with Centre for Robotics and Neural Systems,
         Plymouth University, Plymouth PL48AA, United Kingdom
-        {\tt\small severin.lemaignan@plymouth.ac.uk}}%
+        {\tt\small firstname.surname@plymouth.ac.uk}}%
 }

### 2017-02-24 paper.tex
Adjusted abstract

@@
 
-    We introduce hereafter \uwds, a novel framework for \emph{cascading}
-    spatio-temporal situation assessment in robotics. \uwds allows programmers
+    We introduce \uwds, a novel lightweight framework for \emph{cascading
+    spatio-temporal situation assessment} in robotics. \uwds allows programmers
     to represent the robot's environment as real-time distributed data
-    structures, containing both scene graphs (for 3D representation of geometries)
-    and timelines (for representation of temporal events). \uwds supports
-    \emph{cascading} representations: the environment is viewed as a set of
-    \emph{worlds} that can each have different spatial and temporal
-    granularities, and may depend on each others.
-
-    \uwds also provides a set of high-level client libraries and tools to introspect
-    and manipulate the environment models.
+    structures, containing both scene graphs (for representation of 3D
+    geometries) and timelines (for representation of temporal events). \uwds
+    supports \emph{cascading} representations: the environment is viewed as a
+    set of \emph{worlds} that can each have different spatial and temporal
+    granularities, and may depend on each others.  \uwds also provides a set of
+    high-level client libraries and tools to introspect and manipulate the
+    environment models.
 

### 2017-02-24 paper.tex
Update on Overleaf.

@@
 \uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor} in order to conceptualize and visualize the spatial relations and visibility of the physical objects that users interact with.
-In one of the scenarios of this project, users are instructed by a NAO robot to manipulate a number of Duplo-like animals on top of a printed background paper.
+In one of the scenarios of this project, users are instructed by a NAO robot to manipulate a number of Duplo-like animals on top of a printed background paper which serves as a landscape (See \ref{fig|l2torexample}).
 The robot instructions are based on pedagogical strategies to promote spatial relation learning by object manipulation.
 Such instructions include commands like to place the elephant \textit{on top} or \textit{next to} the cube or put the giraffe \textit{inside} the house.
-
-i will finish this section later today
-...needs for spatial representation
+For that reason, the robot and the operator should be able to access whether the user performs the required task and at the same time envisage their perspective.
+A number of \uwds clients have been developed in order to provide a distributed model of the environment and the robot as described in the next section.
 
 \subsection*{Architecture}
-The scenario consists of a number of \uwds clients that communicate with each other in order to provide an updated virtual representation to the operator.
-As seen in Figure \ref{fig|l2torexample}, the user manipulates the physical objects on top of an RFID mat as instructed by the NAO robot.
-
-L2TOR \uwds clients:
-\subsubsection{Kinect Head Tracker}
-\subsubsection{Object RFID tracker}
-\subsubsection{NAO model}
-\subsubsection{Visibility computation}
-\subsubsection{Spatial Relations computation}
-\subsubsection{Game/Rules computation}
-
 \begin{figure}
@@
 \end{figure}
-...diagram with the various \uwds clients
+
+The scenario consists of a number of \uwds clients distributed accross two computers that communicate with each other in order to provide an updated representation of the world.
+Figure \ref{fig|l2torarchitecture} illustrates the high level communication of the \uwds clients that form the achitecture of the L2TOR spatial relations platform.\\
+\\L2TOR \uwds clients:
+\subsubsection{Kinect Head Tracker} {Captures the location and the direction of the user's head - Camera point}
+\subsubsection{Object RFID tracker} {Updates the locations of the virtual objects based on RFID data}
+\subsubsection{NAO model}{Virtual representation of the Robot - Camera poit}
+\subsubsection{Visibility computation}{Computes visibility based on camera points}
+\subsubsection{Spatial Relations computation}{Extracts spatial relations between the virtual objects}
+\subsubsection{Game/Rules computation}{Interaction manager based on \uwds clients' input}
+
+The L2TOR project features an additional scenario where the user performs the manipulations on a tablet computer instead of using physical objects.
+To accomodate this change, we replaced the object detection \uwds client with a 3D virtual object detection while the rest of the system remains the same.
+Finally, we are evaluating the use of a hybrid platform that utilizes both virtual and physical objects at the same time.
 

### 2017-02-27 paper.tex
WIP on intro, in particular, related work

@@
 to build and refine in real-time spatial and temporal models of the environment
-surrounding a robot. \uwds makes it easy, then, to share these models amongst
-the other software components running on the robot, to eventually provide a
-distributed environment model for the robot.
+surrounding a robot. \uwds make it possible to share these world models amongst
+the software components running on the robot. Critically, \uwds permits to
+represent and manipulate in a distributed manner \emph{multiple alternatives} to
+the current, perceived world model: for instance, the world with some objects filtered out,
+the world `viewed' from the perspective of another agent, an hypothetical world
+resulting from the simulated application of a plan, etc.
 
-The components which require such spatial and temporal models of the environment
-are typically found in the intermediate layers of robotic architectures, between
+\subsection{Situation Assessment}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=\linewidth]{overview}
+    \caption{Schema of a possible \uwds network: eight possible, user-written
+    \emph{clients} (in blue) are exchanging environment models through four shared
+    \emph{worlds} (made of joint spatial and temporal models). This architecture
+    enables successive and modular refinement of the models
+    (\emph{cascading} situation assessment), effectively adapted to each
+    client's needs (yellow nodes represent other possible components in the system that do not
+    interact directly with the \uwds network).}
+
+    \label{fig|scene}
+
+\end{figure}
+
+Anchoring perceptions in a symbolic model suitable for decision-making requires
+perception abilities and their symbolic interpretation. We call \emph{physical
+situation assessment} the cognitive skill that a robot exhibits when it
+represents and assesses the nature and content of its surroundings and monitors
+its evolution.
+
+Numerous approaches exist, like amodal (in the sense of modality-independent)
+\emph{proxies}~\cite{Jacobsson2008}, grounded amodal
+representations~\cite{Mavridis2006}, semantic
+maps~\cite{Nuechter2008, Galindo2008,Blodow2011} or affordance-based planning
+and object classification~\cite{Lorken2008, Varadarajan2011}.
+
+\uwds is specifically inspired by the geometric and temporal reasoning module
+{\sc Spark} (\emph{SPAtial Reasoning \& Knowledge}~\cite{sisbot2011situation}.
+{\sc Spark} acts as a situation assessment reasoner that generates symbolic
+knowledge from the geometry of the environment with respect to relations between
+objects, robots and humans. It also takes into account the different perspective
+that each agent has on the environment. {\sc Spark} embeds a
+modality-independent geometric model of the environment that serves both as
+basis for the fusion of the perception modalities and as bridge with the
+symbolic layer~\cite{lemaignan2016artificial}. This geometric model is built
+from 3D CAD models of the objects, furnitures and robots, and full body, rigged
+models of humans.  It is updated at run-time by the robot's sensors.
+Likewise, \uwds embeds a grounded amodal model of the environment, updated
+online from the robot's sensors (sensor fusion).
+
+However, {\sc Spark} is a monolithic module, that does not aim at sharing its
+internal 3D model to other external components. On the contrary, \uwds focuses
+on offering a shared and distributed representation within the robot's software
+architecture. Previous work on distributed scene graphs~\cite{naef2003blue} has
+been applied to robotics to provide a shared 3D representation of the robot's
+environment (like the \emph{Robot Scene Graph}~\cite{blumenthal2013scene} or 
+\emph{Deep State Representation}~\cite{bustos2016unified}).
+
+\uwds further extends this line of work by providing the ability to manipulate
+several \emph{alternative 3D worlds}, corresponding to filtered or hypothetical
+\emph{views} on the perceived model of the environment.
+
+\subsection{Representing Alternative States of the Environment}
+
+The components which make use of spatial and temporal models of the environment
+are usually found in the intermediate layers of robotic architectures, between
 the low-level perceptual layers, and the high-level decisional layers. They
-typically include geometric reasoners (that compute spatial and topological
-relations between objects), motion planners, gestures/sequence recognizers.
+include modules like geometric reasoners (that compute spatial and topological
+relations between objects), motion planners, action recognition.
 
-Difficulties however arise quickly. These different components might need
-different models, at different spatial and temporal resolutions. A 3D motion
-planner would typically use coarse 3D models of surrounding objects to lower the
-computational load while planning, while a spatial reasoner might need on the
-contrary a high-resolution model of the environment for accurate rendering when
-attempting to estimate the visibility of an object. Spatial models are also
-usually large as they might contain a lot of geometric data (including point
-clouds and meshes). Efficient sharing of such datasets is non-trivial,
-particularly when the said data are dynamically acquired (\eg in the case of
-map building or object recognition).
+These different components typically need different models, at different spatial
+and temporal resolutions. A 3D motion planner would typically use coarse 3D
+models of surrounding objects to lower the computational load while planning,
+while a spatial reasoner might need on the contrary a high-resolution model of
+the environment for accurate rendering when attempting to estimate the
+visibility of an object. Spatial models are also usually large as they might
+contain a lot of geometric data (including point clouds and meshes). Efficient
+sharing of such datasets is non-trivial, particularly when the said data are
+dynamically acquired (\eg in the case of map building or object recognition).
 
@@
 position are non-trivial with ROS, for instance) and representing and reasoning
-on \emph{alternative states} of the environment is not directly doable.
+on \emph{alternative states} of the environment is not directly feasible.
 
@@
 
-\subsection{Cascading Situation Assessment}
-
-Anchoring perceptions in a symbolic model requires perception abilities and
-their symbolic interpretation. We call \emph{physical situation assessment} the
-cognitive skill that a robot exhibits when it assesses the nature and content of its
-surroundings and monitors its evolution.
-
-Numerous approaches exist, like amodal (in the sense of modality-independent)
-\emph{proxies}~\cite{Jacobsson2008}, grounded amodal
-representations~\cite{Mavridis2006}, semantic
-maps~\cite{Nuechter2008, Galindo2008,Blodow2011} or affordance-based planning
-and object classification~\cite{Lorken2008, Varadarajan2011}.
-
 serving a scene from
@@
 
-\section{Related work}
+Relation to Beetz's \emph{deep representation}~\cite{beetz2010towards}
 
-\cite{sisbot2011situation}
-\cite{naef2003blue}
-\cite{bustos2016unified}
+GSM (for \emph{Grounded Situation Model})~\cite{Mavridis2006} is a knowledge
+representation system primarily built to ``facilitate cross-modal
+interoperability'',  especially in the context of verbal interaction with a
+robot.
 
-\uwds is inspired by geometric and temporal reasoning components like {\sc
-Spark} (\emph{SPAtial Reasoning \&
-Knowledge}~\cite{sisbot2011situation} (see~\cite{lemaignan2016artificial} for a
-high-level description of the integration of a component like {\sc Spark} in a
-complete robotic architecture).
+GSM does not rely on any formal language but rather on a layered data structure
+(figure~\ref{fig|gsm}) that organises the surrounding world into agents and
+relations between agents.  Each agent (any animate or inanimate object) is
+attached to a physical model (made of \emph{body parts} that have properties
+like their position, color, etc.) and a mental model (which is a recursively
+embedded GSM, thus allowing a sort of theory of mind).
 
-It acts as a situation assessment reasoner that generates symbolic knowledge from the
-geometry of the environment with respect to relations between objects, robots
-and humans, also
-taking into account the different perspective that each agent has on the
-environment.  {\sc Spark} embeds an \emph{amodal} (as defined by Mavridis and
-Roy in~\cite{Mavridis2006}: the different perceptual modalities are abstracted
-away into a blended spatial model) geometric model of the environment that
-serves both as basis for the fusion of the perception modalities and as bridge
-with the symbolic layer. This geometric model is built from 3D CAD models of the
-objects, furnitures and robots, and full body, rigged models of humans
-(Figure~\ref{fig:sparkScreenshot}).  It is updated at run-time by the robot's
-sensors (usually, a combination of vision-based tracking of 2D fiducial markers
-to identify and localise objects, and Kinect-based skeleton tracking of humans,
-optionally assisted by motion capture to accurately track the head motion, which
-is required to compute what the human is looking at).
+Properties are represented in three layers: a stochastic representation, close
+to sensory percepts, a \emph{continuous single-valued} encoding of the
+stochastic model, and a discrete, categorical model.
 
-\subsection*{Relation to existing robotic middlewares}
+One notable feature of GSM is the \emph{bidirectionality} of the grounding
+process: not only sensor percepts are abstracted into categories suitable for
+human conversation, but human utterance (like ``There is a red ball in the
+center of the table'') can also be turned into property descriptions. This
+basically enable the knowledge representation system of the robot to
+\emph{imagine} entities.
+
+GSM also features several strategies for managing time and events.
+\emph{Moments} are created by storing timestamped snap-shots of GSM, and
+\emph{event classifiers} allow to define and detect events.
+
+\subsection{Relation to existing robotic middlewares}
 
@@
 connected through shared data structures called \emph{worlds}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{overview}
-    \caption{Schema of a possible \uwds network: eight possible, user-written
-    \emph{clients} (in blue) are exchanging environment models through four shared
-    \emph{worlds} (made of joint spatial and temporal models). This architecture
-    enables successive and modular refinement of the models
-    (\emph{cascading} situation assessment), effectively adapted to each
-    client's needs (yellow nodes represent other possible components in the system that do not
-    interact directly with the \uwds network).}
-
-    \label{fig|scene}
-
-\end{figure}
 

### 2017-02-27 paper.tex
Caption of python code snippet

@@
 \end{pythoncode}
-    \caption{Example of a simple \uwds Python client named {\tt walls\_filter}:
-    the client connects to the \uwds network, efficiently accesses the world
-    {\tt example\_in}, filter out some objects, and publish the remaining
-    objects in the world {\tt example\_out}.} \label{lst|pythonapi}
+    \caption{Example of a simple \uwds filter, written in Python, and named {\tt
+    walls\_filter}: the client connects to the \uwds network, walkthrough the
+    scene graph of world {\tt example\_in}, filter out some objects, and publish
+    the remaining objects in the world {\tt example\_out}.}
+
+    \label{lst|pythonapi}
 \end{listing}

### 2017-02-27 paper.tex
More work on intro

@@
     \caption{Schema of a possible \uwds network: eight possible, user-written
-    \emph{clients} (in blue) are exchanging environment models through four shared
-    \emph{worlds} (made of joint spatial and temporal models). This architecture
+    \emph{clients} (in blue) are sharing environment models through four
+    independent \emph{worlds} (made of joint spatial and temporal models). This architecture
     enables successive and modular refinement of the models
@@
 
-However, {\sc Spark} is a monolithic module, that does not aim at sharing its
+However, {\sc Spark} is a monolithic module, that does not support sharing its
 internal 3D model to other external components. On the contrary, \uwds focuses
@@
 environment (like the \emph{Robot Scene Graph}~\cite{blumenthal2013scene} or 
-\emph{Deep State Representation}~\cite{bustos2016unified}).
+\emph{Deep State Representation}~\cite{bustos2016unified} approaches).
 
@@
 
-These different components typically need different models, at different spatial
-and temporal resolutions. A 3D motion planner would typically use coarse 3D
-models of surrounding objects to lower the computational load while planning,
-while a spatial reasoner might need on the contrary a high-resolution model of
-the environment for accurate rendering when attempting to estimate the
-visibility of an object. Spatial models are also usually large as they might
-contain a lot of geometric data (including point clouds and meshes). Efficient
-sharing of such datasets is non-trivial, particularly when the said data are
-dynamically acquired (\eg in the case of map building or object recognition).
+These different components typically exhibit different needs in terms of model,
+like different nominal spatial and/or temporal resolutions. For instance, a 3D
+motion planner would typically use coarse 3D models of surrounding objects to
+lower the computational load while planning, while a module assessing the
+visibility of objects might need on the contrary a high-resolution model of the
+environment for accurate 3D rendering. This requirement of multiple
+task-specific representations has been framed as the need for \emph{deep
+representation} by Beetz~\cite{beetz2010towards}.
+
+%Spatial models are also usually large as they might
+%contain a lot of geometric data (including point clouds and meshes). Efficient
+%real-time sharing of such datasets is non-trivial, particularly when the said
+%data is changing (\eg in the case of map building or object recognition).
 
 Traditional robotic middlewares like ROS are not particularly well-suited to
-deal with these complexities: geometric data can be represented, but are not
-first-class citizens (basic tasks like displaying a 3D mesh at a dynamic
-position are non-trivial with ROS, for instance) and representing and reasoning
-on \emph{alternative states} of the environment is not directly feasible.
+deal with these different needs: full geometric data can be represented, but are not
+first-class citizens (for instance, basic tasks like displaying a 3D mesh at a dynamic
+position are not particularly easy with ROS) and critically, representing and reasoning
+on alternative models of the environment is not directly feasible.
 
-Representing \emph{alternative states} is actually often critical. For instance,
+Representing \emph{alternative states} is however often highly desirable. For instance,
 software components manipulating environment models typically perform better if
 the models are physically consistent. However, low-level perception inaccuracies
-at the sensor level often introduce hard-to-avoid physical inconsistencies (like
+often introduce hard-to-avoid physical inconsistencies (like
 detected objects floating in the air, or on the contrary wrongly inset into
 other objects). Therefore, a post-process stage (for instance, using a physics
-simulation engine) might be desired to move the objects seen
+simulation engine) might be needed to displace the objects seen
 by the robot into physically-correct positions. Implemented with a classical
@@
 
-serving a scene from
-different viewpoints to compute visibility is not directly possible
+A third example relates to human-robot interaction scenarios where \emph{perspective
+taking} is important (a prototypical example being the game `I spy with my
+little eyes', as implemented in~\cite{ros2010which}). Perspective taking is a
+cognitive skill that relies on the ability for an agent to project itself in
+others' `eyes' to estimate what they see from their perspective. While
+perspective taking has been previously implemented in robotics by temporary placing
+virtual cameras at eyes position for each of the humans tracked by the
+robot~\cite{ros2010solving}. a the face of  swapping the different viewpoints
+to compute visibility is not directly possible
 
-Relation to Beetz's \emph{deep representation}~\cite{beetz2010towards}
-
-GSM (for \emph{Grounded Situation Model})~\cite{Mavridis2006} is a knowledge
-representation system primarily built to ``facilitate cross-modal
-interoperability'',  especially in the context of verbal interaction with a
-robot.
-
-GSM does not rely on any formal language but rather on a layered data structure
-(figure~\ref{fig|gsm}) that organises the surrounding world into agents and
-relations between agents.  Each agent (any animate or inanimate object) is
-attached to a physical model (made of \emph{body parts} that have properties
-like their position, color, etc.) and a mental model (which is a recursively
-embedded GSM, thus allowing a sort of theory of mind).
-
-Properties are represented in three layers: a stochastic representation, close
-to sensory percepts, a \emph{continuous single-valued} encoding of the
-stochastic model, and a discrete, categorical model.
-
-One notable feature of GSM is the \emph{bidirectionality} of the grounding
-process: not only sensor percepts are abstracted into categories suitable for
-human conversation, but human utterance (like ``There is a red ball in the
-center of the table'') can also be turned into property descriptions. This
-basically enable the knowledge representation system of the robot to
-\emph{imagine} entities.
-
-GSM also features several strategies for managing time and events.
-\emph{Moments} are created by storing timestamped snap-shots of GSM, and
-\emph{event classifiers} allow to define and detect events.
+Geometric \emph{pre-supposition accommodation} makes another interesting case
+for alternative worlds representation. \emph{Pre-supposition accommodation}
+comes originally from linguistics, where it describes the mechanism by which
+\emph{context is adjusted [...] to accept [...] a sentence that imposes certain
+requirements on the context in which it is
+processed}~\cite{vonfintel2008presupposition}. In the context of spatio-temporal
+representations, we call pre-supposition accommodation the ability for an agent
+to adjust its model so that it matches some external contextual constraint. For
+instance, if A tells B to ``catch the red balloon behind you'', B adds a
+representation of a red balloon behind him, even without actually seeing the
+balloon: B accommodates the pre-supposition of a red balloon being present
+behind himself. Endowing robots with this capability has been touched by
+Mavridis \etal within their multi-modal \emph{Grounded Situation
+Model}~\cite{Mavridis2006}. However, to the best of our knowledge, a general
+framework which would enable robots to derive imaginary worlds by extending of
+modifying existing ones has not been proposed so far.
 

### 2017-02-27 paper.tex
Work on clients

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{uwds-screenshot}
+    \caption{Screenshot of the {\tt uwds view} 3D visualization and manipulation
+    tool. In this particular example, the 3D meshes have been pre-loaded using
+    {\tt uwds load}. Their positions are then updated at run time using the
+    robot's sensors and proprioception (joint state).}
+    \label{fig|uwds-view}
+\end{figure}
+
 \subsection{Distributed spatio-temporal models}
@@
 \end{pythoncode}
-    \caption{Example of a simple \uwds filter, written in Python, and named {\tt
+    \caption{Example of a simple yet complete \uwds filter, written in Python, and named {\tt
     walls\_filter}: the client connects to the \uwds network, walkthrough the
@@
 
-\subsubsection{3D Visualization}
+\subsubsection{3D Visualization and manipulation}
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\linewidth]{uwds-screenshot}
-    \caption{Screenshot of the {\tt uwds view} 3D visualization and manipulation
-    tool.}
-    \label{fig|uwds-view}
-\end{figure}
-
 This tool also support basic object manipulations, that are broadcasted to the
@@
 
-\subsubsection{Assets loading}
+\paragraph{Assets loading}
 
@@
 
-{\tt uwds ls}
-{\tt uwds show}
+\uwds provides a range of tools to inspect a running the network. Graphical
+tools ({\tt uwds explorer} and {\tt uwds timeline}, see Figure~\ref{fig|explorer})
+provide an user-friendly overview of the system's graph with the connections
+between the clients and the worlds as well as their activity.
 
-{\tt uwds explorer}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{tools}
+    \caption{Screenshots of the timeline (top) and network (bottom)
+    introspection tools. Worlds are represented as boxes, clients as ellipses.}
+    \label{fig|explorer}
+\end{figure}
 
-%%TODO add screenshot of uwds explorer
+Specialized command-line tools are also available to list at run-time the worlds
+and their content ({\tt uwds ls}) or to display detailed informations on a
+specific node ({\tt uwds show}).
 
@@
 
+\begin{listing}[h!]
+\begin{pythoncode}
+import copy
+import underworlds
+from underworlds.tools.visibility import VisibilityMonitor
+
+with underworlds.Context("Human PoV") as ctx:
+
+    source = ctx.worlds["base"]
+    target = ctx.worlds["human_perspective"]
+
+    # pick up the first node named 'human'
+    human = source.scene.nodebyname("human")[0]
+    target.scene.nodes.append(human)
+
+    # VisibilityMonitor computes the set of visible 
+    # nodes from a given viewpoint
+    visibility = VisibilityMonitor(ctx, source)
+
+    node_mapping = {}
+
+    while True:
+
+        for node in visibility.from_camera(human):
+
+            newnode = node.copy()
+
+            # track the correspondence between nodes
+            # in source and target scenes
+            if node in node_mapping:
+                newnode.id = node_mapping[node].id
+            else:
+                node_mapping[node] = newnode
+
+            # reparent the nodes to the new root
+            if node.parent == source.scene.rootnode.id:
+                newnode.parent = target.scene.rootnode.id
+
+            target.scene.nodes.update(newnode)
+
+        source.scene.waitforchanges()
+
+\end{pythoncode}
+    \caption{This \uwds client creates a world named {\tt human\_perspective}
+    from a source world {\tt base} by only updating objects that are in the
+    human field of view. As such, {\tt human\_perspective} is a real-time model
+    of the human perspective on the scene.}
+
+    \label{lst|humanperspective} \end{listing}
+
 \section{Application Example}
@@
     \centering
-    \includegraphics[width=\linewidth]{l2tor-timeline}
+    \includegraphics[width=\linewidth]{l2tor}
     \caption{Schema of the \uwds architecture used in L2TOR project pilot study. }
@@
 
+Representation of uncertainty
+
 Representation of fields
@@
 Client libraries in more languages
+
+Interplay with effort like RSG-DSL~\cite{blumenthal2014towards}
 

### 2017-02-27 paper.tex
More on clients + various little bits

@@
 
-\subsection{Relation to existing robotic middlewares}
-
-Like traditional robotic middlewares, \uwds offers a form of distributed
-computation based on message passing. However, it distinguishes itself from
-existing middleware in several significant ways.
-
-First and foremost, \uwds purposefully does not offer any \emph{general}
-capability at distributing computation and data flows amongst independent
-components: it focuses specifically on distributing environment models, both
-spatial (geometric models) and temporal (events and situations).
-
-While using standard middlewares as \emph{underlying transport} for \uwds would be
-technically feasible and relatively easy to implement, it does not offer any
-clear advantage over lighter and dedicated message passing libraries like ZeroMQ
-or gRPC (the later being the one used by \uwds).
+The main contribution of this work is a generic approach to represent and share
+\textbf{multiple parallel representations of the world}. \uwds does so by
+permitting clients to clone an existing world, modify it, and re-share it,
+without the cost of duplicating geometric data (as explained in
+section~\ref{design}). By organizing clients in a network
+(Figure~\ref{fig|scene}), worlds can be made dependent on each other, resulting
+in a loosely-coupled modular approach to spatio-temporal world representation
+that we call \emph{cascading situation assessment}.
 
 \section{Design and Architecture}
+\label{design}
 
@@
 and an optional name. \emph{Object} nodes optionally store as well pointers to their
-associated meshes.
+associated meshes. Importantly, mesh data (or other geometric datasets like
+point clouds) are \emph{not} stored within the nodes. \uwds represents geometric
+data as immutable data, labelled with their hash (preventing data duplication).
+Nodes only store the hash of their geometric data, and the actual data is pulled
+by client whenever they actually need it (for rendering for instance).
 
@@
 
-\paragraph{Assets loading}
+\paragraph*{Assets loading}
 
-{\tt uwds load}
+Geometric data like meshes can be loaded into a world (or rather, into the whole
+network as geometric data are treated as immutable resources shared amongst all
+worlds) by simply sending an array of vertices and faces to the \uwds server.
+As such, any client can load 3D data that is automatically made available to the
+whole network. For example, an object recognition module recognizes a new
+object. It sends the mesh to the \uwds server and creates a new node which
+references the mesh. At the next frame update, the 3D viewer sees a new node
+referencing a yet unknown mesh. The viewer request the mesh data from the
+server, stores it internally (\eg as an OpenGL buffer), and renders it. Whenever
+the node transformation is updated, the 3D viewer can re-render it without
+having to download the mesh from the server again.
+
+Often, objects manipulated by the robot have known meshes with corresponding CAD
+models. In these case, \uwds provides a tool, {\tt uwds load}, that loads a mesh
+into a \uwds network (and optionally, creates a node) from a large range of 3D
+format\footnote{The underlying import capability is provided by the {\sc assimp}
+library.  \url{http://assimp.sourceforge.net/}} (including Collada, FBX, OBJ,
+Blender).
 
@@
 
-\subsection{Spatial Reasoning}
+\subsection{Spatial Reasoning and Perspective Taking}
 
@@
 used for natural language processing for applications such as direction
-recognition ~\cite{Kollar2010,Matuszek2010} or language
-grounding~\cite{Tellex2010}.~\cite{Skubic2004} present a spatial reasoner
-integrated in a robot which computes symbolic positions of objects.
+recognition~\cite{Kollar2010,Matuszek2010} or language
+grounding~\cite{Tellex2010}. Other examples in human-robot interaction include Ros
+\etal~\cite{ros2010solving, ros2010which} which has been recently integrated
+into a full architecture for autonomous human-robot
+interaction~\cite{lemaignan2016artificial}.
 
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{spatialrelations}
+    \includegraphics[width=\linewidth]{spatialrelations}
     \caption{The {\tt spatial\_relations} client computes perspective-aware
-    spatial relations between objects and agents}
+    spatial relations between objects and agents: \emph{allo-centric} relations (like
+    {\tt is in} or {\tt is on}) are independent of the viewpoint, while
+    \emph{ego-centric} relations ({\tt in front of}, {\tt left of}) depend on
+    the viewer perspective.}
     \label{fig|spatialrelations}
@@
 
-\cite{Ros2010}
+\uwds provides a client ({\tt spatial\_relations}) to compute both
+allo-centric and ego-centric (\ie viewer-dependent) spatial relations
+between objects (Figure~\ref{fig|spatialrelations}).
 
-\subsubsection{Visibility computation}
+\uwds also implement an efficient algorithm to assess object visibility from a
+specific viewpoint (\ie from a given \emph{camera} node). The algorithm (color
+picking) enables fast (single pass) computation of the visibility of every
+object in the scene, while providing control regarding how much of an object
+should be actually visible to be considered as globally visible. The
+command-line tool {\tt uwds visibility} returns the list of visible objects from
+the point of view of every camera in a given world, and \uwds also provides the
+helper class {\tt VisibilityMonitor} to programmatically access visibility
+information.
 
-{\tt uwds visibility}
-
-\subsubsection{Perspective Taking}
+When integrated into a filter node, visibility computation enables to easily
+create new worlds representing the estimated perspectives of the different
+agents. Listing~\ref{lst|humanperspective} provides the complete code of such an
+\uwds client.
 
@@
 \end{pythoncode}
-    \caption{This \uwds client creates a world named {\tt human\_perspective}
-    from a source world {\tt base} by only updating objects that are in the
-    human field of view. As such, {\tt human\_perspective} is a real-time model
-    of the human perspective on the scene.}
+    \caption{Example of an \uwds client that creates a world named {\tt
+    human\_perspective} from a source world {\tt base} by only updating objects
+    that are in the field of view of the human. As such, {\tt human\_perspective} is
+    a real-time model of the human perspective on the scene.}
 
@@
 
-\section{Conclusion}
+\section{Discussion and Conclusion}
+
+\subsection{Relation to existing robotic middlewares}
+
+Like traditional robotic middlewares, \uwds offers a form of distributed
+computation based on message passing. However, it distinguishes itself from
+existing middleware in several significant ways.
+
+First and foremost, \uwds purposefully does not offer any \emph{general}
+capability at distributing computation and data flows amongst independent
+components: it focuses specifically on distributing environment models, both
+spatial (geometric models) and temporal (events and situations).
+
+While using standard middlewares as \emph{underlying transport} for \uwds would be
+technically feasible and relatively easy to implement, it does not offer any
+clear advantage over lighter and dedicated message passing libraries like ZeroMQ
+or gRPC (the later being the one used by \uwds).
 
@@
 
-Interplay with effort like RSG-DSL~\cite{blumenthal2014towards}
+\subsection{Conclusion}
+
+Integration with effort like RSG-DSL~\cite{blumenthal2014towards}
 
@@
 
+The authors also acknowledge the contribution of the HRI lab of LAAS-CNRS (where this
+work debuted), and in particular, the influential ideas of Rachid Alami.
+
 \bibliographystyle{IEEEtran}

### 2017-02-27 paper.tex
WIP Conclusion + formatting

@@
 
-\uwds is a distributed and lightweight framework that enables robot programmers
-to build and refine in real-time spatial and temporal models of the environment
-surrounding a robot. \uwds make it possible to share these world models amongst
-the software components running on the robot. Critically, \uwds permits to
-represent and manipulate in a distributed manner \emph{multiple alternatives} to
-the current, perceived world model: for instance, the world with some objects filtered out,
-the world `viewed' from the perspective of another agent, an hypothetical world
-resulting from the simulated application of a plan, etc.
+\uwds is a distributed and lightweight open-source
+framework\footnote{\url{https://github.com/severin-lemaignan/underworlds}} that
+enables robot programmers to build and refine in real-time spatial and temporal
+models of the environment surrounding a robot. \uwds make it possible to share
+these world models amongst the software components running on the robot.
+Critically, \uwds permits to represent and manipulate in a distributed manner
+\emph{multiple alternatives} to the current, perceived world model: for
+instance, the world with some objects filtered out, the world `viewed' from the
+perspective of another agent, an hypothetical world resulting from the simulated
+application of a plan, etc.
 
@@
 As mentioned in section~\ref{arch}, \uwds uses Google's gRPC as message passing
-protocol. The protocol is explicitely defined (using the \emph{protocol
+protocol. The protocol is explicitly defined (using the \emph{protocol
 buffers}\footnote{\url{https://developers.google.com/protocol-buffers/}}
@@
     \caption{Example of a simple yet complete \uwds filter, written in Python, and named {\tt
-    walls\_filter}: the client connects to the \uwds network, walkthrough the
+    walls\_filter}: the client connects to the \uwds network, walk through the
     scene graph of world {\tt example\_in}, filter out some objects, and publish
@@
 
-    \label{lst|humanperspective} \end{listing}
+    \label{lst|humanperspective}
+\end{listing}
 
@@
     \includegraphics[width=\linewidth]{l2tor-photo2}
-    \caption{A user manipulating objects on RFID mat while an operator monitors the interaction via an {\tt uwds view} client}
+    \caption{A user manipulating objects on RFID mat while an operator monitors
+    the interaction via an {\tt uwds view} client}
+
     \label{fig|l2torexample}
 \end{figure}
-\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor} in order to conceptualize and visualize the spatial relations and visibility of the physical objects that users interact with.
-In one of the scenarios of this project, users are instructed by a NAO robot to manipulate a number of Duplo-like animals on top of a printed background paper which serves as a landscape (See \ref{fig|l2torexample}).
-The robot instructions are based on pedagogical strategies to promote spatial relation learning by object manipulation.
-Such instructions include commands like to place the elephant \textit{on top} or \textit{next to} the cube or put the giraffe \textit{inside} the house.
-For that reason, the robot and the operator should be able to access whether the user performs the required task and at the same time envisage their perspective.
-A number of \uwds clients have been developed in order to provide a distributed model of the environment and the robot as described in the next section.
+
+\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor}
+in order to conceptualize and visualize the spatial relations and visibility of
+the physical objects that users interact with. In one of the scenarios of this
+project, users are instructed by a NAO robot to manipulate a number of
+Duplo-like animals on top of a printed background paper which serves as a
+landscape (See \ref{fig|l2torexample}). The robot instructions are based on
+pedagogical strategies to promote spatial relation learning by object
+manipulation. Such instructions include commands like to place the elephant
+\textit{on top} or \textit{next to} the cube or put the giraffe \textit{inside}
+the house.  For that reason, the robot and the operator should be able to access
+whether the user performs the required task and at the same time envisage their
+perspective. A number of \uwds clients have been developed in order to provide a
+distributed model of the environment and the robot as described in the next
+section.
 
@@
 
-The scenario consists of a number of \uwds clients distributed accross two computers that communicate with each other in order to provide an updated representation of the world.
-Figure \ref{fig|l2torarchitecture} illustrates the high level communication of the \uwds clients that form the achitecture of the L2TOR spatial relations platform.\\
-\\L2TOR \uwds clients:
-\subsubsection{Kinect Head Tracker} {Captures the location and the direction of the user's head - Camera point}
-\subsubsection{Object RFID tracker} {Updates the locations of the virtual objects based on RFID data}
-\subsubsection{NAO model}{Virtual representation of the Robot - Camera poit}
-\subsubsection{Visibility computation}{Computes visibility based on camera points}
-\subsubsection{Spatial Relations computation}{Extracts spatial relations between the virtual objects}
-\subsubsection{Game/Rules computation}{Interaction manager based on \uwds clients' input}
+The scenario consists of a number of \uwds clients distributed across two
+computers that communicate with each other in order to provide an updated
+representation of the world. Figure \ref{fig|l2torarchitecture} illustrates the
+high level communication of the \uwds clients that form the architecture of the
+L2TOR spatial relations platform.
 
-The L2TOR project features an additional scenario where the user performs the manipulations on a tablet computer instead of using physical objects.
-To accomodate this change, we replaced the object detection \uwds client with a 3D virtual object detection while the rest of the system remains the same.
-Finally, we are evaluating the use of a hybrid platform that utilizes both virtual and physical objects at the same time.
+L2TOR \uwds clients:
+
+\begin{itemize}
+    \item \textbf{Kinect Head Tracker} Captures the location and the direction
+        of the user's head - Camera point
+
+    \item \textbf{Object RFID tracker} Updates the locations of the virtual
+        objects based on RFID data
+
+    \item \textbf{NAO model} Virtual representation of the Robot - Camera poit
+
+    \item \textbf{Visibility computation} Computes visibility based on camera
+        points
+
+    \item \textbf{Spatial Relations computation} Extracts spatial relations
+        between the virtual objects
+
+    \item \textbf{Game/Rules computation} Interaction manager based on \uwds
+        clients' input
+
+\end{itemize}
+
+The L2TOR project features an additional scenario where the user performs the
+manipulations on a tablet computer instead of using physical objects. To
+accommodate this change, we replaced the object detection \uwds client with a 3D
+virtual object detection while the rest of the system remains the same. Finally,
+we are evaluating the use of a hybrid platform that utilizes both virtual and
+physical objects at the same time.
 
@@
 
+We have introduced \uwds, a novel framework for shared and composable
+spatio-temporal representations of a robot's world. The key contributions of our
+approach are: a composite data-structure for environment representation within a
+robotic software architecture, made of a scene graph and a timeline; a mechanism
+to efficiently and transparently share this data-structure amongst a set of
+clients (the software modules of the robot); a cascading architecture permitting
+the explicit of representation of alternative states of the world while
+maintaining a network of dependencies amongst themselves; a typology of clients
+(\emph{root}, \emph{leaf}, \emph{filter}, \emph{transformer}) to systematically
+describe such a network of representations dependencies.
+
+We have presented as well a concrete instantiation of a system relying on \uwds
+for its representation needs, and we have sketched future directions of
+development.
+
 Integration with effort like RSG-DSL~\cite{blumenthal2014towards}

### 2017-02-27 paper.tex
Proof-reading part 1

@@
 
-\author{Séverin Lemaignan$^{1}$, Fotios Papadopoulos$^{1}$%
-\thanks{$^{1}$Authors are with Centre for Robotics and Neural Systems,
+\author{Séverin Lemaignan$^{*}$, Fotios Papadopoulos$^{*}$%
+\thanks{$^{*}$Authors are with Centre for Robotics and Neural Systems,
         Plymouth University, Plymouth PL48AA, United Kingdom
@@
 enables robot programmers to build and refine in real-time spatial and temporal
-models of the environment surrounding a robot. \uwds make it possible to share
+models of the environment surrounding a robot. \uwds makes it possible to share
 these world models amongst the software components running on the robot.
-Critically, \uwds permits to represent and manipulate in a distributed manner
-\emph{multiple alternatives} to the current, perceived world model: for
-instance, the world with some objects filtered out, the world `viewed' from the
-perspective of another agent, an hypothetical world resulting from the simulated
+Besides, \uwds permits to represent and manipulate in a distributed manner
+\emph{multiple alternatives} to the current, perceived world model. For
+instance, the world with some objects filtered out; the world `viewed' from the
+perspective of another agent; an hypothetical world resulting from the simulated
 application of a plan, etc.
 
-\subsection{Situation Assessment}
+\subsection{Distributed Situation Assessment}
 
@@
 
-\uwds is specifically inspired by the geometric and temporal reasoning module
-{\sc Spark} (\emph{SPAtial Reasoning \& Knowledge}~\cite{sisbot2011situation}.
+\uwds is specifically inspired by the geometric and temporal reasoner
+{\sc Spark} (\emph{SPAtial Reasoning \& Knowledge})~\cite{sisbot2011situation}.
 {\sc Spark} acts as a situation assessment reasoner that generates symbolic
@@
 However, {\sc Spark} is a monolithic module, that does not support sharing its
-internal 3D model to other external components. On the contrary, \uwds focuses
-on offering a shared and distributed representation within the robot's software
-architecture. Previous work on distributed scene graphs~\cite{naef2003blue} has
-been applied to robotics to provide a shared 3D representation of the robot's
-environment (like the \emph{Robot Scene Graph}~\cite{blumenthal2013scene} or 
-\emph{Deep State Representation}~\cite{bustos2016unified} approaches).
+internal 3D model to other external components. In contrast, \uwds focuses on
+offering a shared and distributed representation of the environment within the
+robot's software architecture.
 
-\uwds further extends this line of work by providing the ability to manipulate
-several \emph{alternative 3D worlds}, corresponding to filtered or hypothetical
-\emph{views} on the perceived model of the environment.
+Work on distributed scene graphs~\cite{naef2003blue} has been previously applied
+to robotics to provide a shared 3D representation of the robot's environment
+(for instance, the \emph{Robot Scene Graph}~\cite{blumenthal2013scene} or the
+\emph{Deep State Representation} proposed in~\cite{bustos2016unified}).  \uwds
+offers a similar distribution mechanism for 3D scene graphs and extends it to
+temporal representations. Besides, \uwds further extends this line of work by
+providing the ability to create, manipulate and share \emph{multiple alternative
+worlds}, corresponding for instance to filtered or hypothetical \emph{views} on
+the initial, perceived model of the environment.
 

### 2017-02-27 paper.tex
Proof read part 1

@@
 include modules like geometric reasoners (that compute spatial and topological
-relations between objects), motion planners, action recognition.
+relations between objects), motion planners or action recognition modules.
 
-These different components typically exhibit different needs in terms of model,
+These components exhibit different needs in terms of model,
 like different nominal spatial and/or temporal resolutions. For instance, a 3D
@@
 lower the computational load while planning, while a module assessing the
-visibility of objects might need on the contrary a high-resolution model of the
-environment for accurate 3D rendering. This requirement of multiple
+visibility of objects might need on the contrary  high-resolution models 
+for accurate 3D rendering. This requirement of multiple
 task-specific representations has been framed as the need for \emph{deep
-representation} by Beetz~\cite{beetz2010towards}.
+representations} by Beetz~\cite{beetz2010towards}.
 
@@
 Traditional robotic middlewares like ROS are not particularly well-suited to
-deal with these different needs: full geometric data can be represented, but are not
-first-class citizens (for instance, basic tasks like displaying a 3D mesh at a dynamic
-position are not particularly easy with ROS) and critically, representing and reasoning
-on alternative models of the environment is not directly feasible.
+deal with these different needs: full geometric data can be represented, but are
+not first-class citizens (basic tasks like displaying a 3D mesh at a dynamic
+position are not particularly easy to perform with ROS) and critically,
+representing and reasoning on alternative states of the environment is not
+directly feasible.
 
-Representing \emph{alternative states} is however often highly desirable. For instance,
+Representing alternative states is however often highly desirable. For instance,
 software components manipulating environment models typically perform better if
 the models are physically consistent. However, low-level perception inaccuracies
-often introduce hard-to-avoid physical inconsistencies (like
-detected objects floating in the air, or on the contrary wrongly inset into
-other objects). Therefore, a post-process stage (for instance, using a physics
-simulation engine) might be needed to displace the objects seen
-by the robot into physically-correct positions. Implemented with a classical
-approach (for instance, using ROS TF frames), we would get an object {\tt book}
-located at the frame {\tt book\_frame}, and a secondary frame {\tt
-book\_frame\_corrected} computed by the physics engine. It is easy to see why
-such an approach would be quickly confusing (where is in fact my book?) and would not
-scale well, with the robot's 3D model cluttered with spurious frames.
+often introduce hard-to-avoid physical inconsistencies (like detected objects
+floating in the air, or on the contrary wrongly inset into other objects).
+Therefore, a post-process stage (for instance, using a physics simulation
+engine) is needed to displace the objects seen by the robot into
+physically-correct positions. Implemented with a classical approach (for
+instance, using ROS TF frames), we would represent an object {\tt book} with two
+frames: the original frame (\eg {\tt book\_frame\_raw}) and a second one
+computed by the physics engine (\eg {\tt book\_frame\_corrected}). Such an
+approach leads to the robot's 3D model being cluttered with multiple frames and
+does not scale well.
 
-Another relevant example pertains to geometric task planning: a geometric task
-planner typically need to reason over hypothetical future states of the
-environment (``what happen if I move this glass onto that pile of books?'').
-The planner might want to quickly generate possible future states, that
-themselves might require further processing (for instance, running a physics
-simulation). This use-case evidences the need for a flexible representation
-system, where models are derived from each others, with partial modifications
-and different timescales.
+Another example pertains to geometric task planning: a geometric task planner
+typically needs to reason over hypothetical future states of the environment
+(``what happen if I move this glass onto that pile of books?'').  The planner
+generates many possible future states, that themselves might require further
+processing (for instance, running a physics simulation). Such a tool would
+benefit a flexible representation system, where models are derived from each
+others, with partial modifications and different timescales.
 
@@
 cognitive skill that relies on the ability for an agent to project itself in
-others' `eyes' to estimate what they see from their perspective. While
-perspective taking has been previously implemented in robotics by temporary placing
+others' `eyes' to estimate what they see from their perspectives. 
+Perspective taking has been previously implemented in robotics by temporary placing
 virtual cameras at eyes position for each of the humans tracked by the
-robot~\cite{ros2010solving}. a the face of  swapping the different viewpoints
-to compute visibility is not directly possible
+robot~\cite{ros2010solving}. While acceptable for simple cases, such an approach
+does not permit to maintain truly independent models of the environment for each
+agent, and in particular does not permit false-beliefs
+modelling~\cite{lemaignan2015mutual} whereas separate, independent world models
+would effectively support such a skill.
 
-Geometric \emph{pre-supposition accommodation} makes another interesting case
-for alternative worlds representation. \emph{Pre-supposition accommodation}
+Lastly, geometric \emph{pre-supposition accommodation} makes another interesting
+case for alternative worlds representation. \emph{Pre-supposition accommodation}
 comes originally from linguistics, where it describes the mechanism by which
@@
 representations, we call pre-supposition accommodation the ability for an agent
-to adjust its model so that it matches some external contextual constraint. For
-instance, if A tells B to ``catch the red balloon behind you'', B adds a
-representation of a red balloon behind him, even without actually seeing the
-balloon: B accommodates the pre-supposition of a red balloon being present
-behind himself. Endowing robots with this capability has been touched by
-Mavridis \etal within their multi-modal \emph{Grounded Situation
-Model}~\cite{Mavridis2006}. However, to the best of our knowledge, a general
-framework which would enable robots to derive imaginary worlds by extending of
-modifying existing ones has not been proposed so far.
+to adjust its model so that it matches some contextual constraint. For instance,
+if A tells B to ``catch the red balloon behind you'', B might create a
+representation of an imaginary red balloon, placed behind him, even without
+actually observing the balloon: B \emph{accommodates} the \emph{pre-supposition}
+of a red balloon being present behind himself. Endowing robots with this
+capability has been touched by Mavridis \etal within their multi-modal
+\emph{Grounded Situation Model}~\cite{Mavridis2006}. However, to the best of our
+knowledge, a general framework which would enable robots to accommodate spatial
+and temporal pre-suppositions by deriving imaginary worlds from existing ones
+has not been proposed so far.
 
-The main contribution of this work is a generic approach to represent and share
-\textbf{multiple parallel representations of the world}. \uwds does so by
-permitting clients to clone an existing world, modify it, and re-share it,
-without the cost of duplicating geometric data (as explained in
-section~\ref{design}). By organizing clients in a network
+\uwds addresses this need and the main contribution of this work is a generic
+approach to \textbf{represent and share multiple parallel representations of the
+world}. \uwds does so by permitting clients to clone existing worlds, modify
+them, and re-share them, without the cost of duplicating geometric data (as
+explained in section~\ref{design}). By organizing clients in a network
 (Figure~\ref{fig|scene}), worlds can be made dependent on each other, resulting

### 2017-02-27 paper.tex
Proof read part 2

@@
     \includegraphics[width=\linewidth]{overview}
-    \caption{Schema of a possible \uwds network: eight possible, user-written
-    \emph{clients} (in blue) are sharing environment models through four
-    independent \emph{worlds} (made of joint spatial and temporal models). This architecture
-    enables successive and modular refinement of the models
-    (\emph{cascading} situation assessment), effectively adapted to each
-    client's needs (yellow nodes represent other possible components in the system that do not
-    interact directly with the \uwds network).}
+    \caption{Schema of a possible \uwds network: eight (user-written \&
+    architecture specific) \emph{clients} (in blue) are sharing environment
+    models through four independent \emph{worlds} (made of joint spatial and
+    temporal models). This architecture enables successive and modular
+    refinement of the models (\emph{cascading} situation assessment),
+    effectively adapted to each client's needs. \textit{Dashed yellow nodes represent
+    other possible components in the system that do not interact directly with
+    the \uwds network}.}
 
@@
 them, and re-share them, without the cost of duplicating geometric data (as
-explained in section~\ref{design}). By organizing clients in a network
+explained at section~\ref{design}). By organizing clients in a network
 (Figure~\ref{fig|scene}), worlds can be made dependent on each other, resulting
@@
 
-Figure~\ref{fig|scene} pictures a typically \uwds topology: a graph (that
+Figure~\ref{fig|scene} pictures a typical \uwds topology: a graph (that
 happens to be a directed acyclic graph on Figure~\ref{fig|scene}, but does not
@@
 clients (like a 3D visualization tool) are provided with the \uwds package (see
-section~\ref{std_clients}), but clients are otherwise written by the end users
+section~\ref{std_clients}). Clients are otherwise written by the end users
 using the \uwds client API (see section~\ref{api}).
@@
 automatically see updates broadcasted by other clients connected to the same
-worlds.
+world.
 
@@
 worlds arise from the clients' connections. For instance, filters effectively
-create a dependency between worlds. A client like the \textit{Physics-based
-position correction} client create a dependency between the world {\tt base} (which
+create a dependency between worlds. On Figure~\ref{fig|scene}, the \textit{Physics-based
+position correction} client creates a dependency between the world {\tt base} (which
 represents here the result of raw sensor fusion) and the world {\tt corrected}
@@
 
-\emph{Worlds} contain both a geometric model and a temporal model. The geometric
+Worlds contain both a geometric model and a temporal model. The geometric
 model is represented as a scene graph. The scene graph has a unique root node,
@@
 
-Nodes in the scene graph have four possible types: \textbf{objects} that represent
-concrete physical objects (typically with one or several associated 3D meshes);
-\textbf{entities} that represent abstract entities like reference frames or
-groups of objects; \textbf{perspectives} that represent viewpoints onto the
-scene (like cameras or human gaze); and \textbf{fields} that represent scalar or
-vector fields (like the visibility of an object, the working space of robot,
-etc. -- note that fields are not yet implemented in the current version of
-\uwds, see section~\ref{futurework}).
+Nodes in an \uwds scene graph have four possible types: \textbf{objects} that
+represent concrete physical objects (typically with one or several associated 3D
+meshes); \textbf{entities} that represent abstract entities like reference
+frames or groups of objects; \textbf{perspectives} that represent viewpoints
+onto the scene (like cameras or human gaze); and \textbf{fields} that represent
+scalar or vector fields (like the visibility of an object, the working space of
+robot, etc. -- note that fields are not yet implemented in the current version
+of \uwds, see section~\ref{futurework}).
 
 Every node has a unique ID, a parent, a 3D transformation relative to the parent
-and an optional name. \emph{Object} nodes optionally store as well pointers to their
-associated meshes. Importantly, mesh data (or other geometric datasets like
-point clouds) are \emph{not} stored within the nodes. \uwds represents geometric
-data as immutable data, labelled with their hash (preventing data duplication).
-Nodes only store the hash of their geometric data, and the actual data is pulled
-by client whenever they actually need it (for rendering for instance).
+and an optional name. \emph{Object} nodes optionally store as well pointers to
+their associated meshes. Importantly, mesh data (or other geometric datasets
+like point clouds) are \emph{not} stored within the nodes themselves. \uwds
+represents geometric data as immutable data, labelled with their hash
+(preventing \textit{de facto} data duplication).  Nodes only store the hash
+corresponding to the desired geometric data, and the actual data is pulled from
+the server by the clients whenever they actually need it (for rendering for
+instance).
 
@@
 and synchronized amongst the clients in the same way as the scene graph.
-Client can record in the timeline both \emph{events} (duration-less state) and
+Clients can record in the timeline both \emph{events} (duration-less states) and
 \emph{situations}, \ie states with a start time and a (possibly open-ended) end time.
@@
 detection module, a human skeleton tracker, etc.), and for a {client library}
-that makes it easy to interface existing software components with the network.
+that makes it possible to interface existing software components with the network.
 
 Critically, the network is essentially hidden to the client: from the user
-perspective, the environment model is manipulated as local data structures (the
-\emph{worlds}). Modifications to the model are asynchronously synchronised with
+perspective, the environment model is manipulated as a local data structure (see
+Listing~\ref{lst|pythonapi}). Modifications to the model are asynchronously synchronised with
 a central server (the {\tt underworlded} daemon) and broadcasted to every other
-client in the network.
+client connected to the same world.
 
@@
 Google's gRPC message passing framework\footnote{\url{http://www.grpc.io/}}, ensuring
-high throughput, reliability, cross-platform/cross-language support. The \uwds
-API is specifically discussed hereafter, section~\ref{api}.
-
-\subsubsection*{Performance considerations}
+high throughput, reliability and cross-platform/cross-language support. The \uwds
+API is specifically discussed hereafter, at section~\ref{api}.
 
@@
 including large geometric datasets, like meshes) in real-time. \uwds itself does
-not perform many CPU intensive tasks (CPU intensive processing tasks -- sensor fusion, physics
-simulation, etc. -- are realized by the clients) and as such, the performance
-bottleneck is essentially the network's data throughput.
-
-In that regard, one of the simple yet critical optimization performed by \uwds
-is automatic caching of mesh data. Mesh data are not transmitted when nodes are
-updated; only a hash value of the mesh data. The client can then request the
-full data whenever they actually need it.
+not perform many CPU intensive tasks (CPU intensive processing tasks -- sensor
+fusion, physics simulation, etc. -- are performed by the clients themselves) and
+as such, the performance bottleneck is essentially the network's data
+throughput.  In that regard, one of the simple yet critical optimization
+performed by \uwds is automatic caching of mesh data. Mesh data are not
+transmitted when nodes are updated; only a hash value of the mesh data. The
+client can then request the full data whenever they actually need it.
 
@@
 
-As mentioned in section~\ref{arch}, \uwds uses Google's gRPC as message passing
+As mentioned at section~\ref{arch}, \uwds uses Google's gRPC as message passing
 protocol. The protocol is explicitly defined (using the \emph{protocol

### 2017-02-27 paper.tex
Proof read part 3

@@
 
-As mentioned at section~\ref{arch}, \uwds uses Google's gRPC as message passing
-protocol. The protocol is explicitly defined (using the \emph{protocol
+As mentioned, \uwds uses Google's gRPC as message passing protocol. The protocol
+is explicitly defined (using the \emph{protocol
 buffers}\footnote{\url{https://developers.google.com/protocol-buffers/}}
-interface definition language), and bindings to various languages
-and platforms can be automatically generated from the protocol definition file
-(as of Feb 2017, gRPC can generate bindings for C, C++, C\#, Node.js, PHP, Ruby,
-Python, Go and Java on Windows, Mac and Linux).
-
-The cross-platform/cross-language support of gRPC is especially welcomed in the
-academic context, as it offers ease and flexibility to plug a variety of
-pre-existing components into an \uwds network.
+interface definition language), and bindings to various languages and platforms
+can be automatically generated from the protocol definition file (as of Feb
+2017, gRPC can generate bindings for C, C++, C\#, Node.js, PHP, Ruby, Python, Go
+and Java on Windows, Mac and Linux).  The cross-platform/cross-language support
+of gRPC is especially welcomed in the academic context, as it offers ease and
+flexibility to plug a variety of pre-existing components into an \uwds network.
 
@@
 of the robot environment). In particular, the asynchronous fetching (and
-conversely, remote updating) of nodes and time-related objects should be
-entirely hidden from the user, and managed instead by the \uwds client library.
+conversely, remote updating) of nodes and time-related objects is typically
+hidden from the user, and managed instead by the \uwds client library.
 
-\uwds currently offers such a high-level client library for Python only. A C++
-library is being developed (currently in 'alpha' state).
-Listing~\ref{lst|pythonapi} gives a complete example of an \uwds client
-performing simple filtering: the client continuously ``listen'' for changes in
-an input world, hide some items (in this case, items whose name contain the
-string ``{\tt wall}''), and forward all other changes to an output world,
-effectively making the output world a copy of the input world with all walls
-removed.
+\uwds currently offers such a high-level client library for Python only (a C++
+library is under development).  Listing~\ref{lst|pythonapi} gives a complete
+example of an \uwds client performing simple filtering: the client continuously
+listen for changes in an input world, removes some objects (in this case, items
+whose name contain the string ``{\tt wall}''), and forward all other changes to
+an output world, effectively making the output world a copy of the input world
+with all walls removed.
 
@@
 
+# by default, connect to the server on localhost
 with underworlds.Context("walls_filter") as ctx:
@@
 \end{pythoncode}
-    \caption{Example of a simple yet complete \uwds filter, written in Python, and named {\tt
-    walls\_filter}: the client connects to the \uwds network, walk through the
-    scene graph of world {\tt example\_in}, filter out some objects, and publish
-    the remaining objects in the world {\tt example\_out}.}
+    \caption{Example of a simple yet complete \uwds filter, written in Python:
+    the client connects to the \uwds network, walk through the scene graph of
+    world {\tt example\_in}, filter out some objects, and publishes the
+    remaining objects in the world {\tt example\_out}.}
 
@@
 \label{std_clients}
+
+The \uwds package provides several standard clients to perform common tasks on
+\uwds networks.
 
@@
 to visualize the content of a scene is desirable. \uwds provides a standard
-client, {\tt uwds view}, that performs real-time 3D rendering of the scene,
+client, {\tt uwds view}, that performs real-time 3D rendering of worlds,
 using OpenGL (Figure~\ref{fig|uwds-view}).
@@
 This tool also support basic object manipulations, that are broadcasted to the
-other \uwds clients as expected.
+other \uwds clients connected to the same world.
 
@@
 server, stores it internally (\eg as an OpenGL buffer), and renders it. Whenever
-the node transformation is updated, the 3D viewer can re-render it without
+the node transformation is updated, the 3D viewer can re-render the node without
 having to download the mesh from the server again.
@@
 Often, objects manipulated by the robot have known meshes with corresponding CAD
-models. In these case, \uwds provides a tool, {\tt uwds load}, that loads a mesh
-into a \uwds network (and optionally, creates a node) from a large range of 3D
-format\footnote{The underlying import capability is provided by the {\sc assimp}
-library.  \url{http://assimp.sourceforge.net/}} (including Collada, FBX, OBJ,
-Blender).
+models that can be conveniently pre-loaded. In these case, \uwds provides a
+tool, {\tt uwds load}, that loads a mesh into a \uwds network (and optionally,
+creates a node) from a large range of 3D format (including Collada, FBX, OBJ,
+Blender)\footnote{The underlying import
+capability is provided by the {\sc assimp} library.
+\url{http://assimp.sourceforge.net/}}.
 
@@
 provide an user-friendly overview of the system's graph with the connections
-between the clients and the worlds as well as their activity.
+between the clients and the worlds, as well as their activity.
 
@@
 
-The {\tt uwds tf} root client continuously monitor the ROS TF tree, and mirror
-TF frames as nodes in a given \uwds world. A node is first created if none
+The {\tt uwds tf} client continuously monitor the ROS TF tree, and mirrors
+TF frames as nodes in the desired \uwds world. A node is first created if none
 matches a given TF frame, and its transformation is subsequently updated,
@@
 
-\uwds also implement an efficient algorithm to assess object visibility from a
-specific viewpoint (\ie from a given \emph{camera} node). The algorithm (color
+\uwds also implements an efficient algorithm to assess object visibility from a
+specific viewpoint (\ie from a given \emph{perspective} node). The algorithm (color
 picking) enables fast (single pass) computation of the visibility of every
-object in the scene, while providing control regarding how much of an object
-should be actually visible to be considered as globally visible. The
+object in the scene, while providing control regarding how many pixels should be
+actually visible for the object to be considered globally visible. The
 command-line tool {\tt uwds visibility} returns the list of visible objects from
-the point of view of every camera in a given world, and \uwds also provides the
+the point of view of each camera in a given world, and \uwds also provides the
 helper class {\tt VisibilityMonitor} to programmatically access visibility
@@
 
-When integrated into a filter node, visibility computation enables to easily
+When integrated into a filter node, visibility computation allows to easily
 create new worlds representing the estimated perspectives of the different

### 2017-02-27 paper.tex
Proof read part 4

@@
     \includegraphics[width=\linewidth]{l2tor-photo2}
-    \caption{A user manipulating objects on RFID mat while an operator monitors
-    the interaction via an {\tt uwds view} client}
+    \caption{A user manipulating objects on top of a RFID mat (providing object
+    localization) with a NAO robot, while an operator monitors
+    the interaction on a tablet that display an {\tt uwds view} client.}
 
@@
 in order to conceptualize and visualize the spatial relations and visibility of
-the physical objects that users interact with. In one of the scenarios of this
-project, users are instructed by a NAO robot to manipulate a number of
+the physical objects that participants interact with. In one of the scenarios of this
+project, participants are instructed by a NAO robot to manipulate a number of
 Duplo-like animals on top of a printed background paper which serves as a
-landscape (See \ref{fig|l2torexample}). The robot instructions are based on
+landscape (Figure~\ref{fig|l2torexample}). The robot instructions are based on
 pedagogical strategies to promote spatial relation learning by object
@@
 \textit{on top} or \textit{next to} the cube or put the giraffe \textit{inside}
-the house.  For that reason, the robot and the operator should be able to access
-whether the user performs the required task and at the same time envisage their
-perspective. A number of \uwds clients have been developed in order to provide a
-distributed model of the environment and the robot as described in the next
-section.
+the house.  To assess the performance, the robot and the operator need to
+compute spatial relation between objects from the visual perspective of the
+participant.
 
 \subsection*{Architecture}
+
 \begin{figure}
     \centering
-    \includegraphics[width=\linewidth]{l2tor}
-    \caption{Schema of the \uwds architecture used in L2TOR project pilot study. }
+    \includegraphics[width=0.9\linewidth]{l2tor}
+    \caption{Schema of the \uwds architecture used in the L2TOR project. }
     \label{fig|l2torarchitecture}
@@
 
-The scenario consists of a number of \uwds clients distributed across two
-computers that communicate with each other in order to provide an updated
-representation of the world. Figure \ref{fig|l2torarchitecture} illustrates the
-high level communication of the \uwds clients that form the architecture of the
-L2TOR spatial relations platform.
-
-L2TOR \uwds clients:
+The interaction scenario requires a number of \uwds clients. The client are
+distributed across two computers (the robot and a Microsoft Windows-operated
+tablet) that communicate with each other in order to synchronize the worlds
+representations. Figure~\ref{fig|l2torarchitecture} illustrates the topology of
+the \uwds network, that consists in the following clients:
 
 \begin{itemize}
-    \item \textbf{Kinect Head Tracker} Captures the location and the direction
-        of the user's head - Camera point
+    \item \textbf{Kinect Head Tracker}: captures the location and the direction
+        of the user's head -- maintains the {\tt human} \emph{Perspective} node;
 
-    \item \textbf{Object RFID tracker} Updates the locations of the virtual
-        objects based on RFID data
+    \item \textbf{Object RFID tracker}: updates the locations of the virtual
+        objects based on RFID data;
 
-    \item \textbf{NAO model} Virtual representation of the Robot - Camera poit
+    \item \textbf{NAO joint state publisher}: virtual representation of the
+        robot -- maintains the {\tt robot} \emph{Perspective} node;
 
-    \item \textbf{Visibility computation} Computes visibility based on camera
-        points
+    \item \textbf{Visibility computation}: computes visibility based on
+        \emph{Perspective} nodes;
 
-    \item \textbf{Spatial Relations computation} Extracts spatial relations
-        between the virtual objects
+    \item \textbf{Spatial relations computation}: extracts symbolic spatial relations
+        between the virtual objects;
 
-    \item \textbf{Game/Rules computation} Interaction manager based on \uwds
-        clients' input
+    \item \textbf{Game/Rules computation}: interaction manager based on \uwds
+        clients' input.
 
@@
 accommodate this change, we replaced the object detection \uwds client with a 3D
-virtual object detection while the rest of the system remains the same. Finally,
-we are evaluating the use of a hybrid platform that utilizes both virtual and
+virtual object detection while the rest of the system remains the same.
+
+Finally, we are evaluating the use of a hybrid platform that utilizes both virtual and
 physical objects at the same time.

### 2017-02-28 paper.tex
at->in

@@
 them, and re-share them, without the cost of duplicating geometric data (as
-explained at section~\ref{design}). By organizing clients in a network
+explained in section~\ref{design}). By organizing clients in a network
 (Figure~\ref{fig|scene}), worlds can be made dependent on each other, resulting
@@
 high throughput, reliability and cross-platform/cross-language support. The \uwds
-API is specifically discussed hereafter, at section~\ref{api}.
+API is specifically discussed hereafter, in section~\ref{api}.
 

### 2017-02-28 paper.tex
Update on Overleaf.

@@
     set of \emph{worlds} that can each have different spatial and temporal
-    granularities, and may depend on each others.  \uwds also provides a set of
+    granularities, and may depend on each other.  \uwds also provides a set of
     high-level client libraries and tools to introspect and manipulate the
@@
 framework\footnote{\url{https://github.com/severin-lemaignan/underworlds}} that
-enables robot programmers to build and refine in real-time spatial and temporal
-models of the environment surrounding a robot. \uwds makes it possible to share
+enables robot programmers to build and refine spatial and temporal
+models of the environment surrounding a robot in real-time. \uwds makes it possible to share
 these world models amongst the software components running on the robot.
-Besides, \uwds permits to represent and manipulate in a distributed manner
-\emph{multiple alternatives} to the current, perceived world model. For
+Additionally, \uwds enables users to represent and manipulate
+\emph{multiple alternatives} to the current, perceived world model in a distributed manner. For
 instance, the world with some objects filtered out; the world `viewed' from the
-perspective of another agent; an hypothetical world resulting from the simulated
+perspective of another agent; a hypothetical world resulting from the simulated
 application of a plan, etc.
@@
     \includegraphics[width=\linewidth]{overview}
-    \caption{Schema of a possible \uwds network: eight (user-written \&
-    architecture specific) \emph{clients} (in blue) are sharing environment
-    models through four independent \emph{worlds} (made of joint spatial and
+    \caption{Schema of a possible \uwds network: eight \emph{clients} (user-written \&
+    architecture specific; in blue) are sharing environment
+    models through four independent \emph{worlds} (made from joint spatial and
     temporal models). This architecture enables successive and modular
@@
 symbolic layer~\cite{lemaignan2016artificial}. This geometric model is built
-from 3D CAD models of the objects, furnitures and robots, and full body, rigged
+from 3D CAD models of the objects, furniture and robots, and full body, rigged
 models of humans.  It is updated at run-time by the robot's sensors.
@@
 However, {\sc Spark} is a monolithic module, that does not support sharing its
-internal 3D model to other external components. In contrast, \uwds focuses on
+internal 3D model with other external components. In contrast, \uwds focuses on
 offering a shared and distributed representation of the environment within the
@@
 providing the ability to create, manipulate and share \emph{multiple alternative
-worlds}, corresponding for instance to filtered or hypothetical \emph{views} on
+worlds}. As an example, these could correspond to filtered or hypothetical \emph{views} on
 the initial, perceived model of the environment.
@@
 
-These components exhibit different needs in terms of model,
+These components exhibit different needs in terms of representation,
 like different nominal spatial and/or temporal resolutions. For instance, a 3D
@@
 lower the computational load while planning, while a module assessing the
-visibility of objects might need on the contrary  high-resolution models 
+visibility of objects might need high-resolution models 
 for accurate 3D rendering. This requirement of multiple
@@
 
-Traditional robotic middlewares like ROS are not particularly well-suited to
+% JK: like software, the plural of middleware is middleware :)
+Traditional robotic middleware, like ROS, is not particularly well-suited to
 deal with these different needs: full geometric data can be represented, but are
-not first-class citizens (basic tasks like displaying a 3D mesh at a dynamic
-position are not particularly easy to perform with ROS) and critically,
+not first-class citizens (a basic task like displaying a 3D mesh at a dynamic
+position is not particularly easy to perform with ROS) and critically,
 representing and reasoning on alternative states of the environment is not
@@
 often introduce hard-to-avoid physical inconsistencies (like detected objects
-floating in the air, or on the contrary wrongly inset into other objects).
+floating in the air, or wrongly inset into other objects).
 Therefore, a post-process stage (for instance, using a physics simulation
@@
 (``what happen if I move this glass onto that pile of books?'').  The planner
-generates many possible future states, that themselves might require further
+generates many possible future states, that in turn might require further
 processing (for instance, running a physics simulation). Such a tool would
 benefit a flexible representation system, where models are derived from each
-others, with partial modifications and different timescales.
+other, with partial modifications and different timescales.
 
@@
 taking} is important (a prototypical example being the game `I spy with my
-little eyes', as implemented in~\cite{ros2010which}). Perspective taking is a
+little eye', as implemented in~\cite{ros2010which}). Perspective taking is a
 cognitive skill that relies on the ability for an agent to project itself in
 others' `eyes' to estimate what they see from their perspectives. 
-Perspective taking has been previously implemented in robotics by temporary placing
-virtual cameras at eyes position for each of the humans tracked by the
+Perspective taking has previously been implemented in robotics by temporarily placing
+virtual cameras at eye locations for each of the humans tracked by the
 robot~\cite{ros2010solving}. While acceptable for simple cases, such an approach
-does not permit to maintain truly independent models of the environment for each
-agent, and in particular does not permit false-beliefs
+does not maintain truly independent models of the environment for each
+agent, and in particular does not permit false-belief
 modelling~\cite{lemaignan2015mutual} whereas separate, independent world models
@@
 case for alternative worlds representation. \emph{Pre-supposition accommodation}
-comes originally from linguistics, where it describes the mechanism by which
+originally comes from linguistics, where it describes the mechanism by which
 \emph{context is adjusted [...] to accept [...] a sentence that imposes certain
@@
 processed}~\cite{vonfintel2008presupposition}. In the context of spatio-temporal
-representations, we call pre-supposition accommodation the ability for an agent
+representations, we call pre-supposition accommodation the ability of an agent
 to adjust its model so that it matches some contextual constraint. For instance,
@@
 of a red balloon being present behind himself. Endowing robots with this
-capability has been touched by Mavridis \etal within their multi-modal
+capability has been touched upon by Mavridis \etal within their multi-modal
 \emph{Grounded Situation Model}~\cite{Mavridis2006}. However, to the best of our
@@
 them, and re-share them, without the cost of duplicating geometric data (as
-explained in section~\ref{design}). By organizing clients in a network
+explained in section~\ref{design}). By organising clients in a network
 (Figure~\ref{fig|scene}), worlds can be made dependent on each other, resulting
@@
 model is represented as a scene graph. The scene graph has a unique root node,
-to which a tree of other nodes is parented.
+in which a tree of other nodes is parented.
 
@@
 corresponding to the desired geometric data, and the actual data is pulled from
-the server by the clients whenever they actually need it (for rendering for
+the server by the clients whenever they actually need it (rendering for
 instance).
@@
 world also store the world's \emph{timeline}. This data structure is shared
-and synchronized amongst the clients in the same way as the scene graph.
+and synchronised amongst the clients in the same way as the scene graph.
 Clients can record in the timeline both \emph{events} (duration-less states) and
@@
     \includegraphics[width=\linewidth]{uwds-screenshot}
-    \caption{Screenshot of the {\tt uwds view} 3D visualization and manipulation
+    \caption{Screenshot of the {\tt uwds view} 3D visualisation and manipulation
     tool. In this particular example, the 3D meshes have been pre-loaded using
@@
 As previously mentioned, worlds are composite data structures comprising of a
-scene graph and a timeline. These data structures are synchronized using
+scene graph and a timeline. These data structures are synchronised using
 Google's gRPC message passing framework\footnote{\url{http://www.grpc.io/}}, ensuring
@@
 as such, the performance bottleneck is essentially the network's data
-throughput.  In that regard, one of the simple yet critical optimization
+throughput.  In that regard, one of the simple yet critical optimisation
 performed by \uwds is automatic caching of mesh data. Mesh data are not
@@
 
-\subsubsection{3D Visualization and manipulation}
+\subsubsection{3D Visualisation and manipulation}
 
@@
 graphics library like OpenGL). However, for all practical purposes, the ability
-to visualize the content of a scene is desirable. \uwds provides a standard
+to visualise the content of a scene is desirable. \uwds provides a standard
 client, {\tt uwds view}, that performs real-time 3D rendering of worlds,
@@
 As such, any client can load 3D data that is automatically made available to the
-whole network. For example, an object recognition module recognizes a new
+whole network. For example, an object recognition module recognises a new
 object. It sends the mesh to the \uwds server and creates a new node which
@@
 
-Specialized command-line tools are also available to list at run-time the worlds
+Specialised command-line tools are also available to list at run-time the worlds
 and their content ({\tt uwds ls}) or to display detailed informations on a
@@
     \caption{A user manipulating objects on top of a RFID mat (providing object
-    localization) with a NAO robot, while an operator monitors
+    localisation) with a NAO robot, while an operator monitors
     the interaction on a tablet that display an {\tt uwds view} client.}
@@
 \uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor}
-in order to conceptualize and visualize the spatial relations and visibility of
+in order to conceptualise and visualise the spatial relations and visibility of
 the physical objects that participants interact with. In one of the scenarios of this
@@
 distributed across two computers (the robot and a Microsoft Windows-operated
-tablet) that communicate with each other in order to synchronize the worlds
+tablet) that communicate with each other in order to synchronise the worlds
 representations. Figure~\ref{fig|l2torarchitecture} illustrates the topology of
@@
 
-Finally, we are evaluating the use of a hybrid platform that utilizes both virtual and
+Finally, we are evaluating the use of a hybrid platform that utilises both virtual and
 physical objects at the same time.

### 2017-02-28 paper.tex
Update on Overleaf.

@@
 Software components accessing \uwds worlds are called clients. Some standard
-clients (like a 3D visualization tool) are provided with the \uwds package (see
+clients (like a 3D visualisation tool) are provided with the \uwds package (see
 section~\ref{std_clients}). Clients are otherwise written by the end users
@@
 create and update worlds (`write-only' client, like the \emph{Sensor fusion} and
-\emph{ROS interface} clients on Figure~\ref{fig|scene}); \textbf{leaf clients}
+\emph{ROS interface} clients in Figure~\ref{fig|scene}); \textbf{leaf clients}
 which on the contrary only read worlds, without modifying them (like the
@@
 
-Each world is technically independent from every others. Dependencies between
+Each world is technically independent from all others. Dependencies between
 worlds arise from the clients' connections. For instance, filters effectively
@@
 frames or groups of objects; \textbf{perspectives} that represent viewpoints
-onto the scene (like cameras or human gaze); and \textbf{fields} that represent
+of the scene (like cameras or human gaze); and \textbf{fields} that represent
 scalar or vector fields (like the visibility of an object, the working space of
-robot, etc. -- note that fields are not yet implemented in the current version
+robot, \etc -- note that fields are not yet implemented in the current version
 of \uwds, see section~\ref{futurework}).
@@
 Complementing the spatial representation encapsulated in the scene graph, each
-world also store the world's \emph{timeline}. This data structure is shared
+world also stores the world's \emph{timeline}. This data structure is shared
 and synchronised amongst the clients in the same way as the scene graph.
-Clients can record in the timeline both \emph{events} (duration-less states) and
-\emph{situations}, \ie states with a start time and a (possibly open-ended) end time.
+Clients can record both \emph{events} (duration-less states) and
+\emph{situations} in the timeline, \ie states with a start time and a (possibly open-ended) end time.
 
@@
 snapshot is associated to the event, which allows clients to effectively retrieve past
-states of the world. This capability is thought to be primarly used by \uwds clients
+states of the world. This capability is anticipated to be primarily used by \uwds clients
 performing action recognition.
@@
     tool. In this particular example, the 3D meshes have been pre-loaded using
-    {\tt uwds load}. Their positions are then updated at run time using the
+    {\tt uwds load}. Their positions are then updated at run-time using the
     robot's sensors and proprioception (joint state).}
@@
 client connected to the same world.
+% JK: underworlded - intentional? (I guess yes, but checking)
 
-As previously mentioned, worlds are composite data structures comprising of a
+As previously mentioned, worlds are composite data structures comprised of a
 scene graph and a timeline. These data structures are synchronised using
@@
 not perform many CPU intensive tasks (CPU intensive processing tasks -- sensor
-fusion, physics simulation, etc. -- are performed by the clients themselves) and
+fusion, physics simulation, \etc -- are performed by the clients themselves) and
 as such, the performance bottleneck is essentially the network's data
-throughput.  In that regard, one of the simple yet critical optimisation
+throughput.  In that regard, one of the simple yet critical optimisations
 performed by \uwds is automatic caching of mesh data. Mesh data are not
 transmitted when nodes are updated; only a hash value of the mesh data. The
-client can then request the full data whenever they actually need it.
+client can then request the full data whenever it is actually needed.
 
@@
 2017, gRPC can generate bindings for C, C++, C\#, Node.js, PHP, Ruby, Python, Go
-and Java on Windows, Mac and Linux).  The cross-platform/cross-language support
-of gRPC is especially welcomed in the academic context, as it offers ease and
+and Java, on Windows, Mac and Linux).  The cross-platform/cross-language support
+of gRPC is especially welcome in the academic context, as it offers ease and
 flexibility to plug a variety of pre-existing components into an \uwds network.
@@
 example of an \uwds client performing simple filtering: the client continuously
-listen for changes in an input world, removes some objects (in this case, items
-whose name contain the string ``{\tt wall}''), and forward all other changes to
+listens for changes in an input world, removes some objects (in this case, items
+whose name contain the string ``{\tt wall}''), and forwards all other changes to
 an output world, effectively making the output world a copy of the input world
@@
 
-        for node in in_word.scene.nodes:
+        for node in in_world.scene.nodes:
             if("wall" not in node.name):
@@
     \caption{Example of a simple yet complete \uwds filter, written in Python:
-    the client connects to the \uwds network, walk through the scene graph of
-    world {\tt example\_in}, filter out some objects, and publishes the
+    the client connects to the \uwds network, walks through the scene graph of
+    world {\tt example\_in}, filters out some objects, and publishes the
     remaining objects in the world {\tt example\_out}.}
@@
 Interestingly, while \uwds deals with 3D geometries and scenes, it does
-represent 3D entities purely as data structures: no visual representation is
+represent 3D entities purely as data structures; no visual representation is
 involved (and as such, the \uwds server and core libraries do not depend on any
@@
 
-This tool also support basic object manipulations, that are broadcasted to the
+This tool also supports basic object manipulations, that are broadcasted to the
 other \uwds clients connected to the same world.
@@
 references the mesh. At the next frame update, the 3D viewer sees a new node
-referencing a yet unknown mesh. The viewer request the mesh data from the
+referencing an unknown mesh. The viewer requests the mesh data from the
 server, stores it internally (\eg as an OpenGL buffer), and renders it. Whenever
@@
 Often, objects manipulated by the robot have known meshes with corresponding CAD
-models that can be conveniently pre-loaded. In these case, \uwds provides a
+models that can be conveniently pre-loaded. In these cases, \uwds provides a
 tool, {\tt uwds load}, that loads a mesh into a \uwds network (and optionally,
-creates a node) from a large range of 3D format (including Collada, FBX, OBJ,
+creates a node) from a large range of 3D formats (including Collada, FBX, OBJ,
 Blender)\footnote{The underlying import
@@
 
-\uwds provides a range of tools to inspect a running the network. Graphical
+\uwds provides a range of tools to inspect a running network. Graphical
 tools ({\tt uwds explorer} and {\tt uwds timeline}, see Figure~\ref{fig|explorer})
-provide an user-friendly overview of the system's graph with the connections
+provide a user-friendly overview of the system's graph with the connections
 between the clients and the worlds, as well as their activity.
@@
     \caption{Screenshots of the timeline (top) and network (bottom)
-    introspection tools. Worlds are represented as boxes, clients as ellipses.}
+    introspection tools. Worlds are represented as boxes; clients as ellipses.}
     \label{fig|explorer}
@@
 
-Specialised command-line tools are also available to list at run-time the worlds
-and their content ({\tt uwds ls}) or to display detailed informations on a
+Specialised command-line tools are also available to list the worlds
+and their content ({\tt uwds ls}) at run-time, or to display detailed information for a
 specific node ({\tt uwds show}).
@@
 
-The {\tt uwds tf} client continuously monitor the ROS TF tree, and mirrors
+The {\tt uwds tf} client continuously monitors the ROS TF tree, and mirrors
 TF frames as nodes in the desired \uwds world. A node is first created if none
@@
 Currently, the process is unidirectional: the {\tt uwds tf} client performs TF
-to \uwds updates, but not the opposite.
+to \uwds updates, but not the reverse.
 
@@
 grounding~\cite{Tellex2010}. Other examples in human-robot interaction include Ros
-\etal~\cite{ros2010solving, ros2010which} which has been recently integrated
+\etal~\cite{ros2010solving, ros2010which} which has recently been integrated
 into a full architecture for autonomous human-robot
@@
 picking) enables fast (single pass) computation of the visibility of every
-object in the scene, while providing control regarding how many pixels should be
-actually visible for the object to be considered globally visible. The
+object in the scene, while providing control regarding how many pixels should actually be
+visible for the object to be considered globally visible. The
 command-line tool {\tt uwds visibility} returns the list of visible objects from
@@
 
-When integrated into a filter node, visibility computation allows to easily
-create new worlds representing the estimated perspectives of the different
+When integrated into a filter node, visibility computation allows easy
+creation of new worlds representing the estimated perspectives of the different
 agents. Listing~\ref{lst|humanperspective} provides the complete code of such an
@@
     \includegraphics[width=\linewidth]{l2tor-photo2}
-    \caption{A user manipulating objects on top of a RFID mat (providing object
+    \caption{A user manipulating objects on top of an RFID mat (providing object
     localisation) with a NAO robot, while an operator monitors
-    the interaction on a tablet that display an {\tt uwds view} client.}
+    the interaction on a tablet that displays an {\tt uwds view} client.}
 
@@
 pedagogical strategies to promote spatial relation learning by object
-manipulation. Such instructions include commands like to place the elephant
-\textit{on top} or \textit{next to} the cube or put the giraffe \textit{inside}
+manipulation. Such instructions include commands to place the elephant
+\textit{on top} or \textit{next to} the cube, or put the giraffe \textit{inside}
 the house.  To assess the performance, the robot and the operator need to
-compute spatial relation between objects from the visual perspective of the
+compute spatial relationships between objects from the visual perspective of the
 participant.
@@
 
-The interaction scenario requires a number of \uwds clients. The client are
+The interaction scenario requires a number of \uwds clients. The clients are
 distributed across two computers (the robot and a Microsoft Windows-operated
-tablet) that communicate with each other in order to synchronise the worlds
+tablet) that communicate with each other in order to synchronise the world
 representations. Figure~\ref{fig|l2torarchitecture} illustrates the topology of
@@
 
-\subsection{Relation to existing robotic middlewares}
+\subsection{Relation to existing robotic middleware}
 
-Like traditional robotic middlewares, \uwds offers a form of distributed
+Like traditional robotic middleware, \uwds offers a form of distributed
 computation based on message passing. However, it distinguishes itself from
@@
 First and foremost, \uwds purposefully does not offer any \emph{general}
-capability at distributing computation and data flows amongst independent
-components: it focuses specifically on distributing environment models, both
+capability to distribute computation and data flow amongst independent
+components: it focusses specifically on distributing environment models, both
 spatial (geometric models) and temporal (events and situations).
 
-While using standard middlewares as \emph{underlying transport} for \uwds would be
+While using standard middleware as \emph{underlying transport} for \uwds would be
 technically feasible and relatively easy to implement, it does not offer any
@@
 
-We have presented as well a concrete instantiation of a system relying on \uwds
+We have additionally presented a concrete instantiation of a system relying on \uwds
 for its representation needs, and we have sketched future directions of

### 2017-02-28 paper.tex
Update on Overleaf.
[broadcasted->broadcast]

@@
 
+%CDW - Remove 'Besides'. I think you are trying to establish why uwds is not like previous work but I don't think that word quite works in there. Probably ok with just removing it.
 Work on distributed scene graphs~\cite{naef2003blue} has been previously applied
@@
 % JK: like software, the plural of middleware is middleware :)
+% CDW: I get the analogy with first-class citizens but perhaps better to replace with something like 'but is not a great strength' or 'but is not a highly developed feature'
 Traditional robotic middleware, like ROS, is not particularly well-suited to
@@
 Clients can both read and write onto the worlds they are connected to, and
-automatically see updates broadcasted by other clients connected to the same
+automatically see updates broadcast by other clients connected to the same
 world.
@@
 Listing~\ref{lst|pythonapi}). Modifications to the model are asynchronously synchronised with
-a central server (the {\tt underworlded} daemon) and broadcasted to every other
+a central server (the {\tt underworlded} daemon) and broadcast to every other
 client connected to the same world.
@@
 
-This tool also supports basic object manipulations, that are broadcasted to the
+This tool also supports basic object manipulations, that are broadcast to the
 other \uwds clients connected to the same world.

### 2017-02-28 paper.tex
Cleared other comments from Chris
Added 'underworlds' to title, as per James suggestion
Future work section

@@
 \title{\LARGE \bf
-Cascading Situation Assessment for Robots
+{\sc underworlds}: Cascading Situation Assessment for Robots
 }
@@
 
-%CDW - Remove 'Besides'. I think you are trying to establish why uwds is not like previous work but I don't think that word quite works in there. Probably ok with just removing it.
 Work on distributed scene graphs~\cite{naef2003blue} has been previously applied
@@
 
-% JK: like software, the plural of middleware is middleware :)
-% CDW: I get the analogy with first-class citizens but perhaps better to replace with something like 'but is not a great strength' or 'but is not a highly developed feature'
 Traditional robotic middleware, like ROS, is not particularly well-suited to
@@
 \section{Application Example}
+\label{application}
 
@@
 
-Representation of uncertainty
+As illustrated in section~\ref{application}, \uwds is already deployed and used
+on the field. Several features are however still under development.
 
-Representation of fields
+\subsubsection{Representation capabilities} as presented in
+section~\ref{design}, the current version of \uwds allows to represent
+\emph{objects}, \emph{abstract entities} like groups and \emph{perspectives}.
+\emph{Fields} are also part of the \uwds design, but are not yet implemented.
 
-Client libraries in more languages
+\paragraph*{Representation of fields} fields are commonly used to represent
+continuously-valued spatial entities. Fields might or might not be spatially
+bounded. Examples include the working space of a robot arm (spatially bounded),
+the field of view of a camera (spatially bounded), proxemics (potentially
+unbounded). We plan to represent fields in \uwds using the memory-efficient
+octomaps~\cite{hornung2013octomap}. Similarly to geometric data, octomaps
+will not be directly stored with the nodes (nodes will refer to them through 
+handles), but unlike geometric data, they will not be treated as immutable
+datasets by the server, permitting real-time updates.
+
+\paragraph*{Representation of uncertainty} currently node positions are 
+stored as $4\times4$ transformation matrices, relative to the node parent. This
+representation is efficient, and conveniently matches traditional representation
+systems (including ROS TF frames or OpenGL transformations). However, the explicit management
+of uncertainties is instrumental to many robotic applications, and we plan to
+add full support for position uncertainties to \uwds. We plan to add this
+support by adding a pose covariance matrix to the nodes, and equipping the
+different \uwds helper tools with corresponding support (like covariance
+ellipses visualisation in {\tt uwds view}).
+
+\subsubsection{Manipulation of representations} \uwds makes it possible to
+cheaply clone and modify multiple worlds. However, little is provided to
+jointly manipulate sets of worlds. In that regard, we plan to add a mechanism to
+efficiently \emph{compare} two worlds. This would permit a client
+to quickly identify modifications applied to a world, and could prove especially
+useful when combined with temporal events to answer questions like ``what has
+changed in the scene?''. Comparisons between worlds involves the definition of
+an efficient fuzzy \emph{diff} algorithm for 3D scene graphs, able to
+selectively ignore non-relevant micro displacements.
+
+\subsubsection{Implementation and Integration} finally, we plan to continue to
+improve the integration of \uwds into existing software architectures. A
+short-term goal is to provide excellent C++ support, with a high-level,
+user-friendly C++ client library. This is critical for a broader adoption of \uwds
+within the robot community. Support for other languages might follow, depending
+on demands and open-source contributions.
+
+In term of implementation, we are also investigating alternative back-ends that
+could compete with gRPC to provide the synchronisation mechanisms between
+clients. Due to the high-level nature of \uwds (a set of shared, tree-like data
+structures), the use of the GIT version control system is one particular (if
+unusual) option being studied. The scene graph would be represented as a
+hierarchy of directories, nodes would be files in these directories, creating
+new worlds would consist in creating a branch in the repository, and sharing
+would be supported through the usual pull/push GIT mechanisms. Initial
+prototyping shows that under the right conditions (in particular, an in-memory
+file system), performance is not an issue. Automatic conflicts resolution, on
+the other hand, might prove more challenging.
 
@@
 spatio-temporal representations of a robot's world. The key contributions of our
-approach are: a composite data-structure for environment representation within a
+approach are: a composite data structure for environment representation within a
 robotic software architecture, made of a scene graph and a timeline; a mechanism
-to efficiently and transparently share this data-structure amongst a set of
+to efficiently and transparently share this data structure amongst a set of
 clients (the software modules of the robot); a cascading architecture permitting
 the explicit of representation of alternative states of the world while
-maintaining a network of dependencies amongst themselves; a typology of clients
+maintaining a network of dependencies; a typology of clients
 (\emph{root}, \emph{leaf}, \emph{filter}, \emph{transformer}) to systematically

### 2017-02-28 paper.tex
Conclusion

@@
 
-Integration with effort like RSG-DSL~\cite{blumenthal2014towards}
+We believe this work can practically support existing robotic architectures with
+state-of-the-art spatio-temporal representation capabilities. We also hope that
+this line of research can lead to a better understanding of the representation
+needs of modern robotic systems, and participate to the emergence of a possible
+common representation platform for robots, building on previous formalisation
+efforts like the RSG-DSL domain specific language~\cite{blumenthal2014towards}.
 

### 2017-03-01 paper.tex
Update on Overleaf.

@@
 
-However, {\sc Spark} is a monolithic module, that does not support sharing its
+However, {\sc Spark} is a monolithic module that does not support sharing its
 internal 3D model with other external components. In contrast, \uwds focuses on
@@
 typically needs to reason over hypothetical future states of the environment
-(``what happen if I move this glass onto that pile of books?'').  The planner
-generates many possible future states, that in turn might require further
+(``What happens if I move this glass onto that pile of books?'').  The planner
+generates many possible future states, which in turn might require further
 processing (for instance, running a physics simulation). Such a tool would

### 2017-09-27 paper.tex
Change template to use HRI one (ACM)

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%2345678901234567890123456789012345678901234567890123456789012345678901234567890
-%        1         2         3         4         5         6         7         8
+\documentclass[sigconf]{acmart}
 
-\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
+\usepackage{booktabs} % For formal tables
 
-%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
+% Copyright
+%\setcopyright{none}
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+\setcopyright{rightsretained}
+%\setcopyright{usgov}
+%\setcopyright{usgovmixed}
+%\setcopyright{cagov}
+%\setcopyright{cagovmixed}
 
-\IEEEoverridecommandlockouts                              % This command is only needed if 
-                                                          % you want to use the \thanks command
+% DOI
+\acmDOI{10.475/123_4}
 
-\overrideIEEEmargins                                      % Needed to meet printer requirements.
+% ISBN
+\acmISBN{123-4567-24-567/08/06}
 
-\usepackage[utf8x]{inputenc}
-\usepackage{graphicx}
+%Conference
+\acmConference[HRI2018]{ACM/IEEE Human-Robot Conference}{March 2018}{Chicago USA} 
+\acmYear{2018}
+\copyrightyear{2018}
 
-\usepackage{hyperref}
+\acmArticle{4}
+\acmPrice{15.00}
 
-% See the \addtolength command later in the file to balance the column lengths
-% on the last page of the document
-
-% The following packages can be found on http:\\www.ctan.org
-%\usepackage{graphics} % for pdf, bitmapped graphics files
-%\usepackage{epsfig} % for postscript graphics files
-%\usepackage{mathptmx} % assumes new font selection scheme installed
-%\usepackage{times} % assumes new font selection scheme installed
-%\usepackage{amsmath} % assumes amsmath package installed
-%\usepackage{amssymb}  % assumes amsmath package installed
-
-\title{\LARGE \bf
-{\sc underworlds}: Cascading Situation Assessment for Robots
+\title{{\sc underworlds}: Cascading Situation Assessment for Robots
 }
 
-\author{Séverin Lemaignan$^{*}$, Fotios Papadopoulos$^{*}$%
-\thanks{$^{*}$Authors are with Centre for Robotics and Neural Systems,
-        Plymouth University, Plymouth PL48AA, United Kingdom
-        {\tt\small firstname.surname@plymouth.ac.uk}}%
-}
+%\author{Séverin Lemaignan$^{*}$, Fotios Papadopoulos$^{*}$%
+%\thanks{$^{*}$Authors are with Centre for Robotics and Neural Systems,
+%        Plymouth University, Plymouth PL48AA, United Kingdom
+%        {\tt\small firstname.surname@plymouth.ac.uk}}%
+%}
+
+\author{[Hidden for review]}
+\orcid{1234-5678-9012}
+%\affiliation{%
+%  \institution{xxx}
+%  \streetaddress{xxx}
+%  \city{Dublin} 
+%  \state{Ohio} 
+%  \postcode{43017-6221}
+%}
+%\email{trovato@corporation.com}
+
+%
+% The code below should be generated by the tool at
+% http://dl.acm.org/ccs.cfm
+% Please copy and paste the code instead of the example below. 
+%
+
+ \begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10010520.10010553.10010554</concept_id>
+<concept_desc>Computer systems organization~Robotics</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Computer systems organization~Robotics}
+
+\keywords{Human-Robot Interaction, Situation Assessment, Perspective Taking}
 
@@
 
-\maketitle
-\thispagestyle{empty}
-\pagestyle{empty}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{abstract}
+
+\maketitle
+%\thispagestyle{empty}
+%\pagestyle{empty}
 
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{bibliography}

### 2017-09-27 paper.tex
Minor polish

@@
     set of \emph{worlds} that can each have different spatial and temporal
-    granularities, and may depend on each other.  \uwds also provides a set of
+    granularities, and may inherit from each other.  \uwds also provides a set of
     high-level client libraries and tools to introspect and manipulate the
@@
 
-Traditional robotic middleware, like ROS, is not particularly well-suited to
+Traditional robotic middlewares, like ROS, are not particularly well-suited to
 deal with these different needs: full geometric data can be represented, but are
-not first-class citizens (a basic task like displaying a 3D mesh at a dynamic
+not first-class citizens (a basic task like displaying a 3D mesh at an arbitrary
 position is not particularly easy to perform with ROS) and critically,
@@
 model is represented as a scene graph. The scene graph has a unique root node,
-in which a tree of other nodes is parented.
+to which a tree of other nodes is parented.
 
@@
 like point clouds) are \emph{not} stored within the nodes themselves. \uwds
-represents geometric data as immutable data, labelled with their hash
+represents geometric data as immutable data, identified by their hash value
 (preventing \textit{de facto} data duplication).  Nodes only store the hash
 corresponding to the desired geometric data, and the actual data is pulled from
-the server by the clients whenever they actually need it (rendering for
+the server by the clients whenever they actually need it (for rendering for
 instance).
@@
     \caption{Screenshot of the {\tt uwds view} 3D visualisation and manipulation
-    tool. In this particular example, the 3D meshes have been pre-loaded using
+    client. In this particular example, the 3D meshes have been pre-loaded using
     {\tt uwds load}. Their positions are then updated at run-time using the
@@
 interface definition language), and bindings to various languages and platforms
-can be automatically generated from the protocol definition file (as of Feb
+can be automatically generated from the protocol definition file (as of Sept
 2017, gRPC can generate bindings for C, C++, C\#, Node.js, PHP, Ruby, Python, Go
-and Java, on Windows, Mac and Linux).  The cross-platform/cross-language support
+and Java, on Windows, Mac, Linux and Android).  The cross-platform/cross-language support
 of gRPC is especially welcome in the academic context, as it offers ease and
@@
 
-This tool also supports basic object manipulations, that are broadcast to the
-other \uwds clients connected to the same world.
+This tool also supports basic object manipulations (translations, rotations),
+that are broadcast to the other \uwds clients connected to the same world.
 
@@
 picking) enables fast (single pass) computation of the visibility of every
-object in the scene, while providing control regarding how many pixels should actually be
+object in the scene, while providing control regarding how many pixels should be actually
 visible for the object to be considered globally visible. The

### 2017-09-27 paper.tex
Anonymization

@@
 \uwds is a distributed and lightweight open-source
-framework\footnote{\url{https://github.com/severin-lemaignan/underworlds}} that
+%framework \footnote{\url{https://github.com/severin-lemaignan/underworlds}} that
+framework \footnote{[GitHub URL hidden for blind review]} that
 enables robot programmers to build and refine spatial and temporal
@@
     \centering
-    \includegraphics[width=\linewidth]{l2tor-photo2}
+    \includegraphics[width=\linewidth]{l2tor-photo2-blind}
     \caption{A user manipulating objects on top of an RFID mat (providing object
     localisation) with a NAO robot, while an operator monitors
-    the interaction on a tablet that displays an {\tt uwds view} client.}
+    the interaction on a tablet that displays an {\tt uwds view} client [photo blurred for blind review].}
 
@@
 
-\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor}
+%\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor}
+\uwds has been used within the [hidden for blind review] project
 in order to conceptualise and visualise the spatial relations and visibility of
@@
     \includegraphics[width=0.9\linewidth]{l2tor}
-    \caption{Schema of the \uwds architecture used in the L2TOR project. }
+    %\caption{Schema of the \uwds architecture used in the L2TOR project. }
+    \caption{Schema of the \uwds architecture used in the [hidden for blind review] project. }
     \label{fig|l2torarchitecture}
@@
 
-The L2TOR project features an additional scenario where the user performs the
+%The L2TOR project features an additional scenario where the user performs the
+The [hidden for blind review] project features an additional scenario where the user performs the
 manipulations on a tablet computer instead of using physical objects. To
@@
 
-This work has been supported by the EU H2020 Marie Sk\l odowska-Curie Actions
-project DoRoThy (grant 657227) and the EU H2020 L2TOR project (grant 688014).
+%This work has been supported by the EU H2020 Marie Sk\l odowska-Curie Actions
+%project DoRoThy (grant 657227) and the EU H2020 L2TOR project (grant 688014).
 
-The authors also acknowledge the contribution of the HRI lab of LAAS-CNRS (where this
-work debuted), and in particular, the influential ideas of Rachid Alami.
+%The authors also acknowledge the contribution of the HRI lab of LAAS-CNRS (where this
+%work debuted), and in particular, the influential ideas of Rachid Alami.
+
+[Hidden for blind review]
 

### 2018-01-15 paper.tex
IEEEtrans template (for RSS)

@@
-\documentclass[sigconf]{acmart}
+\documentclass[conference]{IEEEtran}
+\usepackage{times}
 
+% numbers option provides compact numerical references in the text. 
+\usepackage[numbers]{natbib}
+
+\usepackage{graphicx}
 \usepackage{booktabs} % For formal tables
 
-% Copyright
-%\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
+\usepackage{multicol}
+\usepackage[bookmarks=true]{hyperref}
 
-% DOI
-\acmDOI{10.475/123_4}
-
-% ISBN
-\acmISBN{123-4567-24-567/08/06}
-
-%Conference
-\acmConference[HRI2018]{ACM/IEEE Human-Robot Conference}{March 2018}{Chicago USA} 
-\acmYear{2018}
-\copyrightyear{2018}
-
-\acmArticle{4}
-\acmPrice{15.00}
-
-\title{{\sc underworlds}: Cascading Situation Assessment for Robots
+\pdfinfo{
+   /Author (Hidden for blind review)
+   /Title  (underwords: Cascading Situation Assessment for Robots)
+   /CreationDate (D:20180115162200)
+   /Subject (Robots)
+   /Keywords (Situation Assessment;Perspective Taking)
 }
-
-%\author{Séverin Lemaignan$^{*}$, Fotios Papadopoulos$^{*}$%
-%\thanks{$^{*}$Authors are with Centre for Robotics and Neural Systems,
-%        Plymouth University, Plymouth PL48AA, United Kingdom
-%        {\tt\small firstname.surname@plymouth.ac.uk}}%
-%}
-
-\author{[Hidden for review]}
-\orcid{1234-5678-9012}
-%\affiliation{%
-%  \institution{xxx}
-%  \streetaddress{xxx}
-%  \city{Dublin} 
-%  \state{Ohio} 
-%  \postcode{43017-6221}
-%}
-%\email{trovato@corporation.com}
-
-%
-% The code below should be generated by the tool at
-% http://dl.acm.org/ccs.cfm
-% Please copy and paste the code instead of the example below. 
-%
-
- \begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10010520.10010553.10010554</concept_id>
-<concept_desc>Computer systems organization~Robotics</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Computer systems organization~Robotics}
-
-\keywords{Human-Robot Interaction, Situation Assessment, Perspective Taking}
 
@@
 
+\title{{\sc underworlds}: Cascading Situation Assessment for Robots
+}
+
+\author{Author Names Omitted for Anonymous Review. Paper-ID [add your ID here]}
+
+%\author{\authorblockN{Séverin Lemaignan}
+%\authorblockA{Centre for Robotics and Neural Systems\\
+%University of Plymouth\\
+%Plymouth, United Kingdom\\
+%Email: severin.lemaignan@plymouth.ac.uk}
+%\and
+%\authorblockN{Yoan Sallami, Aurélie Clodic and Rachid Alami}
+%\authorblockA{Twentieth Century Fox\\
+%Springfield, USA\\
+%Email: homer@thesimpsons.com}
+%\and
+%\authorblockN{James Kirk\\ and Montgomery Scott}
+%\authorblockA{Starfleet Academy\\
+%San Francisco, California 96678-2391\\
+%Telephone: (800) 555--1212\\
+%Fax: (888) 555--1212}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\maketitle
+
 \begin{abstract}
@@
 
-\maketitle
-%\thispagestyle{empty}
-%\pagestyle{empty}
+\IEEEpeerreviewmaketitle
 
@@
 
-\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{plainnat}
 \bibliography{bibliography}

### 2018-01-15 paper.tex
Expand why ROS not suitable for geometric data

@@
 Traditional robotic middlewares, like ROS, are not particularly well-suited to
-deal with these different needs: full geometric data can be represented, but are
-not first-class citizens (a basic task like displaying a 3D mesh at an arbitrary
-position is not particularly easy to perform with ROS) and critically,
+deal with these different needs: full geometric data can be represented, but is
+not first-class citizen (a basic task like displaying a 3D mesh at an arbitrary
+position is not particularly easy to perform with ROS, requiring the combination
+of static Collada meshes, a URDF kinematic description, TF broadcasters, and a
+3D visualisation tool like RViz) and critically,
 representing and reasoning on alternative states of the environment is not

### 2018-01-15 paper.tex
WIP on formal decription of uwds

@@
 \usepackage{graphicx}
+\usepackage{tikz}
+\usetikzlibrary{arrows}
 \usepackage{booktabs} % For formal tables
@@
 \newcommand{\uwds}{{\sc underworlds}\xspace}
+
+\usepackage{amssymb}
+\usepackage{amsmath}
+\DeclareMathOperator{\transform}{\odot}
+\DeclareMathOperator{\filter}{\ominus}
+\DeclareMathOperator{\W}{\mathcal{W}}
 
@@
 \label{design}
+
+\subsection{Formal description}
+
+An \uwds \emph{world} $\W^k$ consists in a pair $\{\mathcal{S}^k,
+\mathcal{T}^k\}$.
+
+$\mathcal{S}^k=\{N^k, T^k\}$ is a tree representing a spatial
+scene (\emph{scene graph}), with $N^k$ the set of nodes and $T^k$ the set of
+transformations between the nodes. $|T^k| = |N^k| - 1$ and $n^k_0 \in N^k$ is
+called the root node of the scene $\mathcal{S}^k$.
+
+$\mathcal{T}^k=\{Evt^k, Sit^k\}$ is the world's \emph{timeline}, consisting in a
+set $Evt^k$ of events (duration-less states) and a set $Sit^k$ of sitations
+(states with a start time and an option end time).
+
+\uwds \emph{transformations} $\transform$ are operators that transform one world into
+another. They are often unary operator (for instance, $\mathcal{W}^{k'} =
+\transform\mathcal{W}^k$), but do not have to. \emph{Filters} (noted $\filter$)
+are a common type of unary, antiextensive transformations such as
+$\filter\mathcal{W}^k \subseteq \mathcal{W}^k$.
+
+An \uwds \emph{network} $\mathcal{N} = \{W, O\}$ is a directed graph (cycles are
+permitted) of worlds $W=\{\mathcal{W}^k\}, k\in0\ldots n$ and operators
+$O=\{\transform^l\}, l \in 0\ldots m$.
+
+\begin{tikzpicture}[->,>=stealth',level/.style={sibling distance = 5cm/#1,
+    level distance = 1.5cm}] 
+    \node {$\mathcal{W}^0$ {\tt\footnotesize base}}
+    child{ node  {$\W^1$ {\tt\footnotesize corrected}} edge from parent node[above left] {$\filter$}
+            child{ node {$\W^2$ {\tt\footnotesize semantic\_map}} 
+                child{ node {}} %for a named pointer
+                            child{ node {}}
+            }
+            child{ node  {$\W^3$ {\tt\footnotesize post\_planning}}
+                            child{ node {}}
+                            child{ node {}}
+            }                            
+    }
+    ; 
+\end{tikzpicture}
 

### 2018-01-24 paper.tex
Slightly better example of uwds client

@@
 listens for changes in an input world, removes some objects (in this case, items
-whose name contain the string ``{\tt wall}''), and forwards all other changes to
+whose volume is below a threshold), and forwards all other changes to
 an output world, effectively making the output world a copy of the input world
-with all walls removed.
+with all smaller objects removed.
 
@@
 # by default, connect to the server on localhost
-with underworlds.Context("walls_filter") as ctx:
+with underworlds.Context("small_object_filter") as ctx:
 
@@
         for node in in_world.scene.nodes:
-            if("wall" not in node.name):
+            if node.volume > THRESHOLD:
                 out_world.scene.nodes.update(node)
@@
     \caption{Example of a simple yet complete \uwds filter, written in Python:
-    the client connects to the \uwds network, walks through the scene graph of
-    world {\tt example\_in}, filters out some objects, and publishes the
-    remaining objects in the world {\tt example\_out}.}
+    the client connects to the \uwds network, blocks until the world {\tt
+    example\_in} changes, and only propagate nodes that match the condition to
+    the world {\tt example\_out}.}
 

### 2018-01-24 paper.tex
Added performance graph

@@
 
-\section{Application Example}
+\section{Performances and Application Example}
 \label{application}
+
+\subsection{Performances}
+\label{performances}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{performances}
+    \caption{Propagation times of one change (node creation)
+    across $n$ worlds. The test is performed by running $n-1$ pass-through
+    filters (independent processes) which each monitor one world and replicate
+    any change into the next world. Times measured over 20 trials, performed on
+    a 8 core machine.}
+    \label{fig|performances}
+\end{figure}
+
+The main performance metric of \uwds is the \emph{propagation time}: the time is
+take of a change in a given world to cascade to all the dependant
+
+\subsection{Application to service robots and human-robot interaction}
 
@@
 
-\subsection*{Architecture}
+\subsection{Architecture}
 

### 2018-01-24 paper.tex
Cleaned up section to be rewritten by LAAS

@@
     \includegraphics[width=\linewidth]{l2tor-photo2-blind}
-    \caption{A user manipulating objects on top of an RFID mat (providing object
-    localisation) with a NAO robot, while an operator monitors
-    the interaction on a tablet that displays an {\tt uwds view} client [photo blurred for blind review].}
-
-    \label{fig|l2torexample}
+    \caption{TODO LAAS: update photo -> MuMMER picture?}
+    \label{fig|mummer-example}
 \end{figure}
 
-%\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor}
+%\uwds has been used within the European MuMMER project
 \uwds has been used within the [hidden for blind review] project
 in order to conceptualise and visualise the spatial relations and visibility of
-the physical objects that participants interact with. In one of the scenarios of this
-project, participants are instructed by a NAO robot to manipulate a number of
-Duplo-like animals on top of a printed background paper which serves as a
-landscape (Figure~\ref{fig|l2torexample}). The robot instructions are based on
-pedagogical strategies to promote spatial relation learning by object
-manipulation. Such instructions include commands to place the elephant
-\textit{on top} or \textit{next to} the cube, or put the giraffe \textit{inside}
-the house.  To assess the performance, the robot and the operator need to
-compute spatial relationships between objects from the visual perspective of the
-participant.
+the physical objects that participants interact with.
 
-\subsection{Architecture}
+TODO LAAS
 
@@
     \includegraphics[width=0.9\linewidth]{l2tor}
-    %\caption{Schema of the \uwds architecture used in the L2TOR project. }
-    \caption{Schema of the \uwds architecture used in the [hidden for blind review] project. }
-    \label{fig|l2torarchitecture}
+    %\caption{Schema of the \uwds architecture used in the MuMMER project. }
+    \caption{TODO LAAS update archi -- Schema of the \uwds architecture used in the [hidden for blind review] project. }
+    \label{fig|mummerarchitecture}
 \end{figure}
@@
 The interaction scenario requires a number of \uwds clients. The clients are
-distributed across two computers (the robot and a Microsoft Windows-operated
-tablet) that communicate with each other in order to synchronise the world
-representations. Figure~\ref{fig|l2torarchitecture} illustrates the topology of
-the \uwds network, that consists in the following clients:
+distributed across...
 
-\begin{itemize}
-    \item \textbf{Kinect Head Tracker}: captures the location and the direction
-        of the user's head -- maintains the {\tt human} \emph{Perspective} node;
+Figure~\ref{fig|mummerarchitecture} illustrates the topology of
+the \uwds network...
 
-    \item \textbf{Object RFID tracker}: updates the locations of the virtual
-        objects based on RFID data;
-
-    \item \textbf{NAO joint state publisher}: virtual representation of the
-        robot -- maintains the {\tt robot} \emph{Perspective} node;
-
-    \item \textbf{Visibility computation}: computes visibility based on
-        \emph{Perspective} nodes;
-
-    \item \textbf{Spatial relations computation}: extracts symbolic spatial relations
-        between the virtual objects;
-
-    \item \textbf{Game/Rules computation}: interaction manager based on \uwds
-        clients' input.
-
-\end{itemize}
-
-%The L2TOR project features an additional scenario where the user performs the
-The [hidden for blind review] project features an additional scenario where the user performs the
-manipulations on a tablet computer instead of using physical objects. To
-accommodate this change, we replaced the object detection \uwds client with a 3D
-virtual object detection while the rest of the system remains the same.
-
-Finally, we are evaluating the use of a hybrid platform that utilises both virtual and
-physical objects at the same time.
+TODO LAAS
 

### 2018-01-24 paper.tex
Minor rewrite of the 'difference with middlewares' section

@@
 computation based on message passing. However, it distinguishes itself from
-existing middleware in several significant ways.
-
-First and foremost, \uwds purposefully does not offer any \emph{general}
-capability to distribute computation and data flow amongst independent
-components: it focusses specifically on distributing environment models, both
-spatial (geometric models) and temporal (events and situations).
+existing middleware in significant ways. Most importantly, \uwds purposefully
+does not offer any \emph{general} capability to distribute computation and data
+flow amongst independent components: it focusses specifically on distributing
+environment models, both spatial (geometric models) and temporal (events and
+situations). In that sense, \uwds really is a \emph{distributed datastructure}
+with a narrow focus on the needs of spatio-temporal representations.
 

### 2018-01-24 paper.tex
Updated the formal description section

@@
 \usepackage{tikz}
-\usetikzlibrary{arrows}
+\usetikzlibrary{graphs,graphdrawing,arrows}
+\usegdlibrary{trees, layered}
 \usepackage{booktabs} % For formal tables
-
 \usepackage{multicol}
@@
 \usepackage{amsmath}
-\DeclareMathOperator{\transform}{\odot}
-\DeclareMathOperator{\filter}{\ominus}
+\DeclareMathOperator{\op}{\odot}
+\DeclareMathOperator{\filter}{\otimes}
+\DeclareMathOperator{\generator}{\oplus}
+\DeclareMathOperator{\consumer}{\ominus}
 \DeclareMathOperator{\W}{\mathcal{W}}
+\DeclareMathOperator{\Ws}{\boldsymbol{W}}
+\DeclareMathOperator{\Os}{\boldsymbol{O}}
 
@@
 
-An \uwds \emph{world} $\W^k$ consists in a pair $\{\mathcal{S}^k,
+An \uwds \emph{world} $\W^k \in \Ws$ consists in a pair (\emph{scene}, \emph{timeline}),
+noted $\{\mathcal{S}^k,
 \mathcal{T}^k\}$.
 
-$\mathcal{S}^k=\{N^k, T^k\}$ is a tree representing a spatial
-scene (\emph{scene graph}), with $N^k$ the set of nodes and $T^k$ the set of
-transformations between the nodes. $|T^k| = |N^k| - 1$ and $n^k_0 \in N^k$ is
-called the root node of the scene $\mathcal{S}^k$.
+$\mathcal{S}^k=\{N^k, T^k\}$ is a tree representing a spatial scene (\emph{scene
+graph}), with $N^k$ the set of nodes and $T^k$ the set of 6D geometric
+transformations between the nodes ($|T^k| = |N^k| - 1$). $n^k_0 \in N^k$ is
+called the \emph{root node} of the scene $\mathcal{S}^k$.
 
 $\mathcal{T}^k=\{Evt^k, Sit^k\}$ is the world's \emph{timeline}, consisting in a
-set $Evt^k$ of events (duration-less states) and a set $Sit^k$ of sitations
+set of events $Evt^k$ (duration-less states) and a set of situations $Sit^k$
 (states with a start time and an option end time).
 
-\uwds \emph{transformations} $\transform$ are operators that transform one world into
-another. They are often unary operator (for instance, $\mathcal{W}^{k'} =
-\transform\mathcal{W}^k$), but do not have to. \emph{Filters} (noted $\filter$)
-are a common type of unary, antiextensive transformations such as
-$\filter\mathcal{W}^k \subseteq \mathcal{W}^k$.
+\uwds \emph{operators} (noted $\op\colon \Ws^{0\ldots m} \rightarrow
+\Ws^{0\ldots n}$) transform one world into another. Examples of operators
+include \emph{filters} (noted $\filter$), a common type of unary, antiextensive
+transformations such as $\filter\W^k \subseteq \W^k$, and \emph{generators}
+(noted $\generator$), a class of nullary operators that are used to create new
+worlds (for instance, from perception routines).
 
-An \uwds \emph{network} $\mathcal{N} = \{W, O\}$ is a directed graph (cycles are
-permitted) of worlds $W=\{\mathcal{W}^k\}, k\in0\ldots n$ and operators
-$O=\{\transform^l\}, l \in 0\ldots m$.
+Finally, an \uwds \emph{network} $\mathcal{N} = \{\Ws, \Os\}$ is a directed graph (cycles are
+permitted) of worlds $\Ws=\{\W^k\}, k\in0\ldots n$ and operators
+$\Os=\{\op^l\}, l \in 0\ldots m$. As described in the following sections,
+operators are typically implemented as independent software processes that connect at
+runtime to their operands, \ie the \uwds worlds, to read and write them. These
+processes are hereafter called \emph{clients} of the \uwds network.
 
-\begin{tikzpicture}[->,>=stealth',level/.style={sibling distance = 5cm/#1,
-    level distance = 1.5cm}] 
-    \node {$\mathcal{W}^0$ {\tt\footnotesize base}}
-    child{ node  {$\W^1$ {\tt\footnotesize corrected}} edge from parent node[above left] {$\filter$}
-            child{ node {$\W^2$ {\tt\footnotesize semantic\_map}} 
-                child{ node {}} %for a named pointer
-                            child{ node {}}
-            }
-            child{ node  {$\W^3$ {\tt\footnotesize post\_planning}}
-                            child{ node {}}
-                            child{ node {}}
-            }                            
-    }
-    ; 
-\end{tikzpicture}
+%As an example, Figure~\ref{fig|formal-net} shows the
+%formal model of the \uwds network represented in Figure~\ref{fig|scene}.
+%
+%\begin{figure}[!h]
+%    \centering
+%    \begin{tikzpicture}[nodes={text height=0.4cm,align=center,text width=1.5cm,
+%                    draw=black!20, thick, fill=white, font=\footnotesize},
+%                        tr/.style={draw=none,text width=0.3cm},
+%                        >=stealth', rounded corners, semithick]
+%\graph [layered layout, grow=right, level distance=0.5cm, sibling sep=.5em, sibling distance=1cm] {
+%    "$\W^0$ {\tiny\tt base}" -- "$\filter^0$"[tr] -> "$\W^1$ {\tiny\tt corrected}";
+%        "$\W^1$ {\tiny\tt corrected}" -- "$\op^2$"[tr] -> "$\W^3$ {\tiny\tt post\_planning}";
+%        "$\W^1$ {\tiny\tt corrected}" -- "$\op^1$"[tr] -> "$\W^2$ {\tiny\tt semantic\_mapping}";
+%};
+%\end{tikzpicture}
+%\caption{}
+%    \label{fig|formal-net}
+%\end{figure}
+%
 

### 2018-01-24 paper.tex
Several smaller changes

@@
 
-Traditional robotic middlewares, like ROS, are not particularly well-suited to
+Traditional robotic middlewares, like ROS, are not particularly well suited to
 deal with these different needs: full geometric data can be represented, but is
@@
 approach to \textbf{represent and share multiple parallel representations of the
-world}. \uwds does so by permitting clients to clone existing worlds, modify
+world}. \uwds does so by allowing clients to clone existing worlds, modify
 them, and re-share them, without the cost of duplicating geometric data (as
@@
 
-\subsection{Cascading architecture}
+\subsection{Software architecture}
 
-Figure~\ref{fig|scene} pictures a typical \uwds topology: a graph (that
-happens to be a directed acyclic graph on Figure~\ref{fig|scene}, but does not
-have to be in the general case -- cycles are permitted) of \emph{clients}
-connected through shared data structures called \emph{worlds}.
+Figure~\ref{fig|scene} pictures a typical \uwds topology: a graph (that happens
+to be an \emph{acyclic} graph on Figure~\ref{fig|scene}, but does not have to
+be in the general case) of worlds and operators connecting the worlds together.
+In the following, we however tend to adopt a software engineering perspective,
+in which the \uwds network is made of \emph{clients} (which implement the
+operators) connected through shared data structures called \emph{worlds}.
+The formal model and this engineering-oriented model are isomorphic.
 
@@
 
-Software components accessing \uwds worlds are called clients. Some standard
-clients (like a 3D visualisation tool) are provided with the \uwds package (see
-section~\ref{std_clients}). Clients are otherwise written by the end users
-using the \uwds client API (see section~\ref{api}).
+Software components implementing \uwds operators and accessing worlds are
+called clients. Some standard clients (like a 3D visualisation tool) are
+provided with the \uwds package (see section~\ref{std_clients}). Clients are
+otherwise written by the end users using the \uwds client API (see
+section~\ref{api}).
 
@@
 automatically see updates broadcast by other clients connected to the same
-world.
+world. To ensure data consistency, worlds can have many simultaneous readers,
+but only one writer at a given time.
 
-Four specific types of clients can be distinguished: \textbf{root clients} which
-create and update worlds (`write-only' client, like the \emph{Sensor fusion} and
-\emph{ROS interface} clients in Figure~\ref{fig|scene}); \textbf{leaf clients}
-which on the contrary only read worlds, without modifying them (like the
-\emph{Computation of spatial relations} client on Figure~\ref{fig|scene});
-\textbf{filters} that copy an input world into an output world, performing
-some filtering operation in-between (like the client \textit{Physics-based
-position correction}); and \textbf{transformers} which transform one
-representation into another (like the \emph{Semantic mapping} client on
-Figure~\ref{fig|scene}).
+Three specific types of clients can be distinguished: \textbf{generator
+clients} which implement generator operators, and create and update worlds
+(`write-only' client, like the \emph{Sensor fusion} and \emph{ROS interface}
+clients in Figure~\ref{fig|scene}); \textbf{leaf clients} which on the contrary
+only read worlds, without modifying them (like the \emph{Computation of spatial
+relations} client on Figure~\ref{fig|scene}); and \textbf{transformers} which
+transform one representation into another (like the \emph{Semantic mapping}
+client on Figure~\ref{fig|scene}). \textbf{Filters} are a specfic class of
+transformer clients that copy an input world into an output world, performing some
+filtering operation in-between (like the client \textit{Physics-based position
+correction}).
 
@@
 
-Each world is technically independent from all others. Dependencies between
-worlds arise from the clients' connections. For instance, filters effectively
-create a dependency between worlds. On Figure~\ref{fig|scene}, the \textit{Physics-based
-position correction} client creates a dependency between the world {\tt base} (which
-represents here the result of raw sensor fusion) and the world {\tt corrected}
-which would be a physically-consistent copy of {\tt base}.
-As a result, an \uwds network can be seen as a dependency graph between worlds (where
-cyclic dependencies are permissible).
+While each world is technically independent from all the others, dependencies
+(and therefore, coupling) arise between worlds from the clients' connections.
+For instance, filters effectively create a dependency between worlds. On
+Figure~\ref{fig|scene}, the \textit{Physics-based position correction} client
+creates a dependency between the world {\tt base} (which represents here the
+result of raw sensor fusion) and the world {\tt corrected} which would be a
+physically-consistent copy of {\tt base}.  As a result, an \uwds network can
+also be seen as a dependency graph between worlds (where cyclic dependencies
+are permissible).
 
 This architecture enables what we call \emph{cascading situation assessment}:
-independent software components (the clients) build, refine and share successive
-models of the environment by a combination of filtering/transformations steps
-and model branching.
+independent software components (the clients) build, refine and share
+successive models of the environment by a combination of
+filtering/transformations steps and model branching. A change performed by one
+client (for instance, a face tracker updates the pose of the human head) may
+thereby \emph{cascade} to each of the downstream, dependent worlds.
 

### 2018-01-25 paper.tex
minor

@@
 client connected to the same world.
-% JK: underworlded - intentional? (I guess yes, but checking)
 
@@
 interface definition language), and bindings to various languages and platforms
-can be automatically generated from the protocol definition file (as of Sept
-2017, gRPC can generate bindings for C, C++, C\#, Node.js, PHP, Ruby, Python, Go
+can be automatically generated from the protocol definition file (as of Jan
+2018, gRPC can generate bindings for C, C++, C\#, Node.js, PHP, Ruby, Python, Go
 and Java, on Windows, Mac, Linux and Android).  The cross-platform/cross-language support

### 2018-01-29 paper.tex
Mkae lualatex happy
Minor language

@@
 \usepackage[bookmarks=true]{hyperref}
-
-\pdfinfo{
-   /Author (Hidden for blind review)
-   /Title  (underwords: Cascading Situation Assessment for Robots)
-   /CreationDate (D:20180115162200)
-   /Subject (Robots)
-   /Keywords (Situation Assessment;Perspective Taking)
-}
 
@@
     across $n$ worlds. The test is performed by running $n-1$ pass-through
-    filters (independent processes) which each monitor one world and replicate
-    any change into the next world. Times measured over 20 trials, performed on
+    filters that monitor one world and replicate
+    any changes into the next world. Durations measured over 20 runs, performed on
     a 8 core machine.}

### 2018-01-29 paper.tex
[minor] language

@@
 little eye', as implemented in~\cite{ros2010which}). Perspective taking is a
-cognitive skill that relies on the ability for an agent to project itself in
-others' `eyes' to estimate what they see from their perspectives. 
+cognitive skill that relies on the ability for an agent to take someone else's
+point of view to estimate what they see from their perspectives. 
 Perspective taking has previously been implemented in robotics by temporarily placing
@@
 if A tells B to ``catch the red balloon behind you'', B might create a
-representation of an imaginary red balloon, placed behind him, even without
+representation of an imaginary red balloon, placed behind her, even without
 actually observing the balloon: B \emph{accommodates} the \emph{pre-supposition}
-of a red balloon being present behind himself. Endowing robots with this
-capability has been touched upon by Mavridis \etal within their multi-modal
+of a red balloon being present behind herself. Endowing robots with this
+capability has been touched upon by Mavridis~\etal within their multi-modal
 \emph{Grounded Situation Model}~\cite{Mavridis2006}. However, to the best of our
@@
 
-Figure~\ref{fig|scene} pictures a typical \uwds topology: a graph (that happens
+Figure~\ref{fig|scene} depicts a typical \uwds topology: a graph (that happens
 to be an \emph{acyclic} graph on Figure~\ref{fig|scene}, but does not have to

### 2018-01-29 paper.tex
Added short section on complexity. To be revisited

@@
 
+\subsection{Time and space complexity}
+
+With $n$ the number of worlds and $m$ the number of operators in a network
+$\mathcal{N}$, the worst-case (when every world is a parameter of every
+operator) time complexity of creating or updating a node and propagating the
+change across the network is $O(n \times m)$. This effectively corresponds to
+the server performing $n \times m$ requests.  The space complexity is the same
+(as clients own a copy of the worlds they monitor), except for mesh data whose
+complexity is $O(1)$ (only the server stores the mesh data).
+
 \section{API \& Clients}

### 2018-01-29 paper.tex
More on complexity analysis

--- 2018-01-29 paper.tex
@@
 
+\subsection{Time and space complexity analysis}
+
+\uwds is fundamentally about distributing two datastructures: a scene graph
+(with nodes representing spatial entities) and a timeline (where events are
+stored as a flat list). Typical time and space complexities arise from these
+datastructures. In typical usage scenarios (where the number of nodes or events
+remain relatively small), the computational load to manipulate these
+datastructures is however dominated by the actual processings performed by the
+clients with the data.
+
+More interesting is the time complexity of distributing changes across an \uwds
+network. With $n$ the number of worlds and $m$ the number of operators in a
+network $\mathcal{N}$, the worst-case (when every world is a parameter of every
+operator) time complexity of creating or updating a node and propagating the
+change across the network is $O(n \times m)$ (this effectively corresponds to
+the \uwds server performing $n \times m$ requests to notify clients of the
+update).  The space complexity is the same (as clients own a full copy of the
+worlds they monitor), except for mesh data whose space and time complexities are
+$O(1)$ (only the server stores the mesh data).
+
+In the common case of one client performing a full update of a single world
+(with $p$ nodes) at each time step, the complexity of propagating these changes across
+the network would be $O(p \times m)$.
+
 \section{API \& Clients}

### 2018-01-29 paper.tex
WIP update related work

@@
 
-\uwds is specifically inspired by the geometric and temporal reasoner
-{\sc Spark} (\emph{SPAtial Reasoning \& Knowledge})~\cite{sisbot2011situation}.
-{\sc Spark} acts as a situation assessment reasoner that generates symbolic
-knowledge from the geometry of the environment with respect to relations between
-objects, robots and humans. It also takes into account the different perspective
-that each agent has on the environment. {\sc Spark} embeds a
-modality-independent geometric model of the environment that serves both as
-basis for the fusion of the perception modalities and as bridge with the
-symbolic layer~\cite{lemaignan2016artificial}. This geometric model is built
-from 3D CAD models of the objects, furniture and robots, and full body, rigged
-models of humans.  It is updated at run-time by the robot's sensors.
-Likewise, \uwds embeds a grounded amodal model of the environment, updated
-online from the robot's sensors (sensor fusion).
+\uwds is specifically inspired geometric and temporal reasoner like {\sc Spark}
+(\emph{SPAtial Reasoning \& Knowledge})~\cite{sisbot2011situation}.  {\sc Spark}
+acts as a situation assessment reasoner that generates symbolic knowledge from
+the geometry of the environment with respect to relations between objects,
+robots and humans. It also takes into account the different perspective that
+each agent has on the environment. {\sc Spark} embeds a modality-independent
+geometric model of the environment that serves both as basis for the fusion of
+the perception modalities and as bridge with the symbolic
+layer~\cite{lemaignan2016artificial}. This geometric model is built from 3D CAD
+models of the objects, furniture and robots, and full body, rigged models of
+humans. It is updated at run-time by the robot's sensors.  Likewise, \uwds
+embeds a grounded amodal model of the environment, updated online from the
+robot's sensors (sensor fusion).
 
@@
 offering a shared and distributed representation of the environment within the
-robot's software architecture.
+robot's software architecture. This also distinguishes \uwds from complex
+cognitive toolkits like KnowRob (as found in OpenEASE~\ref{beetz2015open}).
+While these tools maintain a spatio-temporal model of the world, this model is
+internal and not meant to be made widely accessible to other external processes.
+\uwds focuses instead on reusability and sharing of distributed spatio-temporal
+models.  As such, \uwds can be seen as a middleware for spatio-temporal world
+models and, contrary to KnowRob, it does \emph{not} provide any high-level
+processing or reasoning capability (this is left to the users).
 
@@
 providing the ability to create, manipulate and share \emph{multiple alternative
-worlds}. As an example, these could correspond to filtered or hypothetical \emph{views} on
-the initial, perceived model of the environment.
+worlds}. As an example, these could correspond to filtered or hypothetical
+\emph{views} on the initial, perceived model of the environment.
 
-\subsection{Representing Alternative States of the Environment}
+\subsection{Representing Alternative States of the World}
 

### 2018-01-29 paper.tex
Mnetion QSRLib

@@
 
-\uwds provides a client ({\tt spatial\_relations}) to compute both
+\uwds provides an exemplary client ({\tt spatial\_relations}) to compute both
 allo-centric and ego-centric (\ie viewer-dependent) spatial relations
 between objects (Figure~\ref{fig|spatialrelations}).
+Other libraries, like QSRLib~\ref{gatsoulis2016qsrlib}, that implement
+computational models of Qualitative Spatial Relations could be trivially
+combined with \uwds to provide more advanced geometric analysis.
 

### 2018-02-23 paper.tex
Back to IROS template

@@
-\documentclass[conference]{IEEEtran}
-\usepackage{times}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%2345678901234567890123456789012345678901234567890123456789012345678901234567890
+%        1         2         3         4         5         6         7         8
 
-% numbers option provides compact numerical references in the text. 
-\usepackage[numbers]{natbib}
+\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
 
+%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
+
+\IEEEoverridecommandlockouts                              % This command is only needed if 
+                                                          % you want to use the \thanks command
+
+\overrideIEEEmargins                                      % Needed to meet printer requirements.
+
+\usepackage[utf8x]{inputenc}
 \usepackage{graphicx}
+
 \usepackage{tikz}
@@
 
-\title{{\sc underworlds}: Cascading Situation Assessment for Robots
+\title{\LARGE \bf
+{\sc underworlds}: Cascading Situation Assessment for Robots
 }
 
-\author{Author Names Omitted for Anonymous Review. Paper-ID [add your ID here]}
-
-%\author{\authorblockN{Séverin Lemaignan}
-%\authorblockA{Centre for Robotics and Neural Systems\\
-%University of Plymouth\\
-%Plymouth, United Kingdom\\
-%Email: severin.lemaignan@plymouth.ac.uk}
-%\and
-%\authorblockN{Yoan Sallami, Aurélie Clodic and Rachid Alami}
-%\authorblockA{Twentieth Century Fox\\
-%Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\and
-%\authorblockN{James Kirk\\ and Montgomery Scott}
-%\authorblockA{Starfleet Academy\\
-%San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212\\
-%Fax: (888) 555--1212}}
+\author{Séverin Lemaignan$^{1}$, Yoan Sallami$^{2}$, Christopher Wallbridge$^{1}$, Aurélie Clodic$^{2}$ and Rachid Alami$^{2}$%
+\thanks{$^{1}$Authors are with Centre for Robotics and Neural Systems,
+        Plymouth University, Plymouth PL48AA, United Kingdom
+        {\tt\small firstname.surname@plymouth.ac.uk},
+	$^{2}$Authors are with CNRS LAAS,
+        Toulouse, France
+        {\tt\small firstname.surname@laas.fr}}%
+}
 
@@
 \maketitle
+\thispagestyle{empty}
+\pagestyle{empty}
 
@@
 
-\IEEEpeerreviewmaketitle
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \uwds is a distributed and lightweight open-source
-%framework \footnote{\url{https://github.com/severin-lemaignan/underworlds}} that
-framework \footnote{[GitHub URL hidden for blind review]} that
+framework \footnote{\url{https://github.com/underworlds-robot/underworlds}} that
 enables robot programmers to build and refine spatial and temporal
@@
 
-%\uwds has been used within the European MuMMER project
-\uwds has been used within the [hidden for blind review] project
+\uwds has been used within the European MuMMER project
 in order to conceptualise and visualise the spatial relations and visibility of
@@
     \includegraphics[width=0.9\linewidth]{l2tor}
-    %\caption{Schema of the \uwds architecture used in the MuMMER project. }
-    \caption{TODO LAAS update archi -- Schema of the \uwds architecture used in the [hidden for blind review] project. }
+    \caption{TODO LAAS update archi -- Schema of the \uwds architecture used in
+    the MuMMER project. }
     \label{fig|mummerarchitecture}
@@
 
-%This work has been supported by the EU H2020 Marie Sk\l odowska-Curie Actions
-%project DoRoThy (grant 657227) and the EU H2020 L2TOR project (grant 688014).
+This work has been supported by the EU H2020 Marie Sk\l odowska-Curie Actions
+project DoRoThy (grant 657227),  the EU H2020 MuMMER project (grant TODO LAAS)
+and the EU H2020 L2TOR project (grant 688014).
 
-%The authors also acknowledge the contribution of the HRI lab of LAAS-CNRS (where this
-%work debuted), and in particular, the influential ideas of Rachid Alami.
-
-[Hidden for blind review]
-
-\bibliographystyle{plainnat}
+\bibliographystyle{IEEEtran}
 \bibliography{bibliography}

### 2018-02-26 paper.tex
Update on Overleaf.

@@
         {\tt\small firstname.surname@plymouth.ac.uk},
-	$^{2}$Authors are with CNRS LAAS,
-        Toulouse, France
+	$^{2}$Authors are with LAAS-CNRS, Université de Toulouse, CNRS, Toulouse, France
         {\tt\small firstname.surname@laas.fr}}%
@@
 \uwds is a distributed and lightweight open-source
-framework \footnote{\url{https://github.com/underworlds-robot/underworlds}} that
+framework\footnote{\url{https://github.com/underworlds-robot/underworlds}} that
 enables robot programmers to build and refine spatial and temporal

### 2018-02-26 paper.tex
Removed dependency on tikz that caused text formatting issues
Fixed missing ref

@@
 
-\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
+\documentclass[letterpaper, 10pt, conference]{ieeeconf}
 
@@
 
-\usepackage{tikz}
-\usetikzlibrary{graphs,graphdrawing,arrows}
-\usegdlibrary{trees, layered}
-\usepackage{booktabs} % For formal tables
-\usepackage{multicol}
+%\usepackage{tikz}
+%\usetikzlibrary{graphs,graphdrawing,arrows}
+%\usegdlibrary{trees, layered}
 \usepackage[bookmarks=true]{hyperref}
@@
 
-\uwds is specifically inspired geometric and temporal reasoner like {\sc Spark}
+\uwds is specifically inspired by geometric and temporal reasoners like {\sc Spark}
 (\emph{SPAtial Reasoning \& Knowledge})~\cite{sisbot2011situation}.  {\sc Spark}
@@
 robot's software architecture. This also distinguishes \uwds from complex
-cognitive toolkits like KnowRob (as found in OpenEASE~\ref{beetz2015open}).
+cognitive toolkits like KnowRob (as found in OpenEASE~\cite{beetz2015open}).
 While these tools maintain a spatio-temporal model of the world, this model is
@@
 between objects (Figure~\ref{fig|spatialrelations}).
-Other libraries, like QSRLib~\ref{gatsoulis2016qsrlib}, that implement
+Other libraries, like QSRLib~\cite{gatsoulis2016qsrlib}, that implement
 computational models of Qualitative Spatial Relations could be trivially
@@
 \end{figure}
-
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{laasudws}
+    \caption{TODO LAAS update archi -- Schema of the \uwds architecture used in
+    the MuMMER project. }
+    \label{fig|mummerarchitecture}
+\end{figure}
 The interaction scenario requires a number of \uwds clients. The clients are

### 2018-02-26 paper.tex
Simplified fig 1

@@
     refinement of the models (\emph{cascading} situation assessment),
-    effectively adapted to each client's needs. \textit{Dashed yellow nodes represent
-    other possible components in the system that do not interact directly with
-    the \uwds network}.}
+    effectively adapted to each client's needs.}
 

### 2018-02-26 paper.tex
Update on Overleaf.
Large clean-up, save ~2 pages

@@
 internal and not meant to be made widely accessible to other external processes.
-\uwds focuses instead on reusability and sharing of distributed spatio-temporal
+\uwds focuses instead on re-usability and sharing of distributed spatio-temporal
 models.  As such, \uwds can be seen as a middleware for spatio-temporal world
-models and, contrary to KnowRob, it does \emph{not} provide any high-level
-processing or reasoning capability (this is left to the users).
+models and, contrary to KnowRob, it does \emph{not} provide any intrinsic high-level
+processing or reasoning capability. Such reasoning skills are implemented in
+loosely-coupled \emph{clients} (see Section~\ref{apiclients} hereafter).
 
@@
 visibility of objects might need high-resolution models 
-for accurate 3D rendering. This requirement of multiple
+for accurate 3D visibility testing. This requirement of multiple
 task-specific representations has been framed as the need for \emph{deep
@@
 deal with these different needs: full geometric data can be represented, but is
-not first-class citizen (a basic task like displaying a 3D mesh at an arbitrary
+not first-class citizen: a basic task like displaying a 3D mesh at an arbitrary
 position is not particularly easy to perform with ROS, requiring the combination
 of static Collada meshes, a URDF kinematic description, TF broadcasters, and a
-3D visualisation tool like RViz) and critically,
-representing and reasoning on alternative states of the environment is not
-directly feasible.
+3D visualisation tool like RViz. Critically, simultaneously representing and
+reasoning on alternative states of the environment is not directly feasible.
 
@@
 Therefore, a post-process stage (for instance, using a physics simulation
-engine) is needed to displace the objects seen by the robot into
+engine) is needed to move the objects seen by the robot into
 physically-correct positions. Implemented with a classical approach (for
@@
 
-\subsection{Formal description}
-
-An \uwds \emph{world} $\W^k \in \Ws$ consists in a pair (\emph{scene}, \emph{timeline}),
-noted $\{\mathcal{S}^k,
-\mathcal{T}^k\}$.
-
-$\mathcal{S}^k=\{N^k, T^k\}$ is a tree representing a spatial scene (\emph{scene
-graph}), with $N^k$ the set of nodes and $T^k$ the set of 6D geometric
-transformations between the nodes ($|T^k| = |N^k| - 1$). $n^k_0 \in N^k$ is
-called the \emph{root node} of the scene $\mathcal{S}^k$.
-
-$\mathcal{T}^k=\{Evt^k, Sit^k\}$ is the world's \emph{timeline}, consisting in a
-set of events $Evt^k$ (duration-less states) and a set of situations $Sit^k$
-(states with a start time and an option end time).
-
-\uwds \emph{operators} (noted $\op\colon \Ws^{0\ldots m} \rightarrow
-\Ws^{0\ldots n}$) transform one world into another. Examples of operators
-include \emph{filters} (noted $\filter$), a common type of unary, antiextensive
-transformations such as $\filter\W^k \subseteq \W^k$, and \emph{generators}
-(noted $\generator$), a class of nullary operators that are used to create new
-worlds (for instance, from perception routines).
-
-Finally, an \uwds \emph{network} $\mathcal{N} = \{\Ws, \Os\}$ is a directed graph (cycles are
-permitted) of worlds $\Ws=\{\W^k\}, k\in0\ldots n$ and operators
-$\Os=\{\op^l\}, l \in 0\ldots m$. As described in the following sections,
-operators are typically implemented as independent software processes that connect at
-runtime to their operands, \ie the \uwds worlds, to read and write them. These
-processes are hereafter called \emph{clients} of the \uwds network.
+%\subsection{Formal description}
+%
+%An \uwds \emph{world} $\W^k \in \Ws$ consists in a pair (\emph{scene}, \emph{timeline}),
+%noted $\{\mathcal{S}^k,
+%\mathcal{T}^k\}$.
+%
+%$\mathcal{S}^k=\{N^k, T^k\}$ is a tree representing a spatial scene (\emph{scene
+%graph}), with $N^k$ the set of nodes and $T^k$ the set of 6D geometric
+%transformations between the nodes ($|T^k| = |N^k| - 1$). $n^k_0 \in N^k$ is
+%called the \emph{root node} of the scene $\mathcal{S}^k$.
+%
+%$\mathcal{T}^k=\{Evt^k, Sit^k\}$ is the world's \emph{timeline}, consisting in a
+%set of events $Evt^k$ (duration-less states) and a set of situations $Sit^k$
+%(states with a start time and an option end time).
+%
+%\uwds \emph{operators} (noted $\op\colon \Ws^{0\ldots m} \rightarrow
+%\Ws^{0\ldots n}$) transform one world into another. Examples of operators
+%include \emph{filters} (noted $\filter$), a common type of unary, antiextensive
+%transformations such as $\filter\W^k \subseteq \W^k$, and \emph{generators}
+%(noted $\generator$), a class of nullary operators that are used to create new
+%worlds (for instance, from perception routines).
+%
+%Finally, an \uwds \emph{network} $\mathcal{N} = \{\Ws, \Os\}$ is a directed graph (cycles are
+%permitted) of worlds $\Ws=\{\W^k\}, k\in0\ldots n$ and operators
+%$\Os=\{\op^l\}, l \in 0\ldots m$. As described in the following sections,
+%operators are typically implemented as independent software processes that connect at
+%runtime to their operands, \ie the \uwds worlds, to read and write them. These
+%processes are hereafter called \emph{clients} of the \uwds network.
 
@@
 to be an \emph{acyclic} graph on Figure~\ref{fig|scene}, but does not have to
-be in the general case) of worlds and operators connecting the worlds together.
-In the following, we however tend to adopt a software engineering perspective,
-in which the \uwds network is made of \emph{clients} (which implement the
-operators) connected through shared data structures called \emph{worlds}.
-The formal model and this engineering-oriented model are isomorphic.
+be in the general case) of worlds, with clients connecting the worlds to each
+others.
+%In the following, we however tend to adopt a software engineering perspective,
+%in which the \uwds network is made of \emph{clients} (which implement the
+%operators) connected through shared data structures called \emph{worlds}.
+%The formal model and this engineering-oriented model are isomorphic.
 
@@
 
-Software components implementing \uwds operators and accessing worlds are
-called clients. Some standard clients (like a 3D visualisation tool) are
-provided with the \uwds package (see section~\ref{std_clients}). Clients are
-otherwise written by the end users using the \uwds client API (see
-section~\ref{api}).
-
+Software components implementing accessing \uwds worlds are
+called clients. 
 Clients can both read and write onto the worlds they are connected to, and
@@
 
-Three specific types of clients can be distinguished: \textbf{generator
-clients} which implement generator operators, and create and update worlds
-(`write-only' client, like the \emph{Sensor fusion} and \emph{ROS interface}
-clients in Figure~\ref{fig|scene}); \textbf{leaf clients} which on the contrary
-only read worlds, without modifying them (like the \emph{Computation of spatial
-relations} client on Figure~\ref{fig|scene}); and \textbf{transformers} which
-transform one representation into another (like the \emph{Semantic mapping}
-client on Figure~\ref{fig|scene}). \textbf{Filters} are a specfic class of
-transformer clients that copy an input world into an output world, performing some
-filtering operation in-between (like the client \textit{Physics-based position
-correction}).
+\uwds provides several standard clients (like a 3D visualisation tool or a
+physics engine simulator). Clients are however typically written by the end
+users, depending on the needs of one's specific architecture.
+
+%Three specific types of clients can be distinguished: \textbf{generator
+%clients} which implement generator operators, and create and update worlds
+%(`write-only' client, like the \emph{Sensor fusion} and \emph{ROS interface}
+%clients in Figure~\ref{fig|scene}); \textbf{leaf clients} which on the contrary
+%only read worlds, without modifying them (like the \emph{Computation of spatial
+%relations} client on Figure~\ref{fig|scene}); and \textbf{transformers} which
+%transform one representation into another (like the \emph{Semantic mapping}
+%client on Figure~\ref{fig|scene}). \textbf{Filters} are a specfic class of
+%transformer clients that copy an input world into an output world, performing some
+%filtering operation in-between (like the client \textit{Physics-based position
+%correction}).
 
@@
 
-Nodes in an \uwds scene graph have four possible types: \textbf{objects} that
+Nodes in an \uwds scene graph have three possible types: \textbf{objects} that
 represent concrete physical objects (typically with one or several associated 3D
@@
 frames or groups of objects; \textbf{perspectives} that represent viewpoints
-of the scene (like cameras or human gaze); and \textbf{fields} that represent
-scalar or vector fields (like the visibility of an object, the working space of
-robot, \etc -- note that fields are not yet implemented in the current version
-of \uwds, see section~\ref{futurework}).
+of the scene (like cameras or human gaze).
+%; and \textbf{fields} that represent
+%scalar or vector fields (like the visibility of an object, the working space of
+%robot, \etc -- note that fields are not yet implemented in the current version
+%of \uwds, see section~\ref{futurework}).
 
@@
 More interesting is the time complexity of distributing changes across an \uwds
-network. With $n$ the number of worlds and $m$ the number of operators in a
-network $\mathcal{N}$, the worst-case (when every world is a parameter of every
-operator) time complexity of creating or updating a node and propagating the
+network. With $n$ the number of worlds and $m$ the number of clients in an \uwds
+network, the worst-case (when every world is a parameter of every
+client) time complexity of creating or updating a node and propagating the
 change across the network is $O(n \times m)$ (this effectively corresponds to
@@
 (with $p$ nodes) at each time step, the complexity of propagating these changes across
-the network would be $O(p \times m)$.
+the network would be $O(p \times m)$. Figure~\ref{fig|performances} shows
+measured propagation time for one change across 20 cascading worlds.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{performances}
+    \caption{Propagation times of one change (node creation)
+    across $n$ worlds. The test is performed by running $n-1$ pass-through
+    filters that monitor one world and replicate
+    any changes into the next world. Durations measured over 20 runs, performed on
+    a 8 core machine.}
+    \label{fig|performances}
+\end{figure}
 
 \section{API \& Clients}
+\label{apiclients}
 
@@
 
-    in_world = ctx.worlds["example_in"]
-    out_world = ctx.worlds["example_out"]
+    in_world = ctx.worlds["world1"]
+    out_world = ctx.worlds["world2"]
 
@@
 
-Geometric data like meshes can be loaded into a world (or rather, into the whole
-network as geometric data are treated as immutable resources shared amongst all
-worlds) by simply sending an array of vertices and faces to the \uwds server.
-As such, any client can load 3D data that is automatically made available to the
-whole network. For example, an object recognition module recognises a new
-object. It sends the mesh to the \uwds server and creates a new node which
-references the mesh. At the next frame update, the 3D viewer sees a new node
-referencing an unknown mesh. The viewer requests the mesh data from the
-server, stores it internally (\eg as an OpenGL buffer), and renders it. Whenever
-the node transformation is updated, the 3D viewer can re-render the node without
-having to download the mesh from the server again.
+%Geometric data like meshes can be loaded into a world (or rather, into the whole
+%network as geometric data are treated as immutable resources shared amongst all
+%worlds) by simply sending an array of vertices and faces to the \uwds server.
+%As such, any client can load 3D data that is automatically made available to the
+%whole network. For example, an object recognition module recognises a new
+%object. It sends the mesh to the \uwds server and creates a new node which
+%references the mesh. At the next frame update, the 3D viewer sees a new node
+%referencing an unknown mesh. The viewer requests the mesh data from the
+%server, stores it internally (\eg as an OpenGL buffer), and renders it. Whenever
+%the node transformation is updated, the 3D viewer can re-render the node without
+%having to download the mesh from the server again.
 
@@
 \url{http://assimp.sourceforge.net/}}.
+
+\subsubsection{Physics simulation}
+
+TODO LAAS
 
@@
 creation of new worlds representing the estimated perspectives of the different
-agents. Listing~\ref{lst|humanperspective} provides the complete code of such an
-\uwds client.
+agents.
+%Listing~\ref{lst|humanperspective} provides the complete code of such an
+%\uwds client.
 
-\begin{listing}[h!]
-\begin{pythoncode}
-import copy
-import underworlds
-from underworlds.tools.visibility import VisibilityMonitor
+%
+%\begin{listing}[h!]
+%\begin{pythoncode}
+%import copy
+%import underworlds
+%from underworlds.tools.visibility import VisibilityMonitor
+%
+%with underworlds.Context("Human PoV") as ctx:
+%
+%    source = ctx.worlds["base"]
+%    target = ctx.worlds["human_perspective"]
+%
+%    # pick up the first node named 'human'
+%    human = source.scene.nodebyname("human")[0]
+%    target.scene.nodes.append(human)
+%
+%    # VisibilityMonitor computes the set of visible 
+%    # nodes from a given viewpoint
+%    visibility = VisibilityMonitor(ctx, source)
+%
+%    node_mapping = {}
+%
+%    while True:
+%
+%        for node in visibility.from_camera(human):
+%
+%            newnode = node.copy()
+%
+%            # track the correspondence between nodes
+%            # in source and target scenes
+%            if node in node_mapping:
+%                newnode.id = node_mapping[node].id
+%            else:
+%                node_mapping[node] = newnode
+%
+%            # reparent the nodes to the new root
+%            if node.parent == source.scene.rootnode.id:
+%                newnode.parent = target.scene.rootnode.id
+%
+%            target.scene.nodes.update(newnode)
+%
+%        source.scene.waitforchanges()
+%
+%\end{pythoncode}
+%    \caption{Example of an \uwds client that creates a world named {\tt
+%    human\_perspective} from a source world {\tt base} by only updating objects
+%    that are in the field of view of the human. As such, {\tt human\_perspective} is
+%    a real-time model of the human perspective on the scene.}
+%
+%    \label{lst|humanperspective}
+%\end{listing}
+%
 
-with underworlds.Context("Human PoV") as ctx:
-
-    source = ctx.worlds["base"]
-    target = ctx.worlds["human_perspective"]
-
-    # pick up the first node named 'human'
-    human = source.scene.nodebyname("human")[0]
-    target.scene.nodes.append(human)
-
-    # VisibilityMonitor computes the set of visible 
-    # nodes from a given viewpoint
-    visibility = VisibilityMonitor(ctx, source)
-
-    node_mapping = {}
-
-    while True:
-
-        for node in visibility.from_camera(human):
-
-            newnode = node.copy()
-
-            # track the correspondence between nodes
-            # in source and target scenes
-            if node in node_mapping:
-                newnode.id = node_mapping[node].id
-            else:
-                node_mapping[node] = newnode
-
-            # reparent the nodes to the new root
-            if node.parent == source.scene.rootnode.id:
-                newnode.parent = target.scene.rootnode.id
-
-            target.scene.nodes.update(newnode)
-
-        source.scene.waitforchanges()
-
-\end{pythoncode}
-    \caption{Example of an \uwds client that creates a world named {\tt
-    human\_perspective} from a source world {\tt base} by only updating objects
-    that are in the field of view of the human. As such, {\tt human\_perspective} is
-    a real-time model of the human perspective on the scene.}
-
-    \label{lst|humanperspective}
-\end{listing}
-
-\section{Performances and Application Example}
+\section{Application Example}
 \label{application}
-
-\subsection{Performances}
-\label{performances}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{performances}
-    \caption{Propagation times of one change (node creation)
-    across $n$ worlds. The test is performed by running $n-1$ pass-through
-    filters that monitor one world and replicate
-    any changes into the next world. Durations measured over 20 runs, performed on
-    a 8 core machine.}
-    \label{fig|performances}
-\end{figure}
-
-The main performance metric of \uwds is the \emph{propagation time}: the time is
-take of a change in a given world to cascade to all the dependant
 
@@
 situations). In that sense, \uwds really is a \emph{distributed datastructure}
-with a narrow focus on the needs of spatio-temporal representations.
+that addresses the specific needs of spatio-temporal modelling,
+including the modelling of hypothetical, alternative world models, something
+that traditional middlewares like ROS do not address adequately.
 
@@
 \emph{Fields} are also part of the \uwds design, but are not yet implemented.
-
-\paragraph*{Representation of fields} fields are commonly used to represent
+Fields are commonly used to represent
 continuously-valued spatial entities. Fields might or might not be spatially
@@
 
-\subsubsection{Manipulation of representations} \uwds makes it possible to
-cheaply clone and modify multiple worlds. However, little is provided to
-jointly manipulate sets of worlds. In that regard, we plan to add a mechanism to
-efficiently \emph{compare} two worlds. This would permit a client
-to quickly identify modifications applied to a world, and could prove especially
-useful when combined with temporal events to answer questions like ``what has
-changed in the scene?''. Comparisons between worlds involves the definition of
-an efficient fuzzy \emph{diff} algorithm for 3D scene graphs, able to
-selectively ignore non-relevant micro displacements.
+%\subsubsection{Manipulation of representations} \uwds makes it possible to
+%cheaply clone and modify multiple worlds. However, little is provided to
+%jointly manipulate sets of worlds. In that regard, we plan to add a mechanism to
+%efficiently \emph{compare} two worlds. This would permit a client
+%to quickly identify modifications applied to a world, and could prove especially
+%useful when combined with temporal events to answer questions like ``what has
+%changed in the scene?''. Comparisons between worlds involves the definition of
+%an efficient fuzzy \emph{diff} algorithm for 3D scene graphs, able to
+%selectively ignore non-relevant micro displacements.
 
-\subsubsection{Implementation and Integration} finally, we plan to continue to
+\subsubsection{Implementation and Integration} we plan to continue to
 improve the integration of \uwds into existing software architectures. A
@@
 
-In term of implementation, we are also investigating alternative back-ends that
-could compete with gRPC to provide the synchronisation mechanisms between
-clients. Due to the high-level nature of \uwds (a set of shared, tree-like data
-structures), the use of the GIT version control system is one particular (if
-unusual) option being studied. The scene graph would be represented as a
-hierarchy of directories, nodes would be files in these directories, creating
-new worlds would consist in creating a branch in the repository, and sharing
-would be supported through the usual pull/push GIT mechanisms. Initial
-prototyping shows that under the right conditions (in particular, an in-memory
-file system), performance is not an issue. Automatic conflicts resolution, on
-the other hand, might prove more challenging.
+%In term of implementation, we are also investigating alternative back-ends that
+%could compete with gRPC to provide the synchronisation mechanisms between
+%clients. Due to the high-level nature of \uwds (a set of shared, tree-like data
+%structures), the use of the GIT version control system is one particular (if
+%unusual) option being studied. The scene graph would be represented as a
+%hierarchy of directories, nodes would be files in these directories, creating
+%new worlds would consist in creating a branch in the repository, and sharing
+%would be supported through the usual pull/push GIT mechanisms. Initial
+%prototyping shows that under the right conditions (in particular, an in-memory
+%file system), performance is not an issue. Automatic conflicts resolution, on
+%the other hand, might prove more challenging.
 
@@
 the explicit of representation of alternative states of the world while
-maintaining a network of dependencies; a typology of clients
-(\emph{root}, \emph{leaf}, \emph{filter}, \emph{transformer}) to systematically
-describe such a network of representations dependencies.
+maintaining a network of dependencies.
+%; a typology of clients
+%(\emph{root}, \emph{leaf}, \emph{filter}, \emph{transformer}) to systematically
+%describe such a network of representations dependencies.
 

### 2018-02-26 paper.tex
Update on Overleaf.

@@
 This work has been supported by the EU H2020 Marie Sk\l odowska-Curie Actions
-project DoRoThy (grant 657227),  the EU H2020 MuMMER project (grant TODO LAAS)
+project DoRoThy (grant 657227),  the EU H2020 MuMMER project (grant 688147)
 and the EU H2020 L2TOR project (grant 688014).

### 2018-02-26 paper.tex
Re-insert section on L2TOR project

@@
 
-\subsection{Application to service robots and human-robot interaction}
+\subsection{Application to human-robot interaction and spatial reasoning}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{l2tor-photo2-blind}
+    \caption{A user manipulating objects on top of an RFID mat (providing object
+    localisation) with a NAO robot, while an operator monitors
+    the interaction on a tablet that displays an {\tt uwds view} client [photo blurred for blind review].}
+
+    \label{fig|l2torexample}
+\end{figure}
+
+\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor}
+in order to conceptualise and visualise the spatial relations and visibility of
+the physical objects that participants interact with. In one of the scenarios of this
+project, participants are instructed by a NAO robot to manipulate a number of
+Duplo-like animals on top of a printed background paper which serves as a
+landscape (Figure~\ref{fig|l2torexample}). The robot instructions are based on
+pedagogical strategies to promote spatial relation learning by object
+manipulation. Such instructions include commands to place the elephant
+\textit{on top} or \textit{next to} the cube, or put the giraffe \textit{inside}
+the house.  To assess the performance, the robot and the operator need to
+compute spatial relationships between objects from the visual perspective of the
+participant.
+
+\subsection{Application to joint-action situations}
 

### 2018-02-26 paper.tex
Update on Overleaf.

@@
 
-\subsubsection{Physics simulation}
+\subsubsection{Physics simulation}\label{physics}
 
@@
 
-\uwds has been used within the European MuMMER project
-in order to conceptualise and visualise the spatial relations and visibility of
-the physical objects that participants interact with.
+%\uwds has been used within the European MuMMER project
+%in order to conceptualise and visualise the spatial relations and visibility of
+%the physical objects that participants interact with.
 
-TODO LAAS
+%TODO LAAS
 
@@
 \end{figure}
-The interaction scenario requires a number of \uwds clients. The clients are
-distributed across...
+%The interaction scenario requires a number of \uwds clients. The clients are
+%distributed across...
 
 Figure~\ref{fig|mummerarchitecture} illustrates the topology of
-the \uwds network...
+the \uwds network. A first client \textit{env\_provider} provides the environment set-up and allows to build a first \texttt{ENV} world where static objects/furnitures and walls are present. 
+Then, the system cascades three worlds successively by the help of three clients.
+\textit{robot\_monitor} provides a first world where the robot is present.
+\textit{humans\_monitor} provides a new world where the humans are present.
+\textit{object\_monitor} provides a new world where the objects are present.
+At this end we obtain a world \texttt{BASE}.
 
-TODO LAAS
+This world \texttt{BASE} is then passed through a \textit{physics filter} client as explained in section \ref{physics} to obtain a \texttt{STABLE} world where all elements are present and consistent: robot, humans, objects and static environment.
+On this world, we use \textit{Computation of spatial relations} client to get information such as onTop, isIn or isAbove.
+
+This world \texttt{STABLE} is then used by a \textit{perspective filter} client to provide different perspectives given the monitored agent (in our case: human 1, human 2 or the robot itself).
+Besides giving a view of the input world in the perspective of the monitored agent, it aggregates all the information gathered by this agent.
+Consequently, it does not offer a snapshot of the agent belief at a particular time but an aggregated version along the time.
 

### 2018-02-26 paper.tex
Updates based on IROS2017 review comments

@@
 
-\subsection{Representing Alternative States of the World}
+\subsection{Representing Alternative States of the World}\label{alter}
 
@@
 Complementing the spatial representation encapsulated in the scene graph, each
-world also stores the world's \emph{timeline}. This data structure is shared
-and synchronised amongst the clients in the same way as the scene graph.
-Clients can record both \emph{events} (duration-less states) and
-\emph{situations} in the timeline, \ie states with a start time and a (possibly open-ended) end time.
+world also stores the world's \emph{timeline}. This data structure is shared and
+synchronised amongst the clients in the same way as the scene graph.  Clients
+can record and query both \emph{events} (duration-less states) and \emph{situations} in
+the timeline, \ie states with a start time and a (possibly open-ended) end time.
 
-Importantly, the \uwds server automatically generates a snapshot of
-the scene graph whenever an event or situation is added to the timeline. The
-snapshot is associated to the event, which allows clients to effectively retrieve past
-states of the world. This capability is anticipated to be primarily used by \uwds clients
-performing action recognition.
+%Importantly, the \uwds server automatically generates a snapshot of
+%the scene graph whenever an event or situation is added to the timeline. The
+%snapshot is associated to the event, which allows clients to effectively retrieve past
+%states of the world. This capability is anticipated to be primarily used by \uwds clients
+%performing action recognition.
 
@@
 the network would be $O(p \times m)$. Figure~\ref{fig|performances} shows
-measured propagation time for one change across 20 cascading worlds.
+measured propagation time for one change across up to 20 cascading worlds.
 
@@
 
-TODO LAAS
+When perception modules provide objects localization, the consistency is not guarantee, objects on a table seem to slightly flight in the air, when dropping them in a bin the localization is not updated any more when the objects are occulted, resulting in objects that fly just above the bin.
+
+Being able to manage that issues by simulating the gravity is the purpose of a dedicated client  
+{\tt physics filter} that have been designed using Bullet RT physics simulation and pybullet \url{https://pybullet.org/} library, it provides from an input world, an output world where objects described by an URDF file are filtered by the simulator, which gives us the near future of the objects based on their position, mass and inertia.
 
@@
 
+Underworlds is used 
+%within the European MuMMER project
+to be able to assess visibility and knowledge about objects, places and agents within the scene.
+
+Let's imagine a little example with two humans and a robot looking at a table.
+On the table there is an object that is only seen by one human and the robot (it is hidden for the second human).
+Figure~\ref{fig|mummerarchitecture} illustrates the topology of the \uwds network we can define for this example.
+
 \begin{figure}
@@
 
-Figure~\ref{fig|mummerarchitecture} illustrates the topology of
-the \uwds network. A first client \textit{env\_provider} provides the environment set-up and allows to build a first \texttt{ENV} world where static objects/furnitures and walls are present. 
+A first client \textit{env\_provider} provides the environment set-up and allows to build a first \texttt{ENV} world where static objects/furnitures and walls are present. 
 Then, the system cascades three worlds successively by the help of three clients.
 \textit{robot\_monitor} provides a first world where the robot is present.
-\textit{humans\_monitor} provides a new world where the humans are present.
-\textit{object\_monitor} provides a new world where the objects are present.
+\textit{humans\_monitor} provides a new world where the humans are present (using~\cite{Khalidov_Idiap-RR-02-2017}).
+\textit{object\_monitor} provides a new world where the objects are present (using {\tt ar\_track\_alvar}\footnote{\url{http://wiki.ros.org/ar\_track\_alvar}).
 At this end we obtain a world \texttt{BASE}.
 
-This world \texttt{BASE} is then passed through a \textit{physics filter} client as explained in section \ref{physics} to obtain a \texttt{STABLE} world where all elements are present and consistent: robot, humans, objects and static environment.
-On this world, we use \textit{Computation of spatial relations} client to get information such as onTop, isIn or isAbove.
+This world \texttt{BASE} is then passed through a \textit{physics filter} client as explained in section~\ref{physics} to obtain a \texttt{STABLE} world where all elements are present and consistent: robot, humans, objects and static environment.
+On this world, we use \textit{Computation of spatial relations} client to get information such as {\tt onTop}, {\tt isIn} or {\tt isAbove}.
 
 This world \texttt{STABLE} is then used by a \textit{perspective filter} client to provide different perspectives given the monitored agent (in our case: human 1, human 2 or the robot itself).
-Besides giving a view of the input world in the perspective of the monitored agent, it aggregates all the information gathered by this agent.
-Consequently, it does not offer a snapshot of the agent belief at a particular time but an aggregated version along the time.
+Besides giving a view of the input world in the perspective of the monitored agent (through the robot's own perspective), it aggregates all the information visibly gathered by this agent.
+Consequently, it does not offer a snapshot of the agent belief at a particular time but an aggregated estimation along the time.
+
+With this network, we are able to assess among other things that the object on the table is seen by the robot and human 1 but not by human 2.
+Moreover, if human 1 moves in a position where the object becomes not visible for him, the \textit{perspective filter} will keep the knowledge about the fact that he has already seen it (and keep the last position where it has been seen).
+
+%This knowledge could then be used by planner such as HATP to plan differently given the knowledge or ignorance of an agent regarding the environment.
+
+We are currently setting up this kind of network in the framework of the European project MuMMER where a Pepper robot is intended to handle interactive situations in a Mall in Finland.
+One of the situation is a guiding task where Pepper will help people to find their route by pointing them landmarks and explaining them how to reach a destination.
+The use of \uwds enables to get knowledge about what the customer sees (or has seen) or not and to handle the task accordingly.
+We plan also to be able to add information coming from the dialogue (e.g.
+if the customer says that he has seen a particular landmark that could consequently be added to his belief world).
+Moreover, we plan also to use alternative world ability given by \uwds (as explained in \ref{alter}) (e.g.
+to create a world where we put the customer in front of the desired shop )would be able to infer a possible world where the customer will be put in front of the desired shop to be able to explain him what he will see at destination.
+
+TODO LAAS: on est en train de reflechir a mettre dans l'exemple un nouvau monde simule (notre cas d'ecole c'est qu'on place le bonhomme en simule devant un magasin et on est capable de calculer ce qu'il verra en face de ce magasin) mais il faut qu'on le fasse coller avec l'exemple qu'on a
 
@@
 unbounded). We plan to represent fields in \uwds using the memory-efficient
-octomaps~\cite{hornung2013octomap}. Similarly to geometric data, octomaps
+octomaps~\cite{hornung2013octomap} or NDT-OM maps~\cite{saarinen20133d}.
+Similarly to geometric data,these datastructures
 will not be directly stored with the nodes (nodes will refer to them through 

### 2018-02-27 paper.tex
Update on Overleaf.

@@
 
-\subsection{Representing Alternative States of the World}\label{alter}
+\subsection{Representing Alternative States of the World}
 
@@
 Complementing the spatial representation encapsulated in the scene graph, each
-world also stores the world's \emph{timeline}. This data structure is shared and
-synchronised amongst the clients in the same way as the scene graph.  Clients
-can record and query both \emph{events} (duration-less states) and \emph{situations} in
-the timeline, \ie states with a start time and a (possibly open-ended) end time.
+world also stores the world's \emph{timeline}. This data structure is shared
+and synchronised amongst the clients in the same way as the scene graph.
+Clients can record both \emph{events} (duration-less states) and
+\emph{situations} in the timeline, \ie states with a start time and a (possibly open-ended) end time.
 
-%Importantly, the \uwds server automatically generates a snapshot of
-%the scene graph whenever an event or situation is added to the timeline. The
-%snapshot is associated to the event, which allows clients to effectively retrieve past
-%states of the world. This capability is anticipated to be primarily used by \uwds clients
-%performing action recognition.
+Importantly, the \uwds server automatically generates a snapshot of
+the scene graph whenever an event or situation is added to the timeline. The
+snapshot is associated to the event, which allows clients to effectively retrieve past
+states of the world. This capability is anticipated to be primarily used by \uwds clients
+performing action recognition.
 
@@
 the network would be $O(p \times m)$. Figure~\ref{fig|performances} shows
-measured propagation time for one change across up to 20 cascading worlds.
+measured propagation time for one change across 20 cascading worlds.
 
@@
 if the customer says that he has seen a particular landmark that could consequently be added to his belief world).
-Moreover, we plan also to use alternative world ability given by \uwds (as explained in \ref{alter}) (e.g.
-to create a world where we put the customer in front of the desired shop )would be able to infer a possible world where the customer will be put in front of the desired shop to be able to explain him what he will see at destination.
+Moreover, we plan also to use imaginary world (as explained in \ref{})would be able to infer a possible world where the customer will be put in front of the desired shop to be able to explain him what he will see at destination.
 
@@
 unbounded). We plan to represent fields in \uwds using the memory-efficient
-octomaps~\cite{hornung2013octomap} or NDT-OM maps~\cite{saarinen20133d}.
-Similarly to geometric data,these datastructures
+octomaps~\cite{hornung2013octomap}. Similarly to geometric data, octomaps
 will not be directly stored with the nodes (nodes will refer to them through 

### 2018-02-27 paper.tex
Update on Overleaf.

@@
 if the customer says that he has seen a particular landmark that could consequently be added to his belief world).
-Moreover, we plan also to use imaginary world (as explained in \ref{})would be able to infer a possible world where the customer will be put in front of the desired shop to be able to explain him what he will see at destination.
+Moreover, we plan also to use alternative world ability given by \uwds (as explained in \ref{alter}) (e.g.
+to create a world where we put )would be able to infer a possible world where the customer will be put in front of the desired shop to be able to explain him what he will see at destination.
 

### 2018-02-27 paper.tex
Update on Overleaf.

@@
 Moreover, we plan also to use alternative world ability given by \uwds (as explained in \ref{alter}) (e.g.
-to create a world where we put )would be able to infer a possible world where the customer will be put in front of the desired shop to be able to explain him what he will see at destination.
+to create a world where we put the customer in front of )would be able to infer a possible world where the customer will be put in front of the desired shop to be able to explain him what he will see at destination.
 

### 2018-02-27 paper.tex
Update on Overleaf.

--- 2018-02-27 paper.tex
@@
 Moreover, we plan also to use alternative world ability given by \uwds (as explained in \ref{alter}) (e.g.
-to create a world where we put )would be able to infer a possible world where the customer will be put in front of the desired shop to be able to explain him what he will see at destination.
+to create a world where we put the customer in front of the desired shop )would be able to infer a possible world where the customer will be put in front of the desired shop to be able to explain him what he will see at destination.
 

### 2018-02-27 paper.tex
Update on Overleaf.

--- 2018-02-27 paper.tex
@@
 
-Underworlds is used 
+\uwds is used 
 %within the European MuMMER project
@@
 \textit{humans\_monitor} provides a new world where the humans are present (using~\cite{Khalidov_Idiap-RR-02-2017}).
-\textit{object\_monitor} provides a new world where the objects are present (using {\tt ar\_track\_alvar}\footnote{\url{http://wiki.ros.org/ar\_track\_alvar}).
+\textit{object\_monitor} provides a new world where the objects are present (using {\tt ar\_track\_alvar}\footnote{\url{http://wiki.ros.org/ar_track_alvar}}).
 At this end we obtain a world \texttt{BASE}.
@@
 
+It has to be noticed that this knowledge could be wrong relatively to the current "real" state of the world, but it represents the monitored agent belief about the world.
+
 %This knowledge could then be used by planner such as HATP to plan differently given the knowledge or ignorance of an agent regarding the environment.
@@
 Moreover, we plan also to use alternative world ability given by \uwds (as explained in \ref{alter}) (e.g.
-to create a world where we put )would be able to infer a possible world where the customer will be put in front of the desired shop to be able to explain him what he will see at destination.
-
-TODO LAAS: on est en train de reflechir a mettre dans l'exemple un nouvau monde simule (notre cas d'ecole c'est qu'on place le bonhomme en simule devant un magasin et on est capable de calculer ce qu'il verra en face de ce magasin) mais il faut qu'on le fasse coller avec l'exemple qu'on a
+to create a world where we put the customer in front of the desired shop to infer (and be able to explain him) what he will see at destination).
 

### 2018-02-28 paper.tex
Updated LAAS archi diagram
Proof-read section on joint action

@@
 \subsection{Spatial Reasoning and Perspective Taking}
+\label{spatial}
 
@@
 
-\uwds is used 
-%within the European MuMMER project
-to be able to assess visibility and knowledge about objects, places and agents within the scene.
+\uwds is used within the European project MuMMER for service robots to compute
+visibility and knowledge about objects, places and agents within a mall
+environment.
 
-Let's imagine a little example with two humans and a robot looking at a table.
-On the table there is an object that is only seen by one human and the robot (it is hidden for the second human).
-Figure~\ref{fig|mummerarchitecture} illustrates the topology of the \uwds network we can define for this example.
+We present here a simplified scenario, yet representative of situations which
+are processed in real-time by MuMMER robots: two humans and a robot are looking
+at a table and have to coordinate joint actions (pick and place).  One object on
+the table (the green box) is only visible to one human and the robot, but hidden
+to the second human. The robot needs to take into account this fact to generate
+appropriate and legible joint manipulation actions.
+Figure~\ref{fig|mummerarchitecture} illustrates the topology of the \uwds
+network that we use to this end.
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{l2tor-photo2-blind}
-    \caption{TODO LAAS: update photo -> MuMMER picture?}
-    \label{fig|mummer-example}
-\end{figure}
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=\linewidth]{l2tor-photo2-blind}
+%    \caption{TODO LAAS: update photo -> MuMMER picture?}
+%    \label{fig|mummer-example}
+%\end{figure}
 
@@
 
-%TODO LAAS
-
 \begin{figure}
     \centering
-    \includegraphics[width=0.9\linewidth]{l2tor}
-    \caption{TODO LAAS update archi -- Schema of the \uwds architecture used in
-    the MuMMER project. }
-    \label{fig|mummerarchitecture}
-\end{figure}
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\linewidth]{laasudws}
-    \caption{TODO LAAS update archi -- Schema of the \uwds architecture used in
-    the MuMMER project. }
+    \includegraphics[width=\linewidth]{laasudws}
+    \caption{Schema of the \uwds architecture used in the MuMMER project.
+    Clients (blue boxes) read and generate worlds (whose names are in typewriter
+    fonts).}
     \label{fig|mummerarchitecture}
@@
 
-A first client \textit{env\_provider} provides the environment set-up and allows to build a first \texttt{ENV} world where static objects/furnitures and walls are present. 
-Then, the system cascades three worlds successively by the help of three clients.
-\textit{robot\_monitor} provides a first world where the robot is present.
-\textit{humans\_monitor} provides a new world where the humans are present (using~\cite{Khalidov_Idiap-RR-02-2017}).
-\textit{object\_monitor} provides a new world where the objects are present (using {\tt ar\_track\_alvar}\footnote{\url{http://wiki.ros.org/ar_track_alvar}}).
-At this end we obtain a world \texttt{BASE}.
+A first client, \textit{env\_provider}, provides the environment models and
+allows to build a first \texttt{ENV} world where static objects, furnitures and
+walls are present.  Then, three worlds cascade through three (independent)
+clients: \textit{robot\_state\_monitor} augments {\tt ENV} with the robot state
+(using underneath the ROS robot state publisher node) and broadcast a new world
+{\tt ENV\_ROBOTS}. \textit{objects\_monitor} then recognises and adds the
+dynamic objects (using {\tt
+ar\_track\_alvar}\footnote{\url{http://wiki.ros.org/ar_track_alvar}}).
+\textit{object\_monitor} finally detects and adds the humans
+(using~\cite{Khalidov_Idiap-RR-02-2017}). It broadcasts a world called {\tt
+BASE} that eventually contains the static environment, the robots, the dynamic
+objects and the detected humans.
 
-This world \texttt{BASE} is then passed through a \textit{physics filter} client as explained in section~\ref{physics} to obtain a \texttt{STABLE} world where all elements are present and consistent: robot, humans, objects and static environment.
-On this world, we use \textit{Computation of spatial relations} client to get information such as {\tt onTop}, {\tt isIn} or {\tt isAbove}.
+The world \texttt{BASE} goes through a \textit{physics filter} client as
+explained in section~\ref{physics} to obtain the \texttt{STABLE} world where all
+elements are present with physically-consistent locations. This
+physically-correct world is used by the
+\textit{computation\_of\_spatial\_relations} client to compute spatial relations
+such as {\tt onTop}, {\tt isIn} or {\tt isAbove} (see Section~\ref{spatial}).
 
-This world \texttt{STABLE} is then used by a \textit{perspective filter} client to provide different perspectives given the monitored agent (in our case: human 1, human 2 or the robot itself).
-Besides giving a view of the input world in the perspective of the monitored agent (through the robot's own perspective), it aggregates all the information visibly gathered by this agent.
-Consequently, it does not offer a snapshot of the agent belief at a particular time but an aggregated estimation along the time.
+The world \texttt{STABLE} is also used by a \textit{perspectives\_filter} client
+to compute the different visual perspectives of each agent (in our case: human
+1, human 2 and the robot itself). In addition to a 3D rendering of the input
+world from the perspective of the agent, it aggregates the history of what was
+visible to the agent at a given point in time. As such, it does not only offer a
+snapshot of the agent visual perspective at the current time but also act as the
+visual memory of each agent.
 
-With this network, we are able to assess among other things that the object on the table is seen by the robot and human 1 but not by human 2.
-Moreover, if human 1 moves in a position where the object becomes not visible for him, the \textit{perspective filter} will keep the knowledge about the fact that he has already seen it (and keep the last position where it has been seen).
+With this network, the robot can easily compute that an object on
+the table is only seen by the human 1 and not the human 2; additionally, if
+human 1 moves in a position where the object is not visible anymore to him, the
+\textit{perspective\_filter} will maintain the knowledge that the human had
+seen it (and keep the last position where it has been seen).
 
-It has to be noticed that this knowledge could be wrong relatively to the current "real" state of the world, but it represents the monitored agent belief about the world.
+\uwds makes it possible to implement such a geometric reasoning pipeline in a
+fully decoupled way, and each intermediary ``world'' can be easily introspected
+at run-time. As such, \uwds facilitates the implementation and debugging of complex
+spatio-temporal reasoning pipelines.
 
@@
 
-We are currently setting up this kind of network in the framework of the European project MuMMER where a Pepper robot is intended to handle interactive situations in a Mall in Finland.
-One of the situation is a guiding task where Pepper will help people to find their route by pointing them landmarks and explaining them how to reach a destination.
-The use of \uwds enables to get knowledge about what the customer sees (or has seen) or not and to handle the task accordingly.
-We plan also to be able to add information coming from the dialogue (e.g.
-if the customer says that he has seen a particular landmark that could consequently be added to his belief world).
-Moreover, we plan also to use alternative world ability given by \uwds (as explained in \ref{alter}) (e.g.
-to create a world where we put the customer in front of the desired shop to infer (and be able to explain him) what he will see at destination).
+We are currently deploying a similar network in the framework of the
+European project MuMMER where a Pepper robot handles interactive
+situations in a large shopping centre in Finland. One of the situation is a guiding task where
+Pepper will help people to find their route by pointing them landmarks and
+explaining them how to reach a destination. \uwds effectively facilitates the
+implementation of the required robotic architecture where perception and high-level
+reasoning (including complex, human-aware reasoning) have to be tightly
+integrated, and yet are developed across multiple partners with different
+expertise.
+
+%enables to get
+%knowledge about what the customer sees (or has seen) or not and to handle the
+%task accordingly. We plan also to be able to add information coming from the
+%dialogue (e.g. if the customer says that he has seen a particular landmark that
+%could consequently be added to his belief world). Moreover, we plan also to use
+%alternative world ability given by \uwds (as explained in \ref{alter}) (e.g. to
+%create a world where we put the customer in front of the desired shop to infer
+%(and be able to explain him) what he will see at destination).
 

### 2018-02-28 paper.tex
Save a bit of space by removing/cropping some figures

@@
     the client connects to the \uwds network, blocks until the world {\tt
-    example\_in} changes, and only propagate nodes that match the condition to
-    the world {\tt example\_out}.}
+    world1} changes, and only propagate nodes that match the condition to
+    the world {\tt world2}.}
 
@@
     \includegraphics[width=\linewidth]{tools}
-    \caption{Screenshots of the timeline (top) and network (bottom)
-    introspection tools. Worlds are represented as boxes; clients as ellipses.}
+    \caption{Screenshot of the network topology introspection tool. Worlds are represented as boxes; clients as ellipses.}
     \label{fig|explorer}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{spatialrelations}
-    \caption{The {\tt spatial\_relations} client computes perspective-aware
-    spatial relations between objects and agents: \emph{allo-centric} relations (like
-    {\tt is in} or {\tt is on}) are independent of the viewpoint, while
-    \emph{ego-centric} relations ({\tt in front of}, {\tt left of}) depend on
-    the viewer perspective.}
-    \label{fig|spatialrelations}
-\end{figure}
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=\linewidth]{spatialrelations}
+%    \caption{The {\tt spatial\_relations} client computes perspective-aware
+%    spatial relations between objects and agents: \emph{allo-centric} relations  are independent of the viewpoint, while
+%    \emph{ego-centric} relations ({\tt in front of}, {\tt left of}) depend on
+%    the viewer perspective.}
+%    \label{fig|spatialrelations}
+%\end{figure}
 
 \uwds provides an exemplary client ({\tt spatial\_relations}) to compute both
-allo-centric and ego-centric (\ie viewer-dependent) spatial relations
-between objects (Figure~\ref{fig|spatialrelations}).
+allo-centric (independent of the viewpoint like {\tt isIn} or {\tt isOn}) and
+ego-centric (\ie viewer-dependent, like {\tt inFrontOf} or {\tt leftOf}) spatial relations
+between objects.
+%(Figure~\ref{fig|spatialrelations}).
 Other libraries, like QSRLib~\cite{gatsoulis2016qsrlib}, that implement

### 2018-02-28 paper.tex
Moved figures around to have them closer to the references in text
Rewritten the paragraph on the physics filter

@@
 \subsection{Distributed Situation Assessment}
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=\linewidth]{overview}
-    \caption{Schema of a possible \uwds network: eight \emph{clients} (user-written \&
-    architecture specific; in blue) are sharing environment
-    models through four independent \emph{worlds} (made from joint spatial and
-    temporal models). This architecture enables successive and modular
-    refinement of the models (\emph{cascading} situation assessment),
-    effectively adapted to each client's needs.}
-
-    \label{fig|scene}
-
-\end{figure}
 
@@
 \label{design}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=\linewidth]{overview}
+    \caption{Schema of a possible \uwds network: eight \emph{clients} (user-written \&
+    architecture specific; in blue) are sharing environment
+    models through four independent \emph{worlds} (made from joint spatial and
+    temporal models). This architecture enables successive and modular
+    refinement of the models (\emph{cascading} situation assessment),
+    effectively adapted to each client's needs.}
+
+    \label{fig|scene}
+
+\end{figure}
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{uwds-screenshot}
-    \caption{Screenshot of the {\tt uwds view} 3D visualisation and manipulation
-    client. In this particular example, the 3D meshes have been pre-loaded using
-    {\tt uwds load}. Their positions are then updated at run-time using the
-    robot's sensors and proprioception (joint state).}
-    \label{fig|uwds-view}
-\end{figure}
-
 \subsection{Distributed spatio-temporal models}
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{uwds-screenshot}
+    \caption{Screenshot of the {\tt uwds view} 3D visualisation and manipulation
+    client. In this particular example, the 3D meshes have been pre-loaded using
+    {\tt uwds load}. Their positions are then updated at run-time using the
+    robot's sensors and proprioception (joint state).}
+    \label{fig|uwds-view}
+\end{figure}
+
 Interestingly, while \uwds deals with 3D geometries and scenes, it does
@@
 
-When perception modules provide objects localization, the consistency is not guarantee, objects on a table seem to slightly flight in the air, when dropping them in a bin the localization is not updated any more when the objects are occulted, resulting in objects that fly just above the bin.
+When perception modules provide objects localisation, the physical consistency
+of the locations is not typically enforced. For instance, objects that are
+supposed to lay on a table might be slightly above (or inset into) the table; or
+when dropping an item into a box, the robot can not update the location of the
+item anymore as it becomes occluded.
 
-Being able to manage that issues by simulating the gravity is the purpose of a dedicated client  
-{\tt physics filter} that have been designed using Bullet RT physics simulation and pybullet \url{https://pybullet.org/} library, it provides from an input world, an output world where objects described by an URDF file are filtered by the simulator, which gives us the near future of the objects based on their position, mass and inertia.
+These issues can be alleviated by relying on a physics simulation to stabilise
+the position of objects: natural physics (including gravity) are simulated for a
+short amount of time (up to one second) ahead of time, and the objects'
+positions are updated accordingly. To this end, \uwds provide a standard filter,
+the {\tt physics\_filter}, based on the Bullet RT physics simulation and the
+{\tt pybullet}\footnote{\url{https://pybullet.org/}} library. It generates an
+output world that mirrors its input world after a specific duration of physics
+simulation,  the physical properties of objects (including mass, friction,
+inertia) being provided from standard URDF descriptions.
 
@@
 
-\subsection{Application to human-robot interaction and spatial reasoning}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{l2tor-photo2-blind}
-    \caption{A user manipulating objects on top of an RFID mat (providing object
-    localisation) with a NAO robot, while an operator monitors
-    the interaction on a tablet that displays an {\tt uwds view} client [photo blurred for blind review].}
-
-    \label{fig|l2torexample}
-\end{figure}
-
-\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor}
-in order to conceptualise and visualise the spatial relations and visibility of
-the physical objects that participants interact with. In one of the scenarios of this
-project, participants are instructed by a NAO robot to manipulate a number of
-Duplo-like animals on top of a printed background paper which serves as a
-landscape (Figure~\ref{fig|l2torexample}). The robot instructions are based on
-pedagogical strategies to promote spatial relation learning by object
-manipulation. Such instructions include commands to place the elephant
-\textit{on top} or \textit{next to} the cube, or put the giraffe \textit{inside}
-the house.  To assess the performance, the robot and the operator need to
-compute spatial relationships between objects from the visual perspective of the
-participant.
-
 \subsection{Application to joint-action situations}
@@
 %(and be able to explain him) what he will see at destination).
+
+\subsection{Application to human-robot interaction and spatial reasoning}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{l2tor-photo2-blind}
+    \caption{A user manipulating objects on top of an RFID mat (providing object
+    localisation) with a NAO robot, while an operator monitors
+    the interaction on a tablet that displays an {\tt uwds view} client [photo blurred for blind review].}
+
+    \label{fig|l2torexample}
+\end{figure}
+
+\uwds is being used within the European L2TOR project~\cite{belpaeme2015l2tor} as part of a system for a large scale study with 150 children.
+\uwds {\tt spatial\_relations} module is used to visualise counting tasks and teach spatial words for a second language.
+\uwds's loose coupling allows for easy integration with the {\tt spatial_relations} between either an object recognition system that would allow for use in a physical environment(Figure~\ref{fig|l2torexample})
+
+%\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor} in order to conceptualise and visualise the spatial relations and visibility of the physical objects that participants interact with.
+
+%In one of the scenarios of this project, participants are instructed by a NAO robot to manipulate a number of Duplo-like animals on top of a printed background paper which serves as a landscape (Figure~\ref{fig|l2torexample}).
+The robot instructions are based on pedagogical strategies to promote spatial relation learning by object manipulation.
+Such instructions include commands to place the elephant \textit{on top} or \textit{next to} the cube, or put the giraffe \textit{inside} the house.
+To assess the performance, the robot and the operator need to compute spatial relationships between objects from the visual perspective of the participant.
 

### 2018-02-28 paper.tex
Update on Overleaf.

@@
 
-When perception modules provide objects localisation, the physical consistency
-of the locations is not typically enforced. For instance, objects that are
-supposed to lay on a table might be slightly above (or inset into) the table; or
-when dropping an item into a box, the robot can not update the location of the
-item anymore as it becomes occluded.
+When perception modules provide objects localization, the consistency is not guarantee, objects on a table seem to slightly flight in the air, when dropping them in a bin the localization is not updated any more when the objects are occulted, resulting in objects that fly just above the bin.
 
-These issues can be alleviated by relying on a physics simulation to stabilise
-the position of objects: natural physics (including gravity) are simulated for a
-short amount of time (up to one second) ahead of time, and the objects'
-positions are updated accordingly. To this end, \uwds provide a standard filter,
-the {\tt physics\_filter}, based on the Bullet RT physics simulation and the
-{\tt pybullet}\footnote{\url{https://pybullet.org/}} library. It generates an
-output world that mirrors its input world after a specific duration of physics
-simulation,  the physical properties of objects (including mass, friction,
-inertia) being provided from standard URDF descriptions.
+Being able to manage that issues by simulating the gravity is the purpose of a dedicated client  
+{\tt physics filter} that have been designed using Bullet RT physics simulation and pybullet \url{https://pybullet.org/} library, it provides from an input world, an output world where objects described by an URDF file are filtered by the simulator, which gives us the near future of the objects based on their position, mass and inertia.
 
@@
 \uwds {\tt spatial\_relations} module is used to visualise counting tasks and teach spatial words for a second language.
-\uwds's loose coupling allows for easy integration with the {\tt spatial_relations} between either an object recognition system that would allow for use in a physical environment(Figure~\ref{fig|l2torexample})
+\uwds's loose coupling allows for easy integration with the {\tt spatial_relations} between either an object recognition system that would allow for use in a physical environment(
 

### 2018-02-28 paper.tex
Update on Overleaf.

@@
 \uwds {\tt spatial\_relations} module is used to visualise counting tasks and teach spatial words for a second language.
-\uwds's loose coupling allows for easy integration with the {\tt spatial_relations} between either an object recognition system that would allow for use in a physical environment(
+\uwds's loose coupling allows for easy integration with the {\tt spatial_relations} between either an object recognition system that would allow for use in a physical environment(Figure~\ref{fig|l2torexample})
 

### 2018-02-28 paper.tex
Update on Overleaf.

--- 2018-02-28 paper.tex
@@
 
-\section{Application Example}
+\section{Application Examples}
 \label{application}
@@
 \uwds is being used within the European L2TOR project~\cite{belpaeme2015l2tor} as part of a system for a large scale study with 150 children.
-\uwds {\tt spatial\_relations} module is used to visualise counting tasks and teach spatial words for a second language.
-\uwds's loose coupling allows for easy integration with the {\tt spatial_relations} between either an object recognition system that would allow for use in a physical environment(Figure~\ref{fig|l2torexample})
+\uwds {\tt spatial\_relations} module is used to visualise counting tasks and teach spatial words for a second language(l2).
+The loose coupling provided by \uwds allows for easy integration with the {\tt spatial\_relations} between either an object localisation system that would allow for use in a physical environment (Figure~\ref{fig|l2torexample}) or a virtual tablet game.
+As well this loose coupling facilitated development on multiple parts of the software individually across multiple international sites.
+
+In the counting task the child is instructed to move toy animals into enclosures.
+The {\tt spatial\_relations} module calculates when the robots have been placed in the enclosure.
+The robot then teaches the children l2 words for the numbers and encourages the children to count along.
+
+While teaching the spatial domain the
 
 %\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor} in order to conceptualise and visualise the spatial relations and visibility of the physical objects that participants interact with.
-
-%In one of the scenarios of this project, participants are instructed by a NAO robot to manipulate a number of Duplo-like animals on top of a printed background paper which serves as a landscape (Figure~\ref{fig|l2torexample}).
+In one of the scenarios of this project, participants are instructed by a NAO robot to manipulate a number of Duplo-like animals on top of a printed background paper which serves as a landscape (Figure~\ref{fig|l2torexample}).
 The robot instructions are based on pedagogical strategies to promote spatial relation learning by object manipulation.

### 2018-03-01 paper.tex
Update on Overleaf.

@@
 
-\author{Séverin Lemaignan$^{1}$, Yoan Sallami$^{2}$, Christopher Wallbridge$^{1}$, Aurélie Clodic$^{2}$ and Rachid Alami$^{2}$%
+\author{S\'everin Lemaignan$^{1}$, Yoan Sallami$^{2}$, Christopher Wallbridge$^{1}$, Aur\'elie Clodic$^{2}$ and Rachid Alami$^{2}$%
 \thanks{$^{1}$Authors are with Centre for Robotics and Neural Systems,
@@
 
-\uwds is specifically inspired by geometric and temporal reasoners like {\sc Spark}
-(\emph{SPAtial Reasoning \& Knowledge})~\cite{sisbot2011situation}.  {\sc Spark}
+\uwds (as well as {\sc TOASTER} (\emph{Tracking Of Agents and Spatio-TEmporal Reasoning})~\cite{milliezframework2014}) is specifically inspired by geometric and temporal reasoners like {\sc Spark}
+(\emph{SPAtial Reasoning \& Knowledge})~\cite{sisbot2011situation}.{\sc Spark}
 acts as a situation assessment reasoner that generates symbolic knowledge from
@@
 \uwds is being used within the European L2TOR project~\cite{belpaeme2015l2tor} as part of a system for a large scale study with 150 children.
-\uwds {\tt spatial\_relations} module is used to visualise counting tasks and teach spatial words for a second language(l2).
+The \uwds {\tt spatial\_relations} module is used to visualise counting tasks and teach spatial words for a second language(l2).
 The loose coupling provided by \uwds allows for easy integration with the {\tt spatial\_relations} between either an object localisation system that would allow for use in a physical environment (Figure~\ref{fig|l2torexample}) or a virtual tablet game.
-As well this loose coupling facilitated development on multiple parts of the software individually across multiple international sites.
+As well, this loose coupling facilitated development on multiple parts of the software individually across multiple international sites.
 
 In the counting task the child is instructed to move toy animals into enclosures.
-The {\tt spatial\_relations} module calculates when the robots have been placed in the enclosure.
+The {\tt spatial\_relations} module calculates when the robots have been placed in the enclosure, and passes this on to another part of the system.
 The robot then teaches the children l2 words for the numbers and encourages the children to count along.
-
-While teaching the spatial domain the
+While teaching the spatial domain the robot ask the children to move objects into positions related to other objects (e.g.
+put the box on the shelf).
+The performance can be evaluated by the {\tt spatial\_relations} module between objects.
+In both tasks the robot is able to provide targeted feedback on the position of an animal or object as the spatial relation is calculated rather than just attempting to fulfil a particular position criteria
 

### 2018-03-01 paper.tex
[minor] toaster ref

@@
 
-\uwds (as well as {\sc TOASTER} (\emph{Tracking Of Agents and Spatio-TEmporal Reasoning})~\cite{milliezframework2014}) is specifically inspired by geometric and temporal reasoners like {\sc Spark}
-(\emph{SPAtial Reasoning \& Knowledge})~\cite{sisbot2011situation}.{\sc Spark}
-acts as a situation assessment reasoner that generates symbolic knowledge from
-the geometry of the environment with respect to relations between objects,
-robots and humans. It also takes into account the different perspective that
-each agent has on the environment. {\sc Spark} embeds a modality-independent
-geometric model of the environment that serves both as basis for the fusion of
-the perception modalities and as bridge with the symbolic
-layer~\cite{lemaignan2016artificial}. This geometric model is built from 3D CAD
-models of the objects, furniture and robots, and full body, rigged models of
-humans. It is updated at run-time by the robot's sensors.  Likewise, \uwds
-embeds a grounded amodal model of the environment, updated online from the
-robot's sensors (sensor fusion).
+\uwds  is specifically inspired by geometric and temporal reasoners like {\sc
+Spark} (\emph{SPAtial Reasoning \& Knowledge})~\cite{sisbot2011situation} or
+{\sc toaster} (\emph{Tracking Of Agents and Spatio-TEmporal
+Reasoning})~\cite{milliezframework2014}. {\sc Spark} acts as a situation
+assessment reasoner that generates symbolic knowledge from the geometry of the
+environment with respect to relations between objects, robots and humans. It
+also takes into account the different perspective that each agent has on the
+environment. {\sc Spark} embeds a modality-independent geometric model of the
+environment that serves both as basis for the fusion of the perception
+modalities and as bridge with the symbolic layer~\cite{lemaignan2016artificial}.
+This geometric model is built from 3D CAD models of the objects, furniture and
+robots, and full body, rigged models of humans. It is updated at run-time by the
+robot's sensors.  Likewise, \uwds embeds a grounded amodal model of the
+environment, updated online from the robot's sensors (sensor fusion).
 

### 2018-03-01 paper.tex
Commented out L2TOR example, partially rewritten MuMMER description

@@
 
-\section{Application Examples}
+\section{Application Example: Perspective-aware joint actions}
 \label{application}
 
-\subsection{Application to joint-action situations}
+%\subsection{Application to joint-action situations}
 
-\uwds is used within the European project MuMMER for service robots to compute
-visibility and knowledge about objects, places and agents within a mall
+\uwds is being used within the large European project
+MuMMER\footnote{\url{http://www.mummer-project.eu}} for service robots to
+compute visibility and knowledge about objects, places and agents within a mall
 environment.
@@
 at a table and have to coordinate joint actions (pick and place).  One object on
-the table (the green box) is only visible to one human and the robot, but hidden
-to the second human. The robot needs to take into account this fact to generate
-appropriate and legible joint manipulation actions.
-Figure~\ref{fig|mummerarchitecture} illustrates the topology of the \uwds
-network that we use to this end.
+the table (the green box in Figure~\ref{fig|mummerarchitecture}) is only visible
+to one human and the robot, but hidden to the second human. The robot needs to
+take into account this fact to generate appropriate and legible joint
+manipulation actions.  Figure~\ref{fig|mummerarchitecture} illustrates the
+topology of the \uwds network that we use to this end.
 
@@
 
-A first client, \textit{env\_provider}, provides the environment models and
+A first client, \textit{static\_env\_provider}, provides the environment models and
 allows to build a first \texttt{ENV} world where static objects, furnitures and
 walls are present.  Then, three worlds cascade through three (independent)
-clients: \textit{robot\_state\_monitor} augments {\tt ENV} with the robot state
+clients: \textit{robots\_state\_monitor} augments {\tt ENV} with the robot state
 (using underneath the ROS robot state publisher node) and broadcast a new world
@@
 ar\_track\_alvar}\footnote{\url{http://wiki.ros.org/ar_track_alvar}}).
-\textit{object\_monitor} finally detects and adds the humans
-(using~\cite{Khalidov_Idiap-RR-02-2017}). It broadcasts a world called {\tt
-BASE} that eventually contains the static environment, the robots, the dynamic
+\textit{humans\_monitor} finally detects and continuously updates the humans
+poses (using~\cite{Khalidov_Idiap-RR-02-2017}). It broadcasts a world called {\tt
+BASE} that contains as a result the static environment, the robots, the dynamic
 objects and the detected humans.
 
-The world \texttt{BASE} goes through a \textit{physics filter} client as
-explained in section~\ref{physics} to obtain the \texttt{STABLE} world where all
+The world \texttt{BASE} goes through a \textit{physics filter} client (as
+explained in section~\ref{physics}) to obtain the \texttt{STABLE} world where all
 elements are present with physically-consistent locations. This
@@
 visible to the agent at a given point in time. As such, it does not only offer a
-snapshot of the agent visual perspective at the current time but also act as the
+snapshot of the agent visual perspective at the current time but also acts as the
 visual memory of each agent.
@@
 \uwds makes it possible to implement such a geometric reasoning pipeline in a
-fully decoupled way, and each intermediary ``world'' can be easily introspected
-at run-time. As such, \uwds facilitates the implementation and debugging of complex
-spatio-temporal reasoning pipelines.
+fully decoupled way, and each intermediary world can be easily introspected at
+run-time. This example shows how \uwds facilitates the implementation and
+debugging of complex spatio-temporal reasoning pipelines.
 
@@
 
-We are currently deploying a similar network in the framework of the
-European project MuMMER where a Pepper robot handles interactive
-situations in a large shopping centre in Finland. One of the situation is a guiding task where
-Pepper will help people to find their route by pointing them landmarks and
-explaining them how to reach a destination. \uwds effectively facilitates the
-implementation of the required robotic architecture where perception and high-level
-reasoning (including complex, human-aware reasoning) have to be tightly
-integrated, and yet are developed across multiple partners with different
-expertise.
+We are currently deploying a similar network in the framework of the European
+project MuMMER where a Pepper robot handles interactive situations in a large
+shopping centre in Finland. One of the situation is a guiding task where Pepper
+help people to find their route by pointing them landmarks and explaining them
+how to reach a destination. To be effevtive, this helping behaviour needs to be
+aware of the visual perspective of the human. \uwds facilitates the
+implementation of such a spatio-temporal reasoning pipeline, where perception
+and high-level reasoning (including complex, human-aware reasoning) have to be
+tightly integrated. Because of the decoupling of each of the clients in the
+network, \uwds also practically supports software development spread across
+multiple partners in different countries, with different expertise.
 
@@
 
-\subsection{Application to human-robot interaction and spatial reasoning}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{l2tor-photo2-blind}
-    \caption{A user manipulating objects on top of an RFID mat (providing object
-    localisation) with a NAO robot, while an operator monitors
-    the interaction on a tablet that displays an {\tt uwds view} client [photo blurred for blind review].}
-
-    \label{fig|l2torexample}
-\end{figure}
-
-\uwds is being used within the European L2TOR project~\cite{belpaeme2015l2tor} as part of a system for a large scale study with 150 children.
-The \uwds {\tt spatial\_relations} module is used to visualise counting tasks and teach spatial words for a second language(l2).
-The loose coupling provided by \uwds allows for easy integration with the {\tt spatial\_relations} between either an object localisation system that would allow for use in a physical environment (Figure~\ref{fig|l2torexample}) or a virtual tablet game.
-As well, this loose coupling facilitated development on multiple parts of the software individually across multiple international sites.
-
-In the counting task the child is instructed to move toy animals into enclosures.
-The {\tt spatial\_relations} module calculates when the robots have been placed in the enclosure, and passes this on to another part of the system.
-The robot then teaches the children l2 words for the numbers and encourages the children to count along.
-While teaching the spatial domain the robot ask the children to move objects into positions related to other objects (e.g.
-put the box on the shelf).
-The performance can be evaluated by the {\tt spatial\_relations} module between objects.
-In both tasks the robot is able to provide targeted feedback on the position of an animal or object as the spatial relation is calculated rather than just attempting to fulfil a particular position criteria
-
-%\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor} in order to conceptualise and visualise the spatial relations and visibility of the physical objects that participants interact with.
-In one of the scenarios of this project, participants are instructed by a NAO robot to manipulate a number of Duplo-like animals on top of a printed background paper which serves as a landscape (Figure~\ref{fig|l2torexample}).
-The robot instructions are based on pedagogical strategies to promote spatial relation learning by object manipulation.
-Such instructions include commands to place the elephant \textit{on top} or \textit{next to} the cube, or put the giraffe \textit{inside} the house.
-To assess the performance, the robot and the operator need to compute spatial relationships between objects from the visual perspective of the participant.
+%\subsection{Application to human-robot interaction and spatial reasoning}
+%
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=\linewidth]{l2tor-photo2-blind}
+%    \caption{A user manipulating objects on top of an RFID mat (providing object
+%    localisation) with a NAO robot, while an operator monitors
+%    the interaction on a tablet that displays an {\tt uwds view} client [photo blurred for blind review].}
+%
+%    \label{fig|l2torexample}
+%\end{figure}
+%
+%\uwds is being used within the European L2TOR project~\cite{belpaeme2015l2tor}
+%to implement the symbolic spatial reasoning pipeline of a system used in a large
+%scale study (150 children). \uwds supports 
+%easy integration with the {\tt spatial\_relations} between either an object
+%localisation system that would allow for use in a physical environment
+%(Figure~\ref{fig|l2torexample}) or a virtual tablet game. 
+%
+%The \uwds\xspace {\tt spatial\_relations} module is
+%used to visualise counting tasks and teach spatial words for a second language.
+%
+%In the counting task the child is instructed to move toy animals into
+%enclosures. The {\tt spatial\_relations} module calculates when the objects have
+%been placed in the enclosure, and passes this on to another part of the system.
+%The robot then teaches the children l2 words for the numbers and encourages the
+%children to count along. While teaching the spatial domain the robot ask the
+%children to move objects into positions related to other objects (\eg put the
+%box on the shelf). The performance can be evaluated by the {\tt
+%spatial\_relations} module between objects. In both tasks the robot is able to
+%provide targeted feedback on the position of an animal or object as the spatial
+%relation is calculated rather than just attempting to fulfil a particular
+%position criteria
+%
+%The loose coupling provided by As well, this loose
+%coupling facilitated development on multiple parts of the software individually
+%across multiple international sites.
+%
+%%\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor}
+%%in order to conceptualise and visualise the spatial relations and visibility of
+%%the physical objects that participants interact with. In one of the scenarios
+%%of this project, participants are instructed by a NAO robot to manipulate a
+%%number of Duplo-like animals on top of a printed background paper which serves
+%%as a landscape (Figure~\ref{fig|l2torexample}). The robot instructions are
+%%based on pedagogical strategies to promote spatial relation learning by object
+%%manipulation. Such instructions include commands to place the elephant
+%%\textit{on top} or \textit{next to} the cube, or put the giraffe
+%%\textit{inside} the house.  To assess the performance, the robot and the
+%%operator need to compute spatial relationships between objects from the visual
+%%perspective of the participant.
 

### 2018-03-01 paper.tex
Rachid comments

@@
 
-A third example relates to human-robot interaction scenarios where \emph{perspective
-taking} is important (a prototypical example being the game `I spy with my
-little eye', as implemented in~\cite{ros2010which}). Perspective taking is a
-cognitive skill that relies on the ability for an agent to take someone else's
-point of view to estimate what they see from their perspectives. 
-Perspective taking has previously been implemented in robotics by temporarily placing
-virtual cameras at eye locations for each of the humans tracked by the
+A third example relates to human-robot interaction scenarios where
+\emph{perspective taking} is important (a prototypical example being the game `I
+spy with my little eye', as implemented in~\cite{ros2010which}). Perspective
+taking is a cognitive skill that relies on the ability for an agent to take
+someone else's point of view to estimate what they see from their perspectives.
+Perspective taking has previously been implemented in robotics by temporarily
+placing virtual cameras at eye locations for each of the humans tracked by the
 robot~\cite{ros2010solving}. While acceptable for simple cases, such an approach
-does not maintain truly independent models of the environment for each
-agent, and in particular does not permit false-belief
-modelling~\cite{lemaignan2015mutual} whereas separate, independent world models
-would effectively support such a skill.
+does not maintain truly independent spatio-temporal models of the environment
+for each agent, and in particular, it does not permit the representation of
+proper false-belief situations. On the contrary, separate, independent world
+models as implemented by \uwds effectively support such a skill, which is an
+important precursor to research and implement human's mind modelling (\ie a
+theory of mind)~\cite{lemaignan2015mutual}.
 
@@
     \caption{Schema of the \uwds architecture used in the MuMMER project.
-    Clients (blue boxes) read and generate worlds (whose names are in typewriter
-    fonts).}
+    Clients (blue boxes) read and generate the worlds whose names are in typewriter
+    fonts. The last two worlds, {\tt HUMAN\{1,2\}\_PERSPECTIVE}, do not only
+    represent the immediate visual perspective of the two humans, but integrate
+    as well their past visual perceptions. As such, they are the \emph{visual
+    memories} of the humans, that the robot can rely on when making decisions.}
     \label{fig|mummerarchitecture}

### 2018-03-01 paper.tex
Update on Overleaf.

@@
 
+\uwds distinguishes itself from DyKnow framework ~\cite{de_leng_dyknow:_2016} that have investigated stream reasoning as key a feature for spatial and temporal reasoning because DyKnow aim to address a large variety of data types, which is why auto configurable systems~\cite{leng_towards_2017} based on the semantic are useful where \uwds only distribute one type of geometric(scene tree) and temporal(timeline) models of the environment as explained above.
+
 \subsection{Future work}

### 2018-03-01 paper.tex
Reworked (simplified, really) reference to DyKnow

--- 2018-03-01 paper.tex
@@
 Other libraries, like QSRLib~\cite{gatsoulis2016qsrlib}, that implement
-computational models of Qualitative Spatial Relations could be trivially
+computational models of Qualitative Spatial Relations, could be trivially
 combined with \uwds to provide more advanced geometric analysis.
+Future developments will also include the results of the more basic
+research on spatio-temporal reasoning for robotics, led by de Leng and
+Heintz~\cite{deleng2016qualitative}.
 
@@
 computation based on message passing. However, it distinguishes itself from
-existing middleware in significant ways. Most importantly, \uwds purposefully
-does not offer any \emph{general} capability to distribute computation and data
-flow amongst independent components: it focusses specifically on distributing
-environment models, both spatial (geometric models) and temporal (events and
-situations). In that sense, \uwds really is a \emph{distributed datastructure}
-that addresses the specific needs of spatio-temporal modelling,
-including the modelling of hypothetical, alternative world models, something
-that traditional middlewares like ROS do not address adequately.
+existing middlewares (including ROS extensions like
+DyKnow~\cite{de_leng_dyknow:_2016}) in significant ways. Most importantly,
+\uwds purposefully does not offer any \emph{general} capability to distribute
+computation and data streams amongst independent components: it focusses
+specifically on distributing environment models, both spatial (geometric
+models) and temporal (events and situations). In that sense, \uwds really is a
+\emph{distributed datastructure} that addresses the specific needs of
+spatio-temporal modelling, including the modelling of hypothetical, alternative
+world models, something that traditional middlewares like ROS do not address
+adequately. Second, and as presented above, \uwds offers specific mechanisms
+for the representation and manipulation of alternative world models that are
+not directly achievable with traditional tools.
 
@@
 or gRPC (the later being the one used by \uwds).
+
+%\uwds also differs from the DyKnow framework~\cite{de_leng_dyknow:_2016}.
+%DyKnow investigates stream reasoning as a central concept for spatial and
+%temporal modelling, with a recent focus on semantic-aware auto-configurable
+%systems~\cite{leng_towards_2017}. \uwds
+%only distribute one type of spatial (scene graph) and temporal (timeline)
+%datastructures, as explained above.
 

### 2018-07-10 paper.tex
Update on Overleaf.
Added Tony as coauthor + updated affiliations

@@
 
-\author{S\'everin Lemaignan$^{1}$, Yoan Sallami$^{2}$, Christopher Wallbridge$^{1}$, Aur\'elie Clodic$^{2}$ and Rachid Alami$^{2}$%
-\thanks{$^{1}$Authors are with Centre for Robotics and Neural Systems,
-        Plymouth University, Plymouth PL48AA, United Kingdom
-        {\tt\small firstname.surname@plymouth.ac.uk},
-	$^{2}$Authors are with LAAS-CNRS, Université de Toulouse, CNRS, Toulouse, France
-        {\tt\small firstname.surname@laas.fr}}%
+\author{S\'everin Lemaignan$^{1}$, Yoan Sallami$^{2}$, Christopher
+Wallbridge$^{3}$,\\
+Aur\'elie Clodic$^{2}$, Tony Belpaeme$^{3}$, and Rachid Alami$^{2}$%
+\thanks{$^{1}$Author is with Bristol Robotics Lab, University of the West of
+England, Bristol, United Kingdom
+        {\tt\small severin.lemaignan@brl.ac.uk},
+       $^{2}$Authors are with LAAS-CNRS, Universit\'e de Toulouse, CNRS, Toulouse, France
+        {\tt\small firstname.surname@laas.fr},
+       $^{3}$Authors are with CRNS, Plymouth University, Plymouth, United Kingdom
+        {\tt\small firstname.surname@plymouth.ac.uk}}%
 }
@@
 help people to find their route by pointing them landmarks and explaining them
-how to reach a destination. To be effevtive, this helping behaviour needs to be
+how to reach a destination. To be effective, this helping behaviour needs to be
 aware of the visual perspective of the human. \uwds facilitates the

### 2018-07-13 paper.tex
Final changes, taking into account the reviewer's comments

@@
 datastructures. In typical usage scenarios (where the number of nodes or events
-remain relatively small), the computational load to manipulate these
+remain under a few hundred relatively small), the computational load to manipulate these
 datastructures is however dominated by the actual processings performed by the
-clients with the data.
+clients with the data. In the current implementation, scene graphs and timelines
+are stored in-memory. Were they required, serialization and persistent storage
+are not anticipated to be difficult to implement.
 
@@
     \includegraphics[width=\linewidth]{tools}
-    \caption{Screenshot of the network topology introspection tool. Worlds are represented as boxes; clients as ellipses.}
+    \caption{Screenshot of the network topology introspection tool, with
+    arbitrary examples of worlds (represented as boxes) and clients (ellipses).
+    CLients are connected to the worlds either as
+    \emph{readers} or \emph{providers} of data. \uwds introspection features
+    make it possible to also visualise when clients were last active.}
     \label{fig|explorer}
@@
     \caption{Schema of the \uwds architecture used in the MuMMER project.
-    Clients (blue boxes) read and generate the worlds whose names are in typewriter
-    fonts. The last two worlds, {\tt HUMAN\{1,2\}\_PERSPECTIVE}, do not only
-    represent the immediate visual perspective of the two humans, but integrate
-    as well their past visual perceptions. As such, they are the \emph{visual
+    Clients read and generate the worlds {\tt ENV} $\rightarrow$ {\tt
+    ENV\_ROBOT} $\rightarrow$ \ldots $\rightarrow$ {\tt
+    HUMAN*\_PERSPECTIVE}. The last two worlds {\tt HUMAN\{1,2\}\_PERSPECTIVE}
+    represent the immediate visual perspective of each of the humans, as well as
+    their past visual perceptions. As such, they are the \emph{visual
     memories} of the humans, that the robot can rely on when making decisions.}
